### 2023-05-08 paper-2023-acsos-pulverization.tex
rename the main file

\documentclass[conference]{IEEEtran}
\IEEEoverridecommandlockouts
% The preceding line is only needed to identify funding in the first footnote. If that is unneeded, please comment it out.
\usepackage{cite}
\usepackage{amsmath,amssymb,amsfonts}
\usepackage{algorithmic}
\usepackage{graphicx}
\usepackage{textcomp}
\usepackage{xcolor}
\def\BibTeX{{\rm B\kern-.05em{\sc i\kern-.025em b}\kern-.08em
    T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}
\begin{document}

\title{Declarative runtime reconfiguration of pulverised distributed systems
\thanks{Identify applicable funding agency here. If none, delete this.}
}

\author{\IEEEauthorblockN{1\textsuperscript{st} Danilo Pianini}
\IEEEauthorblockA{\textit{Department of Computer Science and Engineering} \\
\textit{Alma Mater Studiorum---Università di Bologna}\\
Cesena, Italy \\
danilo.pianini@unibo.it}
\and
\IEEEauthorblockN{2\textsuperscript{nd} Roberto Casadei}
\IEEEauthorblockA{\textit{Department of Computer Science and Engineering} \\
\textit{Alma Mater Studiorum---Università di Bologna}\\
Cesena, Italy \\
email address or ORCID}
\and
\IEEEauthorblockN{3\textsuperscript{rd} Nicolas Farabegoli}
\IEEEauthorblockA{\textit{Department of Computer Science and Engineering} \\
\textit{Alma Mater Studiorum---Università di Bologna}\\
Cesena, Italy \\
email address or ORCID}
\and
\IEEEauthorblockN{4\textsuperscript{th} Mirko Viroli}
\IEEEauthorblockA{\textit{Department of Computer Science and Engineering} \\
\textit{Alma Mater Studiorum---Università di Bologna}\\
Cesena, Italy \\
email address or ORCID}
}

\maketitle

\begin{abstract}
TBD
\end{abstract}

\begin{IEEEkeywords}
runtime reconfiguration, distributed systems, self-adaptation, self-organisation, pulverisation
\end{IEEEkeywords}

\section{Introduction}

\nocite{*}

\bibliographystyle{IEEEtran}
\bibliography{bibliography}

\vspace{12pt}
\color{red}
IEEE conference templates contain guidance text for composing and formatting conference papers. Please ensure that all template text is removed from your conference paper prior to submission to the conference. Failure to remove the template text from your paper may result in your paper not being published.

\end{document}

### 2023-05-09 paper-2023-acsos-pulverization.tex
stub an abstract

@@
 \begin{abstract}
-TBD
+In recent years, we witnessed a radical change in the deployed form of distributed systems.
+%
+Modern applications are designed to be executed on very diverse devices
+and to be deployed on heterogeneous infrastructures,
+ranging from cloud servers to mobile and IoT devices.
+%
+De facto, complex distributed computation is happening across a cloud-edge continuum composed of heterogeneous devices and infrastructures.
+%
+The availablity of such an infrastructure opens to new possibilities in terms of better resource utilisation and performance,
+but also poses new challenges to the application designer,
+as the application must be conceived to be able to adapt its deployment to changing conditions.
+%
+In this paper we present a framework for the development of distributed systems
+based on the concept of \emph{pulverisation},
+which is meant to neatly separate business logic and deployment concerns,
+allowing applications to be defined independent of the infrastructure they will execute upon.
+%
+The framework is based on a domain-specific language capturing,
+in a declarative fashion,
+the concepts composing a pulverised application and any potential target network,
+also allowing for the specification of reconfiguration policies.
+%
+The framework, implemented in Kotlin multiplatform and available as open source,
+is then evaluated in a small-scale real-world demo and in a city-scale simulated scenario.
 \end{abstract}
@@
 
-\section{Introduction}
+\section{Introduction}\label{sec:introduction}
 
-\nocite{*}
+\begin{itemize}
+    \item need for deployment-independent specs
+    \item need for runtime reconfiguration (achieve QoS in face of changing conditions --- green computing?)
+    \item declarativity over imperative approaches (note: can be compared with the trend in general softeng/build systems)
+\end{itemize}
+
++ related work
+
+\section{A framework for declarative runtime reconfiguration}\label{sec:contribution}
+
+\subsection{Domain model}
+
+\subsection{Domain-specific language}
+
+\subsection{Implementation details}
+
+Also mention that is multiplatform
+
+\section{Evaluation}\label{sec:evaluation}
+
+\subsection{Small scale: crowd alert}
+
+\subsection{Large scale: smart city application}
+\cite{PianiniJOS2013}
+
+\section{Conclusion and future work}\label{sec:conclusion}
+
+global policies
 

### 2023-05-09 paper-2023-acsos-pulverization.tex
Some rw from pulv paper

@@
     T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}
+
+\newcommand{\meta}[1]{{\color{blue}#1}}
+
 \begin{document}
@@
 Cesena, Italy \\
-email address or ORCID}
+roby.casadei@unibo.it}
 \and
@@
 
+\section{Related Work}
+
+\meta{
+%
+%\subsection{Approaches to Deployment Independence}
+%\label{s:rw:deployment-independence}
+%
+%\meta{TODO: add works on deployment independence, e.g., osmotic~\citep{DBLP:journals/computer/VillariFDRJR19}, BIP~\citep{lekidis2015bip-wsn,bastarrica2001optimization-techniques-deployment-components-bip}}
+%
+We highlight a number of representatieve research efforts on deployment and automatic reconfiguration of systems.
+%
+%For instance, 
+\emph{Osmotic computing}~\cite{DBLP:journals/computer/VillariFDRJR19} is an approach to opportunistic deployment of microservices on the edge-fog-cloud platform.
+An osmotic platform aims to reach and maintain an ``osmotic equilibrium'' 
+between infrastructural and application requirements
+by automatically migrating microservices to deployment locations.
+%
+However, the approach mainly targets centrally orchestrated systems.
+%
+Other approaches leverage component-based, architectural descriptions to decouple application logic and deployment.
+%
+For instance, \emph{DR-BIP (Dynamically Reconfigurable - Behaviour Interaction Priority model)}~\cite{,Ballouli18dr-bip}
+and \emph{DReAM (Dynamically Reconfigurable Architectural Modelling)}~\cite{denicola2020dream-dynamic-reconfig-arch-modelling}
+use \emph{components} (capturing behaviour), \emph{connectors} (capturing interaction between components' \emph{ports}), 
+\emph{maps} (logical topologies),
+and \emph{deployments} (associating components to map locations),
+overall organised in \emph{motifs} (dynamic architectural configurations),
+to model and analyse dynamic architectures.
+%
+
+These approaches have some similarity with the approach presented in this paper, but they are arguably more complex and our  
+%than the one introduced in this paper,
+ %, though also fostering an exogenous, declarative (logics-based) modelling approach.
+approach explicitly addresses self-organising CPS.
+%
+%
+%By constrast, our approach explicitly addresses self-organising CPS.
+%
+%A more in-depth, formal comparison is left as future work.
+}
+
 \section{Conclusion and future work}\label{sec:conclusion}
@@
 
-\vspace{12pt}
-\color{red}
-IEEE conference templates contain guidance text for composing and formatting conference papers.
+%\vspace{12pt}
+%\color{red}
+%IEEE conference templates contain guidance text for composing and formatting conference papers.
 Please ensure that all template text is removed from your conference paper prior to submission to the conference.

### 2023-05-09 paper-2023-acsos-pulverization.tex
write some intro

@@
 
+\newcommand{\ourframework}{Porbione}
+
 \section{Introduction}\label{sec:introduction}
+
+Recent research trends in topics related to the Internet of Things promote the idea of a cloud-edge continuum~\cite{DBLP:journals/iot/BittencourtISFM18},
+an infrastructure composed of heterogeneous devices ranging from large and powerful cloud servers to small connected things
+on which software computes in a distributed fashion,
+optimising for performance and resource utilisation.
+%
+Exploiting this intriguing infrastructure, however, poses new challenges to the application designer.
+%
+Commonly, in fact, applications are designed with a specific infrastructure in mind,
+whose assumptions leak into the application logic.
+
+Previous work suggests a technique for decoupling business logic and deployment concerns
+called \emph{pulverization}~\cite{FI2020-pulverization}.
+%
+In short, the core idea is to consider devices as \emph{logical} entities,
+whose software can be designed as (or broken down to, if already existing) an ensemble of five micro-components
+(behaviour, state, communication, sensors, and actuators)
+whose actual host can be decided at deployment time.
+%
+This way, the designer can focus on the business logic at hand,
+delaying an optimised deployment to later stages of the sofware design,
+thus gaining resilience to changing infrastructures or non-functional requirements.
+%
+One limitation of the current pulverisation approach regards dynamicity:
+although the application can be deployed on arbitrary systems,
+there is no support for the reconfiguration of components in face of changing conditions or requirements.
+
+In this work,
+we continue the research on pulverised systems by providing two key contributions:
+first, we extend the theoretical framework of pulverisation by adding support for
+\emph{runtime configuration rules}, allowing pulverised components to move at runtime across different hosts;
+the providing a practical implementation called \ourframework{} which,
+to the best of our knowledge,
+is the first \emph{practical framework supporting pulverisation} for a general-purpose language.
+%
+We exercise \ourframework{} with two case studies:
+a small-scale real-wold demonstration, and a city-scale simulated scenario.
 

### 2023-05-09 paper-2023-acsos-pulverization.tex
related survey on reconfiguration verification
docs: description of the first dsl

@@
 \usepackage{xcolor}
+\usepackage{listings}
+\usepackage{listing-style}
+
+\usepackage{cleveref}
+
 \def\BibTeX{{\rm B\kern-.05em{\sc i\kern-.025em b}\kern-.08em
@@
 Cesena, Italy \\
-email address or ORCID}
+nicolas.farabegoli@unibo.it}
 \and
@@
 
+The framework allows the system configuration via two distinct DSLs, one for defining the system's devices and their components,
+and one for defining the runtime configuration of the system.
+
+The first domain-specific language allows to define the system in terms of
+\emph{devices}, \emph{components} and which \emph{capabilities} requires to be executed.
+%
+In this way, infrastructural and deployment aspects are abstracted, allowing the focus to be on functional aspects of the system.
+%
+The DSL enables the definition of the type of devices that took part in the system,
+and for each device can be specified which components constitute it and where those component must run on (e.g. which host must execute the component).
+%
+The execution of a component is constrained to hosts with predetermined characteristics:
+the framework models the constraint with the \texttt{Capability} concept,
+namely the property that a host must to ``implement'' to run the component.
+%
+Different capabilities can be specified for each component to provide a more flexible and scalable deployment of the system.
+
+The~\Cref{lst:system-configuration} provides an example of the usage of the domain-specific language where is defined a system composed of two kind of devices.
+%
+The DSL returns the system's configuration that will be used by the runtime.
+
+One key-point of the domain-specific language can be observed in the definition of the \texttt{device-2} device:
+the \emph{Behaviour} component can be deployed on hosts that implement the \texttt{HighCPU} and \texttt{EmbeddedDevice} capabilities,
+in this way the behaviour can be moved (at runtime) from a host to another one that implements the same capabilities.
+%
+The association between a component and a single capability do not necessarily imply that the component can be executed only on a single host,
+but it is the way to specify the minimum requirements for the host that will execute the component.
+
+\lstinputlisting[
+    float=ht,
+    language=Kotlin,
+    caption={TODO},
+    label=lst:system-configuration,
+]{listings/SystemConfig.kt}
+
+\lstinputlisting[
+    float=ht,
+    language=Kotlin,
+    caption={TODO},
+    label=lst:runtime-configuration,
+]{listings/RuntimeConfig.kt}
+
 \subsection{Implementation details}
@@
 \section{Related Work}
+
+A recent survey on 
+ formal techniques for verifying the correctness of reconfigurations
+ in component-based distributed software systems is provided in~\cite{coullon2023swreconfig}.
 

### 2023-05-09 paper-2023-acsos-pulverization.tex
Descripbed runtime DSL

@@
 
+The second domain-specific language allows to define the concrete implementation of each component associating it to a specific start host;
+moreover, it allows to define the creteria on which the system should be reconfigured by specifying which events should trigger a reconfiguration.
+%
+This DSL binds the more abstrac definition specified in the previous DSL to the topology of the available infrastracture,
+defining the actual deployment of the system, and the reconfiguration policies.
+%
+The~\Cref{lst:runtime-configuration} provides an example of the usage of the domain-specific language where is defined the runtime configuration of the system.
+%
+The framework define the concept of \texttt{Host} that is the representation of an available physical host where the components can be deployed;
+for each host is possible to specify the capabilities that it implements.
+%
+In this way, the framework can check if a host can execute a component by verifying if the host implements the required capabilities.
+%
+The runtime is configured be specifying for each implementation of a \emph{component},
+its associated startup host enabling the definition of the initial deployment of the system.
+%
+The most relevant part of the domain-specific language, is the definition of the reconfiguration rules:
+the framework models the concept of \texttt{ReconfigurationEvent}
+that is the event that triggers the reconfiguration of the system.
+%
+The DSL allows the association of a \texttt{ReconfigurationEvent}
+to a new deployment of the system which is defined as the new association between component and host.
+
 \lstinputlisting[
@@
 ]{listings/RuntimeConfig.kt}
+
+\lstinputlisting[
+    float=ht,
+    language=Kotlin,
+    caption={TODO},
+    label=lst:reconfiguration-events,
+]{listings/ReconfigurationEvents.kt}
 

### 2023-05-10 paper-2023-acsos-pulverization.tex
Add metrics

@@
     language=Kotlin,
-    caption={TODO},
+    caption={
+        Example of the usage of the domain-specific language for the system configuration.
+        The example configures two types of devices, and for each device defines the components that compose it and which capabilities requires for the execution.
+    },
     label=lst:system-configuration,
@@
     language=Kotlin,
-    caption={TODO},
+    caption={
+        Example of the usage of the domain-specific language for the runtime configuration.
+        The example defines the runtime configuration of the system, specifying the initial deployment and the reconfiguration rules.
+    },
     label=lst:runtime-configuration,
 ]{listings/RuntimeConfig.kt}
+
+A \texttt{ReconfigurationEvent} is characterized by the \emph{events}
+that it produces and the \emph{condition} that must be satisfied to trigger the reconfiguration.
+%
+The \emph{events} are represented as an asyncronous stream of valued produced by the device and the
+\emph{condition} is a predicate that is evaluated on the stream of events.
+%
+At the first match of the condition, the reconfiguration is triggered and the system is reconfigured with the new deployment.
+%
+The~\Cref{lst:reconfiguration-events} shows an example of the definition of a reconfiguration event.
 
@@
     language=Kotlin,
-    caption={TODO},
+    caption={Example of the definition of a reconfiguration event.},
     label=lst:reconfiguration-events,
@@
 \section{Evaluation}\label{sec:evaluation}
+
+The pulverization approach tries to solve the problem of the deployment independence of the system, in heterogeneous infrastructures.
+%
+The dynamic reconfiguration of the system is a key point to achieve flexibility and adaptability of the system on the requirements changing.
+%
+In this section are presented two case studies that show the effectiveness of the approach:
+the first case study shows the operation of the framework in a small-scale scenario, while the second case study simulate a large-scale scenario
+using the framework in a simulated environment showing the effectiveness of the approach in many different conditions.
+
+\subsection{Metrics}
+\meta{descrizione}
+
+As follow are reported the free variables used in the evaluation:
+\begin{itemize}
+    \item let $n$ be the number of devices;
+    \item let $t_d$ be the discharge time of the battery of the devices;
+    \item let $B_l$ be the low battery threshold;
+    \item let $B_h$ be the high battery threshold;
+    \item let $EPI_{cloud}$ be the energy per instruction of the cloud;
+    \item let $EPI_{ratio}$ be the ratio of the consumption of the cloud and the device;
+\end{itemize}
+
+From the variables above are derived the following metrics:
+
+\begin{itemize}
+    \item let $\beta_{device}$ be the number of \emph{behaviour} executed in the device;
+    \item let $\beta_{cloud}$ be the number of \emph{behaviour} executed in the cloud;
+    \item let $C_{\%}$ be the average battery percentage of the devices;
+    \item let $P_{devices}$ be the consumed power of the devices;
+    \item let $P_{cloud}$ be the consumed power of the cloud;
+    \item let $P_{total}$ be the total consumed power of the system;
+    \item let $I_{cloud}$ be the number of server instances of the cloud;
+    \item let $\$_{cloud}$ be the cost of the cloud; 
+\end{itemize}
 

### 2023-05-10 paper-2023-acsos-pulverization.tex
More related?

@@
 
+A related approach is also provided by the so-called \emph{multi-tier programming} paradigm~\cite{DBLP:journals/csur/WeisenburgerWS20},
+ whereby the code for multiple tiers (e.g., client and server tiers; or view, business logic, and data tiers) is specified in a single codebase,
+ from which the code of each individual tier is compiled.
+
 \meta{
+MuScADeL (Multi-Scale Autonomic Deployment Language)~\cite{DBLP:conf/compsac/BoujbelRLTAL14/muscadel}
+
+ADSL (AWaRE DSL)~\cite{DBLP:conf/aswec/ChhetriLUVKNR18/adsl} % and see corresp RW
+
 %

### 2023-05-10 paper-2023-acsos-pulverization.tex
introduce eval

@@
 \usepackage{listing-style}
+\usepackage[hidelinks]{hyperref}
 
@@
 
-The pulverization approach tries to solve the problem of the deployment independence of the system, in heterogeneous infrastructures.
+In this section,
+we exercise the proposed framework with two case studies.
 %
-The dynamic reconfiguration of the system is a key point to achieve flexibility and adaptability of the system on the requirements changing.
+In the former,
+we show a small-scale demonstrator of a safety system warning people when a room capacity is exceeded
+to showcase the integration of the framework with established technologies\footnote{
+    \url{https://github.com/nicolasfara/ACSOS-2023-pulverization-crowd-room}
+};
+in the latter,
+we simulate a city-scale distributed application deployed for a urban event
+and show how the dynamic reconfiguration may help achieving balance between energy consumption and QoS\footnote{
+    \url{https://github.com/nicolasfara/ACSOS-2023-pulverization-evaluation}
+}.
 %
-In this section are presented two case studies that show the effectiveness of the approach:
-the first case study shows the operation of the framework in a small-scale scenario, while the second case study simulate a large-scale scenario
-using the framework in a simulated environment showing the effectiveness of the approach in many different conditions.
+Both experiments have been released with a permissive open-source license and have been archived for future reference on Zenodo~\cite{TBD}.
 
-\subsection{Metrics}
-\meta{descrizione}
+\subsection{Small scale: crowd alert}
+
+\subsection{Large scale: smart city application}
+
+\cite{PianiniJOS2013}
+
+\subsubsection{Free variables}
 
@@
 \end{itemize}
+
+\subsubsection{Metrics}
 
@@
 
-\subsection{Small scale: crowd alert}
-
-\subsection{Large scale: smart city application}
-\cite{PianiniJOS2013}
+\subsubsection{Results}
 

### 2023-05-10 paper-2023-acsos-pulverization.tex
Comment on rw

@@
 
-ADSL (AWaRE DSL)~\cite{DBLP:conf/aswec/ChhetriLUVKNR18/adsl} % and see corresp RW
+ADSL (AWaRE DSL)~\cite{DBLP:conf/aswec/ChhetriLUVKNR18/adsl} % and see corresp RW: the most related DSLs are those in the area of automated deployment [17].
+Such DSLs include Deladas [31], [32], ConfSolve [15], [33], j-ASD’s DSL [34] and MuScADeL [35], [36].
+% [17]: J.-P.
+Arcangeli, R.
+Boujbel, and S.
+Leriche, "Automatic deployment of distributed software systems: Definitions and state of the art," Journal of Systems and Software, vol.
+103, pp.
+198–218, 2015
 

### 2023-05-11 paper-2023-acsos-pulverization.tex
Add initial description of the small case study

@@
 \usepackage[hidelinks]{hyperref}
+\usepackage{svg}
 
@@
 \subsection{Small scale: crowd alert}
+\label{sec:small-scale-crowd-alert}
+
+In this section we present a simple scenario where the pulverization framework is used to implement a crowd alert system:
+we simulate a laboratory where a safety system warns people when they are too close to each other.
+%
+Each person is equipped with a wearable device that is able to detect the distance from other devices and send the distances data to the
+alert system.
+%
+The people in the laboratory must observe a safety distance, and the system warns them when the distance is not respected by showing a red light
+when they are too close to each other, and a green color when the distance is respected (see~\Cref{fig:crowd-alert}).
+%
+\begin{figure*}[ht]
+    \centering
+    \includesvg[width=.8\textwidth]{figures/crow-laboratory-demo.drawio.svg}
+    \caption{
+        Representation of the scenario described in the~\Cref{sec:small-scale-crowd-alert}
+        where more the people are far from each other, more the color of the screen is green.
+        More the people are close to each other, more the color of the screen is red.
+        The blue dot represents the wearable device and the link between the devices represents the distance between them.
+        The rectangle represents the monitor in the laboratory which shows the congestion level of the laboratory (via the color described above).
+    }
+    \label{fig:crowd-alert}
+\end{figure*}
+%
+The system is composed of two kind of devices: \emph{wearable} and \emph{laboratory}.
+%
+The former represents the wearable devices that is worn by the people, and consists of the following pulverized components:
+$\beta$, $\chi$, and $\sigma$.
+%
+The \emph{behaviour} ($\beta$) component is responsible for converting the raw data from the sensor into the distance
+between the device and the others; the \emph{sensors} ($\sigma$) component is responsible for perceiving the other devices and
+the \emph{communication} ($\chi$) component is responsible for sending the data to the alert system.
+%
+The latter represents the alert system and consists of the following pulverized components:
+$\beta$, $\kappa$, $\chi$, and $\alpha$.
+%
+The \emph{behaviour} ($\beta$) component is responsible for evaluating the meand distance between the devices,
+the \emph{actuators} ($\alpha$) is responsible for triggering the alarm showing a color in the screen,
+and the \emph{communication} ($\chi$) component is responsible for receiving the data from the wearable devices.
+
+The infrastructure defined for this scenario is described as follow:
+
+\begin{itemize}
+    \item \emph{Smartphones} are the devices used by the people to detect the distance from the others;
+    \item \emph{Local PC} is the device used to show the alarm in the laboratory;
+    \item \emph{Server} is the device used run the crowd algorithm and the wearable logic. 
+\end{itemize}
+
+The initial deployment of the system starts with the $\beta_{werable}$ executed in the \emph{Server},
+the $\sigma_{wearable}$ and $\chi_{werable}$ executed in the \emph{Smartphones},
+and all the \emph{Laboratory} components are executed in the \emph{Local PC}.
+%
+Given the infrastracture defined above, we simulate also a condition where the \emph{Server} is overloaded and, via pulverization,
+we reconfigure the execution of all the $\beta_{werable}$ component to the \emph{Smartphone}.
+%
 

### 2023-05-11 paper-2023-acsos-pulverization.tex
Structuring related

@@
 \usepackage{listings}
+\usepackage{acronym}
 \usepackage{listing-style}
@@
 \usepackage{svg}
+
+\acrodef{dsl}[DSL]{domain-specific language}
 
@@
 
-The framework allows the system configuration via two distinct DSLs, one for defining the system's devices and their components,
+The framework allows the system configuration via two distinct \acp{dsl}, one for defining the system's devices and their components,
 and one for defining the runtime configuration of the system.
@@
 
-A recent survey on 
- formal techniques for verifying the correctness of reconfigurations
- in component-based distributed software systems is provided in~\cite{coullon2023swreconfig}.
+This work provides 
+ \acp{dsl}
+ for specifying 
+ distributed systems and
+ corresponding deployment plans,
+ and a framework for executing
+ deployment plans and reconfigurations.
+%
+Therefore, this work deals 
+ with multiple problems,
+ investigated by multiple research threads.
+%
 
-A related approach is also provided by the so-called \emph{multi-tier programming} paradigm~\cite{DBLP:journals/csur/WeisenburgerWS20},
- whereby the code for multiple tiers (e.g., client and server tiers; or view, business logic, and data tiers) is specified in a single codebase,
- from which the code of each individual tier is compiled.
+%
+The approach of specifying in a single codebase 
+ parts of the structure, behaviour, or interaction 
+ for a whole distributed system
+ is also related to \emph{macroprogramming}~\cite{Casadei2023macro}
+ and, in particular, to \emph{multi-tier programming} paradigm~\cite{DBLP:journals/csur/WeisenburgerWS20},
+ where the deployment units for different tiers (e.g., client and server tiers; or view, business logic, and data tiers) are obrained by compilation or interpretation of a single codebase.
+
+\subsection{Application description languages: component models and pulverisation}
+%
+
+\subsection{Infrastructure and deployment description languages}
+
+\subsection{Automatic deployment and reconfiguration}
+%
+A comprehensive conceptual framework
+ on automatic deployment of distributed systems
+ is provided in a 2015 survey by Arcangeli et al. \cite{DBLP:journals/jss/ArcangeliBL15}.
+%
+A more recent survey by Coullon et al. focusses
+ on formal techniques for verifying the correctness of reconfigurations
+ in component-based distributed software systems~\cite{coullon2023swreconfig}.
+%
+In the \emph{osmotic computing} approach~\cite{DBLP:journals/computer/VillariFDRJR19},
+ microservices (the solvent)
+ can migrate across 
+ the edge-cloud infrastructure (the solution),
+ passing through layer boundaries (the semi-permeable membranes),
+ in order to keep a balance in the desired properties (the solute).
 
@@
 %
-We highlight a number of representatieve research efforts on deployment and automatic reconfiguration of systems.
-%
-%For instance, 
-\emph{Osmotic computing}~\cite{DBLP:journals/computer/VillariFDRJR19} is an approach to opportunistic deployment of microservices on the edge-fog-cloud platform.
-An osmotic platform aims to reach and maintain an ``osmotic equilibrium'' 
-between infrastructural and application requirements
-by automatically migrating microservices to deployment locations.
+%We highlight a number of representatieve research efforts on deployment and automatic reconfiguration of systems.
+%%
+%%For instance, 
+%\emph{Osmotic computing}~\cite{DBLP:journals/computer/VillariFDRJR19} is an approach to opportunistic deployment of microservices on the edge-fog-cloud platform.
+%An osmotic platform aims to reach and maintain an ``osmotic equilibrium'' 
+%between infrastructural and application requirements
+%by automatically migrating microservices to deployment locations.
 %

### 2023-05-11 paper-2023-acsos-pulverization.tex
Complete small demo description

@@
 Given the infrastracture defined above, we simulate also a condition where the \emph{Server} is overloaded and, via pulverization,
-we reconfigure the execution of all the $\beta_{werable}$ component to the \emph{Smartphone}.
+we want to lighteninig the load of the \emph{Server} by moving the $\beta_{werable}$ in the \emph{Smartphones}.
 %
+Formally, we can write the reconfiguration event as follow: $\beta^{werable}_{server} \xrightarrow{\mathcal{E}} \beta^{werable}_{smartphone}$
+where $\mathcal{E}$ represents the reconfiguration event due to the high load in the \emph{Server}.
+
+At the start of the system, each \emph{Werable} device, with the $\sigma$ component,
+start to perceive the other devices in the laboratory using the RSSI signal from the Bluetooth.
+%
+The $\beta$ component, initially executed in the \emph{Server}, starts to compute the distances from the other devices
+based on the received RSSI signal from the $\sigma$ component.
+%
+Once the distances are computed, the $\chi$ component sends the data to the
+\emph{Laboratory} containing the dinstances perceived by the device.
+%
+
+Similarity, the \emph{Laboratory} device has all the components executed in the \emph{Local PC},
+where the $\chi$ component receives the data from the \emph{Wearable},
+the $\beta$ component computes the mean distance between the devices.
+%
+The $\kappa$ component persist in memory the updated devices grpah based on the distances retreived by the $\chi$ component.
+and the $\alpha$ component show in the monitor the color based on the mean distance computed by the $\beta$ component.
+
+As soon as the reconfiguration event is triggered, the $\beta$ component of the \emph{Werable} is moved from the \emph{Server} to the \emph{Smartphone}
+lighteninig the server and preserving from one side the functional semantic of the system and from the other side maintain the system operational.
+%
+The reconfiguration of the system in entirely managed by the framework, which handle all the machinery and communication to move the component from
+one host to another.
+%
+With this simple example we show how the pulverization framework can be used in practice to manage a system with different devices and an heterogeneous
+infrastructure.
+%
+Moreover, we show how the pulverization framework can be used adapt the system to changing requirement without loosing the funcional semantic of the system
+by describing the system in a declarative way.
 

### 2023-05-11 paper-2023-acsos-pulverization.tex
Structuring rw sec

@@
 Therefore, this work deals 
- with multiple problems,
+ with multiple problems and perspectives,
  investigated by multiple research threads.
 %
+In~\cite{DBLP:journals/jss/ArcangeliBL15},
+the problem of (automatic) deployment of distributed software system is studied by considering
+ (i) the nature of software to be deployed,
+ (ii) the nature of the deployment domain (i.e. the available infrastructure),
+ (iii) how the deployment is designed,
+ and 
+ (iv) how the deployment is performed.
+%
+Accordingly, 
+ we cover related works about
+ languages for 
+ specifying applications (\Cref{sec:rw:whatsw}),
+ languages for specifying infrastructures and deployments (\Cref{sec:rw:depdesc}),
+ and
+ techniques for automatic deployment and reconfiguration (\Cref{sec:rw:autodep}).
+
+\subsection{Application description languages: component models and pulverisation}
+\label{sec:rw:whatsw}
+%
+A comprehensive survey on component-based software engineering is provided in~\cite{vale2016component-based-se}.
 
@@
 
-\subsection{Application description languages: component models and pulverisation}
-%
-
 \subsection{Infrastructure and deployment description languages}
+\label{sec:rw:depdesc}
 
 \subsection{Automatic deployment and reconfiguration}
+\label{sec:rw:autodep}
 %

### 2023-05-11 paper-2023-acsos-pulverization.tex
Use pdf instead svg
rename section
refine abstract

@@
 \usepackage[hidelinks]{hyperref}
-\usepackage{svg}
 
@@
 \begin{abstract}
-In recent years, we witnessed a radical change in the deployed form of distributed systems.
+In recent years, we witnessed a fast-paced evolution of the infrastructure hosting distributed computations.
 %
 Modern applications are designed to be executed on very diverse devices
-and to be deployed on heterogeneous infrastructures,
-ranging from cloud servers to mobile and IoT devices.
+and to be deployed on infrastructures composed of devices ranging from cloud servers to mobile and IoT devices,
+constituting together a cloud-edge continuum.
 %
-De facto, complex distributed computation is happening across a cloud-edge continuum composed of heterogeneous devices and infrastructures.
+The availability of such an infrastructure opens to better resource utilisation and performance,
+but, contextually, introduces new challenges to software designers,
+as applications must be conceived to be able to adapt their deployment to changing conditions.
 %
-The availablity of such an infrastructure opens to new possibilities in terms of better resource utilisation and performance,
-but also poses new challenges to the application designer,
-as the application must be conceived to be able to adapt its deployment to changing conditions.
-%
-In this paper we present a framework for the development of distributed systems
+In this paper,
+we introduce a framework for the development of distributed systems
 based on the concept of \emph{pulverisation},
-which is meant to neatly separate business logic and deployment concerns,
+meant to neatly separate business logic and deployment concerns,
 allowing applications to be defined independent of the infrastructure they will execute upon.
@@
 in a declarative fashion,
-the concepts composing a pulverised application and any potential target network,
-also allowing for the specification of reconfiguration policies.
+pulverised application components, device capabilities, resource allocation, and (runtime re-) configuration policies.
 %
@@
     \centering
-    \includesvg[width=.8\textwidth]{figures/crow-laboratory-demo.drawio.svg}
+    \includegraphics[width=.8\textwidth]{figures/crow-laboratory-demo.drawio.pdf}
     \caption{
@@
 
-\subsection{Large scale: smart city application}
+\subsection{Large scale: urban event}
 

### 2023-05-11 paper-2023-acsos-pulverization.tex
better introduce pulverisation and glue with the DSL part

@@
 \usepackage[hidelinks]{hyperref}
+\usepackage[inline]{enumitem}
 
 \acrodef{dsl}[DSL]{domain-specific language}
+\acrodef{iot}[IoT]{Internet of Things}
 
@@
 delaying an optimised deployment to later stages of the sofware design,
-thus gaining resilience to changing infrastructures or non-functional requirements.
+thus gaining resilience to changes in the infrastructure or in non-functional requirements.
 %
@@
 
-\subsection{Domain model}
+The goal of pulverisation~\cite{FI2020-pulverization}
+is to provide application designers a way to specify the functional semantics of their software in a deployment-independent way.
+%
+To do so,
+the application logic should be designed considering a \emph{logical system},
+namely, a network of \emph{logical} devices forming a network with \emph{arbitrary} topology.
+%
+For a pulverisation system to work,
+the application of every single logical device must be decomposable into an ensemble of \emph{pulverised components}
+representing, respectively:
+\begin{itemize}
+    \item a set $\sigma$ of logical sensors;
+    \item a set $\alpha$ of logical actuators;
+    \item a state $\kappa$, representing the logical device's knowledge;
+    \item a communication component $\chi$,
+    handling interaction with reachable devices in the logical system, and
+    \item a computation component $\beta$, modelling the behaviour of the logic device.
+\end{itemize}
+%
+Decomposition of an application into pulverised components can be achieved in two ways:
+either the application is designed with pulverisation in mind,
+or the application is developed using a framework supporting automatic decomposition
+(one notable example are aggregate computing frameworks~\cite{BealIEEEComputer2015}).
+%
+Once the application has been pulverised,
+a mapping must be provided between the pulverised logical system and the hosts that will execute the pulverised components.
+%
+In this process,
+a single logical component could be (and usually does) end up being executed on multiple hosts.
 
-\subsection{Domain-specific language}
+For instance, let us assume a minimal logical system counting two devices:
+two rain gauges logging the water level for future reference,
+and opening a valve when the water level crosses a threshold in both devices.
+%
+Once pulverised, the logical system would be split in ten deployable pulverised components
+(that we indicate with their component symbol and an index, e.g., $\sigma_1$ is the logical sensor of the first device).
+%
+These pulverised components can be deployed in arbitrary hosts,
+as far as they have the \emph{capabilities} required to host them
+(for instance, the component hosting a rain gauge logical sensor must have an actual sensor,
+while the component hosting the behaviour must be sufficiently powerful to execute the logic).
+%
+Thus, with \emph{no change to the application logic},
+the system could be deployed on very different systems:
+let us assume, for instance,
+that our actual target infrastructure is not composed of two connected devices,
+but it is an existing \ac{iot} system
+where the rain gauges are connected to two LoRaWAN motes
+(which are too weak and energy-critical to execute the logic),
+the valves are controlled by a dedicated device,
+the levels must be logged on a database hosted on a cloud server reachable via HTTPS,
+and we have an internal edge server that we can use as we please.
+%
+In this case, we would deploy $\sigma_1$ and $\sigma_2$ on the motes,
+$\alpha_1$ and $\alpha_2$ on the valve controller,
+$\kappa_1$ and $\kappa_2$ on the cloud server,
+and the remaining components on the edge server.
 
-The framework allows the system configuration via two distinct \acp{dsl}, one for defining the system's devices and their components,
-and one for defining the runtime configuration of the system.
+\meta{maybe import the image from the pulverisation paper?}
+
+\subsection{Domain-specific languages for pulverisation}
+
+From the previous section,
+we can identify two main phases in the pulverisation process:
+\begin{enumerate*}[label=\it{(\roman*)}]
+  \item the definition of the available devices and their capabilities; and
+  \item the allocation of the pulverised components to them.
+\end{enumerate*}
+%
+We decided to capture these phases into two distinct \acp{dsl}.
 

### 2023-05-11 paper-2023-acsos-pulverization.tex
RW: more on app description and pulverisation

@@
 %
-Therefore, this work deals 
- with multiple problems and perspectives,
- investigated by multiple research threads.
+Therefore, we provide an original solution
+ to different issues.
 %
@@
  languages for 
- specifying applications (\Cref{sec:rw:whatsw}),
+ specifying distributed systems (\Cref{sec:rw:whatsw}),
  languages for specifying infrastructures and deployments (\Cref{sec:rw:depdesc}),
@@
 
-\subsection{Application description languages: component models and pulverisation}
+\subsection{Application description languages: component-based software engineering and the pulverisation model}
 \label{sec:rw:whatsw}
 %
-A comprehensive survey on component-based software engineering is provided in~\cite{vale2016component-based-se}.
+%% A first distinction revolves around whether 
+%% the distributed software to be deployed 
+%% is a monolith 
+%% or is already broken into different components.
+%
+We consider a distributed application 
+ as a graph of deployable units.
+%
+This partitioning may be 
+ manually specified at development time
+ (e.g., by explicitly defining and packaging different components)
+ or automatically defined
+ through \emph{application partitioning} approaches~\cite{DBLP:journals/jnca/LiuASGBQ15}. %exploiting the natural modularity of code (depending on the paradigm and language adopted).
 
+In our approach, the application partitioning into components is manually specified through a \ac{dsl}.
 %
-The approach of specifying in a single codebase 
- parts of the structure, behaviour, or interaction 
+Therefore, it can be framed in the context of \emph{component-based software engineering},
+ with ~\cite{vale2016component-based-se} providing a comprehensive review.
+%
+The specification or design of distributed systems
+ may leverage
+ component models~\cite{DBLP:journals/tse/CrnkovicSVC11},
+ architectural description languages~\cite{DBLP:journals/tse/MedvidovicT00},
+ or frameworks.
+%
+
+Specifically,
+ this work generalises over the \emph{pulverisation model}~\cite{FI2020-pulverization,IEEE-IoTJ-pulverization-simulation},
+ whereby a large-scale cyber-physical system
+ is partitioned into a graph of devices
+ where each device is split into five deployable components:
+ (i) sensors interface, 
+ (ii) actuators interface, 
+ (iii) behaviour, 
+ (iv) state, and 
+ (v) communication component.
+%
+In~\cite{IEEE-IoTJ-pulverization-simulation},
+ a methodology on top of the pulverisation model
+ is proposed
+ for generating and testing deployments through simulation.
+
+Also,
+ the approach of specifying in a single codebase 
+ parts of the structure, behaviour, and/or interaction 
  for a whole distributed system
@@
 %
+\meta{TODO: say something about our positioning wrt verification of reconfigurations}
+
 In the \emph{osmotic computing} approach~\cite{DBLP:journals/computer/VillariFDRJR19},

### 2023-05-12 paper-2023-acsos-pulverization.tex
Meta with limit and deadline
More related work

@@
 \newcommand{\ourframework}{Porbione}
+
+\meta{LIMIT 10 pages incl. refs; DEADLINE 19 may}
 
@@
  for specifying 
- distributed systems and
+ distributed systems (partitioned according to the 
+ pulverisation model)
+ and
  corresponding deployment plans,
@@
 %
-Therefore, we provide an original solution
- to different issues.
-%
-In~\cite{DBLP:journals/jss/ArcangeliBL15},
-the problem of (automatic) deployment of distributed software system is studied by considering
+A way to structure all this 
+ is by using the analytical framework of~\cite{DBLP:journals/jss/ArcangeliBL15},
+  where the problem of (automatic) deployment of distributed software system is studied by considering
  (i) the nature of software to be deployed,
@@
 Specifically,
- this work generalises over the \emph{pulverisation model}~\cite{FI2020-pulverization,IEEE-IoTJ-pulverization-simulation},
+ this work builds on the \emph{pulverisation} component model~\cite{FI2020-pulverization,IEEE-IoTJ-pulverization-simulation},
  whereby a large-scale cyber-physical system
@@
 
-Also,
+Additionally,
  the approach of specifying in a single codebase 
@@
 \label{sec:rw:depdesc}
+
+A \emph{deployment plan} is a configured mapping of a software system onto a deployment domain.
+%
+Given a deployable software system developed using the techniques of the previous subsection,
+ what is needed now is a language to describe a deployment domain 
+ and the deployment mapping.
+%
+Languages have been proposed 
+ targetting specific kinds of deployment domains
+ such as the smart grid (cf.
+the deployment specification language \emph{dspec} in the \emph{Resilient Information Architecture Platform for Smart Grids [RIAPS]} platform~\cite{DBLP:conf/coins/GhoshTKKL22})
+ or the cloud (cf. the \emph{Cloud Application Modelling and Execution Language [CAMEL]} multi-\ac{dsl}~\cite{DBLP:journals/jcloudc/AchilleosKRKDOS19}).
+
+In this work,
+ we are especially interested deployments over large-scale cyber-physical systems and the edge-fog-cloud continuum.
+%
+The \emph{MultiScale Autonomic Deployment Language (MuScADeL)}~\cite{DBLP:conf/compsac/BoujbelRLTAL14/muscadel}
+ is a \ac{dsl} that allows to express deployment properties of 
+ componentised applications on multi-scale deployment domains, i.e., large and heterogeneous infrastructures considered under several viewpoints (device, geography, network, administrative domains, etc.).
+%
+In~\cite{DBLP:journals/sosym/SongDFSF22},
+  a model-based deployment approach is proposed
+  targetting so-called \emph{fleets}, i.e.,
+  distributed and heterogeneous devices at the edge chaaracterised by different cyber-physical contexts (cf. resources, connectivity, etc.).
+%
+It is based on the \emph{Generation and Deployment
+of Smart IoT Systems (GeneSIS)} modelling language, where a \texttt{DeploymentModel} consists of sets of \texttt{Resources} (e.g., \texttt{Component}s, specialised by \texttt{InfrastructureComponent}s and \texttt{SoftwareComponent}s) with associated \texttt{Property}s and \texttt{Link}s.
 
@@
 %
+Two formal models for specifying reconfigurable architectures are \emph{DR-BIP (Dynamically Reconfigurable - Behaviour Interaction Priority model)}~\cite{,Ballouli18dr-bip}
+and \emph{DReAM (Dynamically Reconfigurable Architectural Modelling)}~\cite{denicola2020dream-dynamic-reconfig-arch-modelling}.
+%
+Both are based on the same conceptual model: they 
+ feature \emph{components} (capturing behaviour), \emph{connectors} (capturing interaction between components' \emph{ports}), 
+\emph{maps} (logical topologies),
+and \emph{deployments} (associating components to map locations),
+overall organised in \emph{motifs} (dynamic architectural configurations),
+to model and analyse dynamic architectures.
+%
 \meta{TODO: say something about our positioning wrt verification of reconfigurations}
@@
 
-\meta{
-MuScADeL (Multi-Scale Autonomic Deployment Language)~\cite{DBLP:conf/compsac/BoujbelRLTAL14/muscadel}
-
-ADSL (AWaRE DSL)~\cite{DBLP:conf/aswec/ChhetriLUVKNR18/adsl} % and see corresp RW: the most related DSLs are those in the area of automated deployment [17].
+%\meta{
+%MuScADeL (Multi-Scale Autonomic Deployment Language)~\cite{DBLP:conf/compsac/BoujbelRLTAL14/muscadel}
+%
+%ADSL (AWaRE DSL)~\cite{DBLP:conf/aswec/ChhetriLUVKNR18/adsl} % and see corresp RW: the most related DSLs are those in the area of automated deployment [17].
 Such DSLs include Deladas [31], [32], ConfSolve [15], [33], j-ASD’s DSL [34] and MuScADeL [35], [36].
-% [17]: J.-P.
+%% [17]: J.-P.
 Arcangeli, R.
@@
 198–218, 2015
-
-%
-%\subsection{Approaches to Deployment Independence}
-%\label{s:rw:deployment-independence}
-%
-%\meta{TODO: add works on deployment independence, e.g., osmotic~\citep{DBLP:journals/computer/VillariFDRJR19}, BIP~\citep{lekidis2015bip-wsn,bastarrica2001optimization-techniques-deployment-components-bip}}
-%
-%We highlight a number of representatieve research efforts on deployment and automatic reconfiguration of systems.
 %%
-%%For instance, 
-%\emph{Osmotic computing}~\cite{DBLP:journals/computer/VillariFDRJR19} is an approach to opportunistic deployment of microservices on the edge-fog-cloud platform.
-%An osmotic platform aims to reach and maintain an ``osmotic equilibrium'' 
-%between infrastructural and application requirements
-%by automatically migrating microservices to deployment locations.
-%
-However, the approach mainly targets centrally orchestrated systems.
-%
-Other approaches leverage component-based, architectural descriptions to decouple application logic and deployment.
-%
-For instance, \emph{DR-BIP (Dynamically Reconfigurable - Behaviour Interaction Priority model)}~\cite{,Ballouli18dr-bip}
-and \emph{DReAM (Dynamically Reconfigurable Architectural Modelling)}~\cite{denicola2020dream-dynamic-reconfig-arch-modelling}
-use \emph{components} (capturing behaviour), \emph{connectors} (capturing interaction between components' \emph{ports}), 
-\emph{maps} (logical topologies),
-and \emph{deployments} (associating components to map locations),
-overall organised in \emph{motifs} (dynamic architectural configurations),
-to model and analyse dynamic architectures.
-%
-
-These approaches have some similarity with the approach presented in this paper, but they are arguably more complex and our  
-%than the one introduced in this paper,
- %, though also fostering an exogenous, declarative (logics-based) modelling approach.
-approach explicitly addresses self-organising CPS.
-%
-%
-%By constrast, our approach explicitly addresses self-organising CPS.
-%
-%A more in-depth, formal comparison is left as future work.
-}
+%%%
+%%%\subsection{Approaches to Deployment Independence}
+%%%\label{s:rw:deployment-independence}
+%%%
+%%%\meta{TODO: add works on deployment independence, e.g., osmotic~\citep{DBLP:journals/computer/VillariFDRJR19}, BIP~\citep{lekidis2015bip-wsn,bastarrica2001optimization-techniques-deployment-components-bip}}
+%%%
+%%%We highlight a number of representatieve research efforts on deployment and automatic reconfiguration of systems.
+%%%%
+%%%%For instance, 
+%%%\emph{Osmotic computing}~\cite{DBLP:journals/computer/VillariFDRJR19} is an approach to opportunistic deployment of microservices on the edge-fog-cloud platform.
+%%%An osmotic platform aims to reach and maintain an ``osmotic equilibrium'' 
+%%%between infrastructural and application requirements
+%%%by automatically migrating microservices to deployment locations.
+%%%
+%%However, the approach mainly targets centrally orchestrated systems.
+%%%
+%%Other approaches leverage component-based, architectural descriptions to decouple application logic and deployment.
+%%%
+%%For instance, \emph{DR-BIP (Dynamically Reconfigurable - Behaviour Interaction Priority model)}~\cite{,Ballouli18dr-bip}
+%%and \emph{DReAM (Dynamically Reconfigurable Architectural Modelling)}~\cite{denicola2020dream-dynamic-reconfig-arch-modelling}
+%%use \emph{components} (capturing behaviour), \emph{connectors} (capturing interaction between components' \emph{ports}), 
+%%\emph{maps} (logical topologies),
+%%and \emph{deployments} (associating components to map locations),
+%%overall organised in \emph{motifs} (dynamic architectural configurations),
+%%to model and analyse dynamic architectures.
+%%%
+%%
+%%These approaches have some similarity with the approach presented in this paper, but they are arguably more complex and our  
+%%%than the one introduced in this paper,
+%% %, though also fostering an exogenous, declarative (logics-based) modelling approach.
+%%approach explicitly addresses self-organising CPS.
+%%%
+%%%
+%%%By constrast, our approach explicitly addresses self-organising CPS.
+%%%
+%%%A more in-depth, formal comparison is left as future work.
+%}
 

### 2023-05-12 paper-2023-acsos-pulverization.tex
RW: Also mention service-based service composition

@@
 
-\newcommand{\ourframework}{Porbione}
+\newcommand{\ourframework}{\texttt{pulse-refrakt}}
 
@@
 \begin{enumerate*}[label=\it{(\roman*)}]
-  \item the definition of the available devices and their capabilities; and
-  \item the allocation of the pulverised components to them.
+  \item the definition of the available devices and their capabilities (infrastructure); and
+  \item the allocation of the pulverised components to them (runtime).
 \end{enumerate*}
@@
 
-The first domain-specific language allows to define the system in terms of
-\emph{devices}, \emph{components} and which \emph{capabilities} requires to be executed.
-%
-In this way, infrastructural and deployment aspects are abstracted, allowing the focus to be on functional aspects of the system.
+The infrastructure \ac{dsl} captures the concepts of
+\emph{logical device}, \emph{component}, \emph{capability}, and \emph{requirement}.
 %
@@
  architectural description languages~\cite{DBLP:journals/tse/MedvidovicT00},
+ service-based compositions~\cite{DBLP:journals/csur/LemosDB16}, 
  or frameworks.

### 2023-05-12 paper-2023-acsos-pulverization.tex
minor

@@
  architectural description languages~\cite{DBLP:journals/tse/MedvidovicT00},
- service-based compositions~\cite{DBLP:journals/csur/LemosDB16}, 
  or frameworks.

### 2023-05-12 paper-2023-acsos-pulverization.tex
RW: + AASL autonomic spec lang
RW: + ConfSolve + ADSL

--- 2023-05-12 paper-2023-acsos-pulverization.tex
@@
 
-\subsection{Automatic deployment and reconfiguration}
+\subsection{Automatic/autonomic deployment and reconfiguration}
 \label{sec:rw:autodep}
@@
  in order to keep a balance in the desired properties (the solute).
+%
+The \emph{ASSL (Autonomic System Specification Language)}~\cite{DBLP:conf/birthday/VassevH97}
+ supports the specification
+ of autonomic elements,
+ interaction protocols,
+ and \emph{self-management policies} based on fluents (states), events, and actions.
+%
+In \emph{ConfSolve}~\cite{DBLP:conf/saso/HewsonAG13},
+ the language supports the specification of logical \emph{reconfiguration constraints}, which are then solved as a constraint satisfaction problem.
+%
+The \emph{AWaRE DSL (ADSL)}~\cite{DBLP:conf/aswec/ChhetriLUVKNR18/adsl} also supports constraint-based self-management, leveraging managing agents.
+The language enables the specification of the domain model (in terms of components), the problem structure model (in terms of constraints), the agent architecture model (in terms of agents, roles, and their coordination), and the assignment model (in terms of management problem assignment strategies).
 

### 2023-05-12 paper-2023-acsos-pulverization.tex
RW: + DELADAS but somewhat shrinking

@@
 %
-In \emph{ConfSolve}~\cite{DBLP:conf/saso/HewsonAG13},
- the language supports the specification of logical \emph{reconfiguration constraints}, which are then solved as a constraint satisfaction problem.
+%In \emph{ConfSolve}~\cite{DBLP:conf/saso/HewsonAG13},
+% the language supports the specification of logical \emph{reconfiguration constraints}, which are then solved as a constraint satisfaction problem.
 %
-The \emph{AWaRE DSL (ADSL)}~\cite{DBLP:conf/aswec/ChhetriLUVKNR18/adsl} also supports constraint-based self-management, leveraging managing agents.
+The \emph{AWaRE DSL (ADSL)}~\cite{DBLP:conf/aswec/ChhetriLUVKNR18/adsl} supports constraint-based self-management, leveraging managing agents.
 The language enables the specification of the domain model (in terms of components), the problem structure model (in terms of constraints), the agent architecture model (in terms of agents, roles, and their coordination), and the assignment model (in terms of management problem assignment strategies).
+%
+Similar, earlier approaches based on the specification of \emph{constraint-sets} include \emph{DELADAS}~\cite{Dearle2004deladas}
+and \emph{ConfSolve}~\cite{DBLP:conf/saso/HewsonAG13}.
 

### 2023-05-12 paper-2023-acsos-pulverization.tex
Minor fix

@@
 %
-Two formal models for specifying reconfigurable architectures are \emph{DR-BIP (Dynamically Reconfigurable - Behaviour Interaction Priority model)}~\cite{,Ballouli18dr-bip}
+Two formal models for specifying reconfigurable architectures are \emph{DR-BIP (Dynamically Reconfigurable - Behaviour Interaction Priority model)}~\cite{Ballouli18dr-bip}
 and \emph{DReAM (Dynamically Reconfigurable Architectural Modelling)}~\cite{denicola2020dream-dynamic-reconfig-arch-modelling}.

### 2023-05-15 paper-2023-acsos-pulverization.tex
RW: + Ctrl-F
change title + minors

@@
 
-\title{Declarative runtime reconfiguration of pulverised distributed systems
+\title{Declarative deployment reconfiguration of pulverised collective adaptive systems at runtime
 \thanks{Identify applicable funding agency here. If none, delete this.}
@@
 \begin{abstract}
-In recent years, we witnessed a fast-paced evolution of the infrastructure hosting distributed computations.
+In recent years,
+the infrastructure supporting the execution of situated distributed computations
+evolved with a fast pace.
 %
-Modern applications are designed to be executed on very diverse devices
-and to be deployed on infrastructures composed of devices ranging from cloud servers to mobile and IoT devices,
+Modern collective adaptive applications are designed to be executed on very diverse devices
+and to be deployed on infrastructures composed of devices ranging from cloud servers to wearable devices,
 constituting together a cloud-edge continuum.
@@
 In this paper,
-we introduce a framework for the development of distributed systems
+we introduce a practical framework for the development of collective adaptive systems
 based on the concept of \emph{pulverisation},
@@
 and \emph{ConfSolve}~\cite{DBLP:conf/saso/HewsonAG13}.
+%
+\emph{Ctrl-F}~\cite{DBLP:journals/jss/AlvaresRS17}
+ is an architectural description language
+ with constructs for specifying adaptive behaviour 
+ and policies (constraints) for the reconfiguration of system components.
+%
+These approaches, however, typically suffer from scalability issues.
+
+%%high-level constructs to describe adaptation in software components by means of behavioural programs, i.e., in terms of order and/or conditions under which reconfigurations take place; and policies, i.e., constraints that have to be enforced all along the execution.
+In other words, Ctrl-Fis a first-class language that is applicable to any Component-based application requiring self-adaptation with formal guarantees.
 
@@
 %ADSL (AWaRE DSL)~\cite{DBLP:conf/aswec/ChhetriLUVKNR18/adsl} % and see corresp RW: the most related DSLs are those in the area of automated deployment [17].
-Such DSLs include Deladas [31], [32], ConfSolve [15], [33], j-ASD’s DSL [34] and MuScADeL [35], [36].
+Such DSLs include Deladas [31], [32], ConfSolve [15], [33], j-ASD’s DSL [34] and MuScADeL [35], [36].DBLP:journals/jss/AlvaresRS17
 %% [17]: J.-P.

### 2023-05-15 paper-2023-acsos-pulverization.tex
add some details in sections ii and iii

@@
 \acrodef{iot}[IoT]{Internet of Things}
+\acrodef{cas}[CAS]{collective adaptive system}
+\acrodef{qos}[QoS]{quality of service}
 
@@
 %
+This kind of infrastructure is ideal for the execution of \ac{cas},
+which have components that,
+depending con the conditions at hand,
+could benefit from being deployed on different devices.
+%
 Exploiting this intriguing infrastructure, however, poses new challenges to the application designer.
@@
 Commonly, in fact, applications are designed with a specific infrastructure in mind,
-whose assumptions leak into the application logic.
+whose assumptions unavoidably leak into the application logic~\cite{Spolsky2004},
+unless captured and encapsulated away from it:
+the higher the coupling between the application and the infrastructure,
+the harder it is for the former to exploit the latter's full potential and adapt to changes.
 
@@
 we continue the research on pulverised systems by providing two key contributions:
-first, we extend the theoretical framework of pulverisation by adding support for
+\textbf{first}, we extend the theoretical framework of pulverisation by adding support for
 \emph{runtime configuration rules}, allowing pulverised components to move at runtime across different hosts;
-the providing a practical implementation called \ourframework{} which,
+\textbf{second}, we provide a practical implementation called \ourframework{} which,
 to the best of our knowledge,
@@
     \item need for deployment-independent specs
-    \item need for runtime reconfiguration (achieve QoS in face of changing conditions --- green computing?)
+    \item need for runtime reconfiguration (achieve \ac{qos} in face of changing conditions --- green computing?)
     \item declarativity over imperative approaches (note: can be compared with the trend in general softeng/build systems)
 \end{itemize}
-
-+ related work
 
@@
 For instance, let us assume a minimal logical system counting two devices:
-two rain gauges logging the water level for future reference,
+two rain gauges logging the water level for future reference
 and opening a valve when the water level crosses a threshold in both devices.
@@
 
+Although the idea behind pulverisation is simple,
+finding a way to implement it effectively requires tackling several challenges at different levels:
+\begin{itemize}
+    \item \textbf{communication}: spliting logical devices into small deployment units implies communication among them,
+    thus requiring to consider networking at two levels:
+    across pulverised components (intra-device) and among logical devices (inter-device);
+    \item \textbf{portability}: hosts will have very diverse hardware specifications, operang systems, and software stacks,
+    \item \textbf{runtime}: runtime reconf how to handle the communication between pulverised components across different runtimes?
+    \item \textbf{language}: for the idea to be exploitable by designers is quintessential that
+    the pulverised configuration
+    and the mapping of pulverised components to the underlying infrastructure
+    can be expressed nimbly and, possibly, declaratively.
+\end{itemize}
+
 \subsection{Domain-specific languages for pulverisation}
+
+In principle, there are several possibilities to implement a pulverisation configurator and runtime:
+as a library
+(configured and executed by a target programming language),
+as a framework/runtime (configured by means of configuration files and with its own runtime),
+or as a \ac{dsl} (configured with its own custom syntax).
+%
+The approaches have different trade-offs.
+%
+A (well designed) library can be flexible and easier to adopt
+as it does not require to learn new languages;
+but it will be usable only in the language/runtime it has been designed for,
+and will require its own framework/runtime to be executed
+(or appropriate bindings for existing ones).
+%
+A framework/runtime configured via declarative files enforces declarativity at the configuration level,
+but it usually dictates network protocols and features limited flexibility,
+as options that have not been accounted for at design time can hardly be injected through configuration files.
+%
+Finally, a \ac{dsl} can be designed to be as expressive as needed,
+but has high maintenance cost and, potentially, a steep learning curve due to the need of learning a new (custom) language.
+%
+
+TODO: recently, internal DSLs
 
@@
 we simulate a city-scale distributed application deployed for a urban event
-and show how the dynamic reconfiguration may help achieving balance between energy consumption and QoS\footnote{
+and show how the dynamic reconfiguration may help achieving balance between energy consumption and \ac{qos}\footnote{
     \url{https://github.com/nicolasfara/ACSOS-2023-pulverization-evaluation}

### 2023-05-16 paper-2023-acsos-pulverization.tex
write something before evacuation

@@
 as a library
-(configured and executed by a target programming language),
-as a framework/runtime (configured by means of configuration files and with its own runtime),
-or as a \ac{dsl} (configured with its own custom syntax).
+configured and executed by a target programming language,
+as a framework/runtime configured by means of configuration files and with its own runtime,
+or as a \ac{dsl} configured with its own custom syntax.
 %
-The approaches have different trade-offs.
+These approaches have different trade-offs.
 %
@@
 Finally, a \ac{dsl} can be designed to be as expressive as needed,
-but has high maintenance cost and, potentially, a steep learning curve due to the need of learning a new (custom) language.
+but it has high maintenance cost (as the language maintenance costs stack upon those of the library/API) and, potentially,
+a steep learning curve due to the need of learning a new (custom) language.
 %
+Recent evolution in programming languages,
 
+such as Kotlin, Ruby, and Scala~\cite{Riti2018}
+however,
+a trend has emerged that
 TODO: recently, internal DSLs
+adopted by gradle
 

### 2023-05-16 paper-2023-acsos-pulverization.tex
Capitalise title
Fix alignment of authors in front matter

@@
 
+\makeatletter
+\newcommand{\linebreakand}{%
+  \end{@IEEEauthorhalign}
+  \hfill\mbox{}\par
+  \mbox{}\hfill\begin{@IEEEauthorhalign}
+}
+\makeatother
+
 \begin{document}
 
-\title{Declarative deployment reconfiguration of pulverised collective adaptive systems at runtime
+\title{Declarative Deployment Reconfiguration of Pulverised Collective Adaptive Systems at Runtime
 \thanks{Identify applicable funding agency here. If none, delete this.}
@@
 roby.casadei@unibo.it}
-\and
+\linebreakand
 \IEEEauthorblockN{3\textsuperscript{rd} Nicolas Farabegoli}
@@
 Cesena, Italy \\
-email address or ORCID}
+mirko.viroli@unibo.it}
 }

### 2023-05-16 paper-2023-acsos-pulverization.tex
Abstract: + examples of domains

@@
 %
-Modern collective adaptive applications are designed to be executed on very diverse devices
+Modern collective adaptive applications -- as found in the Internet of Things, swarm robotics, and social computing -- are designed to be executed on very diverse devices
 and to be deployed on infrastructures composed of devices ranging from cloud servers to wearable devices,
@@
 
-\section{A framework for declarative runtime reconfiguration}\label{sec:contribution}
+\section{Background}\label{sec:background}
+
+\subsection{Pulverisation}
 
@@
 
-\subsection{Domain-specific languages for pulverisation}
+\subsection{Domain-specific languages}
 
@@
 as a library
-configured and executed by a target programming language,
-as a framework/runtime configured by means of configuration files and with its own runtime,
+configured by a program written in a general-purpose language,
+as a framework/runtime configured by means of configuration files,
 or as a \ac{dsl} configured with its own custom syntax.
@@
 %
-A (well designed) library can be flexible and easier to adopt
-as it does not require to learn new languages;
-but it will be usable only in the language/runtime it has been designed for,
-and will require its own framework/runtime to be executed
-(or appropriate bindings for existing ones).
+A (well designed) library can be flexible as the host language can be used to express peculiar configurations
+and easier to adopt to users acquainted with the host language;
+but its configuration may quickly become imperative and harder to maintain as complexity grows.
 %
 A framework/runtime configured via declarative files enforces declarativity at the configuration level,
-but it usually dictates network protocols and features limited flexibility,
+but its flexibility is limited,
 as options that have not been accounted for at design time can hardly be injected through configuration files.
@@
 Recent evolution in programming languages,
+however,
+open to an additional strategy:
+the construction of \emph{internal} \acp{dsl},
+hybrids between libraries and stand-alone (external) \acp{dsl}.
+%
+In fact, modern languages such as Kotlin, Groovy, Ruby, and Scala~\cite{Riti2018}
+provide specific syntactic features to enable the construction of APIs whose ergonomy is akin to the one of a dedicated language,
+but that are valid fragments in the host language.
+%
+Although internal \acp{dsl} are de facto libraries in the host programming language
+(there is no clear boundary defining when a library becomes a \ac{dsl}\footnote{\url{https://archive.is/wip/xAeiX}}),
+from a practical perspective
+they allow for great flexibility and ergonomy
+(although not total as the one of a custom \ac{dsl}, as they are subject to the syntactic constraints of the host language)
+while retaining a reduced maintenance cost
+(as the host language ecosystem and tooling can be reused directly)
+and a gentle learning curve
+(as the syntax will be largely familiar to the host language users).
+%
+The internal DSL approach has been successfully applied in several domains,
+spanning from
+build systems\footnote{\url{https://archive.is/5xtaN}}
+and ontologies~\cite{DBLP:journals/jossw/Balhoff16} to
+hardware design~\cite{DBLP:journals/trets/SerreP20} to
+logic~\cite{2pkt-swx16} and aggregate programming~\cite{DBLP:journals/softx/CasadeiVAP22}.
 
-such as Kotlin, Ruby, and Scala~\cite{Riti2018}
-however,
-a trend has emerged that
-TODO: recently, internal DSLs
-adopted by gradle
+\section{A \ac{dsl} for runtime reconfiguration of pulverised systems}\label{sec:contribution}
 
@@
 \begin{enumerate*}[label=\it{(\roman*)}]
-  \item the definition of the available devices and their capabilities (infrastructure); and
-  \item the allocation of the pulverised components to them (runtime).
+  \item the definition of the available devices and their capabilities (\emph{infrastructure}); and
+  \item the allocation of the pulverised components to them (\emph{runtime}).
 \end{enumerate*}
 %
-We decided to capture these phases into two distinct \acp{dsl}.
+Coherently with the single responsibility principle,
+we decided to capture these phases into two distinct \acp{dsl}.
 

### 2023-05-16 paper-2023-acsos-pulverization.tex
restructure the paper, write the background

@@
 %
-Modern collective adaptive applications -- as found in the Internet of Things, swarm robotics, and social computing -- are designed to be executed on very diverse devices
+Modern collective adaptive applications are designed to be executed on very diverse devices
 and to be deployed on infrastructures composed of devices ranging from cloud servers to wearable devices,

### 2023-05-16 paper-2023-acsos-pulverization.tex
Abstract: minors + @danysk could you complete pls? (what does eval show?)

--- 2023-05-16 paper-2023-acsos-pulverization.tex
@@
 but, contextually, introduces new challenges to software designers,
-as applications must be conceived to be able to adapt their deployment to changing conditions.
+as applications must be conceived to be able to adapt their deployment to changing deployment domains and conditions.
 %
@@
 The framework, implemented in Kotlin multiplatform and available as open source,
-is then evaluated in a small-scale real-world demo and in a city-scale simulated scenario.
+is then evaluated in a small-scale real-world demo and in a city-scale simulated scenario,
+ \meta{showing that ???}.
 \end{abstract}

### 2023-05-16 paper-2023-acsos-pulverization.tex
keywords: +deployment
fix snippet 1

@@
 \begin{IEEEkeywords}
-runtime reconfiguration, distributed systems, self-adaptation, self-organisation, pulverisation
+runtime reconfiguration, distributed systems, self-adaptation, self-organisation, pulverisation, deployment.
 \end{IEEEkeywords}
@@
 %
-Coherently with the single responsibility principle,
+Coherently withush the single responsibility principle,
 we decided to capture these phases into two distinct \acp{dsl}.
@@
 %
-The DSL enables the definition of the type of devices that took part in the system,
+
+The DSL enables the definition of the type of devices that take part in the system,
 and for each device can be specified which components constitute it and where those component must run on (e.g. which host must execute the component).

### 2023-05-16 paper-2023-acsos-pulverization.tex
pass on 3.A

@@
 
+\subsection{Infrastructure and requirements}
+
 The infrastructure \ac{dsl} captures the concepts of
-\emph{logical device}, \emph{component}, \emph{capability}, and \emph{requirement}.
+\emph{device type}, \emph{component}, \emph{capability}, and \emph{requirement}.
 %
-
-The DSL enables the definition of the type of devices that take part in the system,
-and for each device can be specified which components constitute it and where those component must run on (e.g. which host must execute the component).
+A reference sample is presented in \Cref{lst:system-configuration}.
 %
-The execution of a component is constrained to hosts with predetermined characteristics:
-the framework models the constraint with the \texttt{Capability} concept,
-namely the property that a host must to ``implement'' to run the component.
+The DSL provides simple means to define capabilities as types in the host language,
+and then to associate them to pulverised components for each device type.
 %
-Different capabilities can be specified for each component to provide a more flexible and scalable deployment of the system.
-
-The~\Cref{lst:system-configuration} provides an example of the usage of the domain-specific language where is defined a system composed of two kind of devices.
+This \ac{dsl} is meant to define constraints that the application imposes on the infrastructure,
+ruling out configurations that are not capable to support the system
+(for instance, deploying the sensor component of a device in a host that exposes no sensor).
 %
-The DSL returns the system's configuration that will be used by the runtime.
-
-One key-point of the domain-specific language can be observed in the definition of the \texttt{device-2} device:
-the \emph{Behaviour} component can be deployed on hosts that implement the \texttt{HighCPU} and \texttt{EmbeddedDevice} capabilities,
-in this way the behaviour can be moved (at runtime) from a host to another one that implements the same capabilities.
+All combinations of components and capabilities that are defined as supported become instead amenable to
+\emph{runtime reconfiguration}.
 %
-The association between a component and a single capability do not necessarily imply that the component can be executed only on a single host,
-but it is the way to specify the minimum requirements for the host that will execute the component.
+Notice, for instance, that \texttt{Behavior} in \texttt{iot-sensor} is defined as executable on any of host exposing
+\texttt{HighCPU} and/or \texttt{EmbeddedDevice} as capabilities:
+it implies that the component can be moved from at runtime across any host implementing any such capability.
 
@@
     caption={
-        Example of the usage of the domain-specific language for the system configuration.
-        The example configures two types of devices, and for each device defines the components that compose it and which capabilities requires for the execution.
+        Example infrastructure configuration,
+        defining three capabilities and a system composed of two device types,
+        each component of which is then bound to the set of capabilities it needs to execute.
     },
@@
 ]{listings/SystemConfig.kt}
+
+\subsection{Runtime and reconfiguration}
 

### 2023-05-16 paper-2023-acsos-pulverization.tex
Intro: adjusting the incipits, minors, and +toc

@@
 
-\newcommand{\ourframework}{\texttt{pulse-refrakt}}
+\newcommand{\ourframework}{\textsc{pulse-refrakt}}
 
@@
 
-Recent research trends in topics related to the Internet of Things promote the idea of a cloud-edge continuum~\cite{DBLP:journals/iot/BittencourtISFM18},
-an infrastructure composed of heterogeneous devices ranging from large and powerful cloud servers to small connected things
-on which software computes in a distributed fashion,
+Recent technological and scientific advances
+ are extending 
+ both the \emph{kinds of applications and systems} that are being addressed  
+ as well as their \emph{supporting infrastructure}.
+%
+Emerging applications
+ are, for instance, those based on \emph{\acp{cas}}~\cite{DBLP:journals/sttt/NicolaJW20},
+ namely collections of devices and agents
+ that interact to solve problems or provide services while adapting as a whole to dynamic environments.
+%
+Examples include 
+  large-scale sensor-actuator networks, 
+  swarms of robots,
+  social computing systems,
+  crowds of wearable-augmented people, etc.
+  supporting activities like monitoring,
+  transportation, coordination, and collective intelligence.
+%
+Regarding infrastructure,
+ it is mounting the idea of the \emph{cloud-edge continuum}~\cite{DBLP:journals/iot/BittencourtISFM18}: 
+a multi-layer heterogeneous network of devices (ranging from large and powerful cloud servers to small connected \emph{things})
+on which software can compute, store, and exchange data in a distributed fashion, while 
 optimising for performance and resource utilisation.
 %
-This kind of infrastructure is ideal for the execution of \ac{cas},
-which have components that,
-depending con the conditions at hand,
-could benefit from being deployed on different devices.
+
+This kind of infrastructure is also valuable for  \acp{cas}~\cite{DBLP:journals/tpds/HongCHGZ19,DBLP:journals/comsur/WangZZMLW20,DBLP:journals/comsur/AfrinJRRWH21,IEEE-IoTJ-pulverization-simulation},
+which generally feature components that,
+depending on the conditions at hand,
+could benefit from being deployed on different devices or from \emph{offloading} some of their  tasks.
 %
-Exploiting this intriguing infrastructure, however, poses new challenges to the application designer.
+Exploiting this %intriguing infrastructure, 
+ however, poses new challenges to \ac{cas} application designers.
 %
@@
 Previous work suggests a technique for decoupling business logic and deployment concerns
-called \emph{pulverization}~\cite{FI2020-pulverization}.
+in \acp{cas} called \emph{pulverisation}~\cite{FI2020-pulverization}.
 %
 In short, the core idea is to consider devices as \emph{logical} entities,
-whose software can be designed as (or broken down to, if already existing) an ensemble of five micro-components
-(behaviour, state, communication, sensors, and actuators)
-whose actual host can be decided at deployment time.
+whose software can be designed as (or broken down to, if already existing) an ensemble of \emph{five components
+(behaviour, state, communication, sensors, and actuators)}
+ that can be deployed with flexibility on available infrastructure without (ideally) affecting application functionality. %whose actual host can be decided at deployment time.
 %
@@
 although the application can be deployed on arbitrary systems,
-there is no support for the reconfiguration of components in face of changing conditions or requirements.
+there is no support for the reconfiguration of components at runtime, as a reaction to changing conditions or requirements.
 
@@
 \emph{runtime configuration rules}, allowing pulverised components to move at runtime across different hosts;
-\textbf{second}, we provide a practical implementation called \ourframework{} which,
+\textbf{second}, we provide a practical implementation in Kotlin called \ourframework{} which,
 to the best of our knowledge,
-is the first \emph{practical framework supporting pulverisation} for a general-purpose language.
+is the first \emph{practical framework supporting pulverisation}. % for a general-purpose language.
 %
-We exercise \ourframework{} with two case studies:
+To evaluate \ourframework{}, we exercise it through two case studies:
 a small-scale real-wold demonstration, and a city-scale simulated scenario.
 
+\meta{
 \begin{itemize}
@@
 \end{itemize}
+}
+
+The manuscript is organised as follows.
+%
+\Cref{sec:background} provides background on \meta{deployment and} pulverisation.
+%
+\Cref{sec:contribution} describes the proposed \ac{dsl} and platform.
+%
+\Cref{sec:evaluation} provides an evaluation of the approach.
+%
+\Cref{sec:rw} covers related work.
+%
+Finally, \Cref{sec:conclusion} concludes the paper and highlights directions for future work.
 
@@
 \section{Related Work}
+\label{sec:rw}
 

### 2023-05-16 paper-2023-acsos-pulverization.tex
Introductory concepts on deployment and reconfiguration -- possono essere utili @danysk ?

@@
 \section{Background}\label{sec:background}
+
+\subsection{Deployment and Reconfiguration: Basic Concepts}
+
+The \emph{deployment view} is a well-known architectural viewpoint for software systems,
+concerned with the mapping of \emph{software components} to \emph{physical machines}~\cite{DBLP:journals/software/Kruchten95}
+ and supported by modelling notations like \emph{UML Deployment diagrams}.
+%
+Here, we provide a brief introduction to deployment and reconfiguration, based on the conceptual characterisation of~\cite{DBLP:journals/jss/ArcangeliBL15,carzaniga1998characterization}.
+%
+A \emph{site} is a set of computers (\emph{hosts}) that may host a \emph{software system}, i.e., a coherent collection of artifacts.
+%
+\emph{(Software) Deployment} is the process of moving and making a software product available and operational from one or more \emph{producer sites} to a target \emph{deployment domain} (i.e., a set of \emph{consumer sites}).
+%
+The \emph{deployment plan} defines 
+ the mapping between the software system
+ and the deployment domain,
+ augmented with further information (e.g., metadata).
+%
+Deployment-related activities 
+ include: (i) \emph{release/update} of the software system at the producer site;
+ (ii) \emph{installation/deinstallation} at/from the consumer sites;
+ (iii) \emph{activation/deactivation}, for starting/stopping the components;
+ (iv) \emph{reorganisation} of the software system;
+ and (v) \emph{redistribution}, i.e., changing the deployment plan.
 

### 2023-05-16 paper-2023-acsos-pulverization.tex
revise listing 2

@@
 
-The second domain-specific language allows to define the concrete implementation of each component associating it to a specific start host;
-moreover, it allows to define the creteria on which the system should be reconfigured by specifying which events should trigger a reconfiguration.
-%
-This DSL binds the more abstrac definition specified in the previous DSL to the topology of the available infrastracture,
-defining the actual deployment of the system, and the reconfiguration policies.
+The second \ac{dsl} supports the definition of the mapping between pulverised components and specific hosts,
+as well as the definition of reactive reconfiguration policies.
 %

### 2023-05-17 paper-2023-acsos-pulverization.tex
Improved flow in intro

@@
 
-\meta{LIMIT 10 pages incl. refs; DEADLINE 19 may}
+%\meta{LIMIT 10 pages incl. refs; DEADLINE 19 may}
 
@@
 %
-Exploiting this %intriguing infrastructure, 
- however, poses new challenges to \ac{cas} application designers.
+However, exploiting this infrastructure,
+ poses new challenges to \ac{cas} application designers.
 %
@@
 
-Previous work suggests a technique for decoupling business logic and deployment concerns
-in \acp{cas} called \emph{pulverisation}~\cite{FI2020-pulverization}.
+The general solution is to devise 
+ a reasonable partitioning of the software system
+ and a corresponding (dynamic) deployment plan 
+ defining the mapping between the software components and the target deployment domain~\cite{DBLP:journals/jss/ArcangeliBL15}.
+%
+In the context of \acp{cas},
+ previous work proposes an application partitioning schema,
+ called \emph{pulverisation}~\cite{FI2020-pulverization},
+ the fosters the decoupling of business logic and deployment concerns.
 %

### 2023-05-17 paper-2023-acsos-pulverization.tex
Intro: minors + some citations on context

@@
  both the \emph{kinds of applications and systems} that are being addressed  
- as well as their \emph{supporting infrastructure}.
+ as well as their \emph{supporting infrastructure}~\cite{DBLP:journals/iot/GillXOPBSGSWASM22}.
 %
 Emerging applications
- are, for instance, those based on \emph{\acp{cas}}~\cite{DBLP:journals/sttt/NicolaJW20},
+ are, for instance, those based on \emph{\acp{cas}}~\cite{DBLP:journals/computer/Abowd16,DBLP:journals/sttt/NicolaJW20},
  namely collections of devices and agents
- that interact to solve problems or provide services while adapting as a whole to dynamic environments.
+ that interact to solve problems or provide services while adapting \emph{as a whole} to dynamic environments.
 %
@@
   social computing systems,
-  crowds of wearable-augmented people, etc.
-  supporting activities like monitoring,
-  transportation, coordination, and collective intelligence.
+  crowds of wearable-augmented people, and so on---supporting activities like monitoring,
+  transportation, coordination, and other forms of collective intelligence~\cite{casadei2023artl-ci}.
 %

### 2023-05-17 paper-2023-acsos-pulverization.tex
Background on deployment and its relationshop with RW sec

@@
 
-\subsection{Deployment and Reconfiguration: Basic Concepts}
+\subsection{Deployment and Reconfiguration: Basic Concepts}\label{sec:background:dep}
 
 The \emph{deployment view} is a well-known architectural viewpoint for software systems,
-concerned with the mapping of \emph{software components} to \emph{physical machines}~\cite{DBLP:journals/software/Kruchten95}
+concerned with the mapping of \emph{software components} to \emph{physical machines}~\cite{DBLP:journals/software/Kruchten95}, 
  and supported by modelling notations like \emph{UML Deployment diagrams}.
@@
 %
-\emph{(Software) Deployment} is the process of moving and making a software product available and operational from one or more \emph{producer sites} to a target \emph{deployment domain} (i.e., a set of \emph{consumer sites}).
+\emph{(Software) Deployment} is the process of moving and making a software system available and operational from one or more \emph{producer sites} to a target set of \emph{consumer sites}, also called the \emph{deployment domain}.
 %
@@
  and the deployment domain,
- augmented with further information (e.g., metadata).
+ possibly augmented with further information (e.g., metadata, constraints, preferences).
 %
-Deployment-related activities 
+Common deployment-related activities 
  include: (i) \emph{release/update} of the software system at the producer site;
@@
  and (v) \emph{redistribution}, i.e., changing the deployment plan.
+%
+According to the analytical framework of~\cite{DBLP:journals/jss/ArcangeliBL15},
+  the problem of (automatic) deployment of distributed software system can be addressed by considering
+ (i) the nature of the software to be deployed (e.g., how the software system is split into components);
+ (ii) the nature of the deployment domain (i.e. the characteristics and topology of the available infrastructure);
+ (iii) how the deployment is designed (e.g., how the deployment plan is specified);
+ and 
+ (iv) how the deployment is performed (e.g., how deployment activities are carried out).
+%
+Related work covering these issues is provided in \Cref{sec:rw}.
 
@@
 
-This work provides 
- \acp{dsl}
- for specifying 
- distributed systems (partitioned according to the 
- pulverisation model)
- and
- corresponding deployment plans,
- and a framework for executing
- deployment plans and reconfigurations.
+%This work provides 
+% \acp{dsl}
+% for specifying 
+% distributed systems (partitioned according to the 
+% pulverisation model)
+% and
+% corresponding deployment plans,
+% and a framework for executing
+% deployment plans and reconfigurations.
 %
-A way to structure all this 
- is by using the analytical framework of~\cite{DBLP:journals/jss/ArcangeliBL15},
-  where the problem of (automatic) deployment of distributed software system is studied by considering
- (i) the nature of software to be deployed,
- (ii) the nature of the deployment domain (i.e. the available infrastructure),
- (iii) how the deployment is designed,
- and 
- (iv) how the deployment is performed.
-%
-Accordingly, 
- we cover related works about
+Following the conceptual framework introduced in \Cref{sec:background:dep},
+ in this section we cover related work about
  languages for 

### 2023-05-17 paper-2023-acsos-pulverization.tex
Some glue
Mentioning middleware
Picture of pulv + better intro

@@
 \usepackage{listing-style}
+\usepackage{subcaption}
+\usepackage{tikz}
+\usetikzlibrary{arrows,shapes.geometric,positioning,automata,calc}
 \usepackage[hidelinks]{hyperref}
@@
 %
+By an operational point of view, 
+ deployment may be supported by a \emph{middleware}~\cite{DBLP:journals/cacm/GazisK22}.
+%
 Related work covering these issues is provided in \Cref{sec:rw}.
+%
+In the following, we introduce the deployment approach 
+  that we extend and upon which we develop \ourframework{}.
+
+\begin{figure}
+\tikzset{-,
+  host/.style={rectangle,draw,line width={2pt},inner sep=10pt,
+  	outer sep=0, minimum height=1.5cm, minimum width=1.8cm, %text height=0.2cm, 
+  	text depth=0.5cm,
+  	fill=black!10!white
+  },
+  node/.style={rectangle,draw,dotted,line width={1pt}, inner sep=2pt, 
+  	fill=blue!20!white,
+  	font=\large
+  },
+  nodeA/.style={node,fill=red!20!white},
+  nodeB/.style={node,fill=green!20!white},
+  nodeC/.style={node,fill=black!30!white},
+  nodeD/.style={node,fill=white!20!white},
+  plink/.style={line width=2pt},
+  llink/.style={dotted,line width=2pt,red},
+  hostThin/.style={rectangle,draw,line width={0.5pt},inner sep=10pt,
+  	outer sep=0, minimum height=1.1cm, minimum width=1.8cm, %text height=0.2cm, 
+  	text depth=0.5cm,
+  	fill=black!10!white
+  },
+  lnode/.style={node,minimum width=0.55cm,minimum height=0.55cm},
+  loglink/.style={->,line width=1.5pt}
+}
+\def\nm{0.35cm} %nm = node margin offset
+\def\tpscale{0.7}
+\newcommand{\agent}{node}
+\newcommand{\LSens}{\boldsymbol{\sigma}}
+\newcommand{\LComp}{\boldsymbol{\beta}}
+\newcommand{\LComm}{\boldsymbol{\chi}}
+\newcommand{\LAct}{\boldsymbol{\alpha}}
+\newcommand{\LState}{\boldsymbol{\kappa}}
+
+\begin{minipage}{\columnwidth}
+\centering
+\begin{tikzpicture}[every node/.style={scale=1}]
+\node[hostThin,minimum width=3.4cm,minimum height=3cm,dashed]
+ (h1) [label={[yshift=0.35cm]above:{\textbf{logical \agent{}}}}] {};
+
+\node[lnode] (d1) at (h1.north west) [xshift=\nm,yshift=-\nm,label=above:{behaviour}] {$\LComp$};
+\node[lnode] (d2) at (h1.north east) [xshift=-\nm,yshift=-\nm,label=above:{communication}] {$\LComm$};
+\node[lnode] (d3) at (h1.center) [xshift=0,yshift=0,label=right:{state/knowledge}] {$\LState$};
+\node[lnode] (d4) at (h1.south west) [xshift=\nm,yshift=\nm,label=below:{sensors}] {$\LSens$};
+\node[lnode] (d5) at (h1.south east) [xshift=-\nm,yshift=\nm,label=below:{actuators}] {$\LAct$};
+
+\node[hostThin,minimum width=2.3cm,minimum height=2.3cm,dashed]
+ (h2) [right=2cm of h1, label={above:{\textbf{neighbour \agent{}}}}] {};
+
+\node[lnode] (d21) at (h2.north west) [xshift=\nm,yshift=-\nm] {$\LComm$};
+\node[lnode] (d22) at (h2.north east) [xshift=-\nm,yshift=-\nm] {$\LComp$};
+\node[lnode] (d23) at (h2.center) [xshift=0,yshift=0] {$\LState$};
+\node[lnode] (d24) at (h2.south west) [xshift=\nm,yshift=\nm] {$\LSens$};
+\node[lnode] (d25) at (h2.south east) [xshift=-\nm,yshift=\nm] {$\LAct$};
+
+\draw[loglink] (d1) -- (d3);
+\draw[loglink] (d2) -- (d3);
+\draw[loglink] (d3) -- (d1);
+\draw[loglink] (d3) -- (d2);
+\draw[loglink] (d4) -- (d3);
+\draw[loglink] (d3) -- (d5);
+
+\draw[loglink] (d21) -- (d23);
+\draw[loglink] (d22) -- (d23);
+\draw[loglink] (d23) -- (d21);
+\draw[loglink] (d23) -- (d22);
+\draw[loglink] (d24) -- (d23);
+\draw[loglink] (d23) -- (d25);
+
+\draw[loglink] (d2.east) -- (d21.west);
+\draw[loglink] (d21.west) -- (d2.east);
+
+\end{tikzpicture}
+\subcaption{A logical device, split into sub-components, and one of its neighbours.\label{fig:pulv:dev}}
+\end{minipage}
+\\[0.2cm]
+\begin{minipage}{\columnwidth}
+\def\nm{0.3cm} %nm = node margin offset
+
+\begin{minipage}{0.48\columnwidth}\centering
+\begin{tikzpicture}[node distance=1.0cm and 0.5cm,every node/.style={scale=\tpscale}]
+% physical nodes
+\node[host] (h1) [anchor=north,label=above:{}] {};
+\node[host] (h2) [right=of h1,label=above:{}] {};
+\node[host] (h3) [below=of h1, label=above:{}] {};
+\node[host] (h4) [right=of h3, label=above:{}] {};
+
+\node[node] (n11) at (h1.south west) [xshift=\nm,yshift=\nm] {$\LSens$};
+\node[node] (n12) at (h1.south east) [xshift=-\nm,yshift=\nm] {$\LComm$};
+\node[node] (n13) at (h1.north east) [xshift=-\nm,yshift=-\nm] {$\LAct$};
+\node[node] (n14) at (h1.north west) [xshift=\nm,yshift=-\nm] {$\LComp$};
+\node[node] (n15) at (h1) [] {$\LState$};
+% logical node 2
+\node[nodeA] (n21) at (h2.south west) [xshift=\nm,yshift=\nm] {$\LComm$};
+\node[nodeA] (n22) at (h2.south east) [xshift=-\nm,yshift=\nm] {$\LAct$};
+\node[nodeA] (n23) at (h2.north east) [xshift=-\nm,yshift=-\nm] {$\LSens$};
+\node[nodeA] (n24) at (h2.north west) [xshift=\nm,yshift=-\nm] {$\LComp$};
+\node[nodeA] (n25) at (h2) [] {$\LState$};
+% logical node 3
+\node[nodeB] (n31) at (h3.south west) [xshift=\nm,yshift=\nm] {$\LSens$};
+\node[nodeB] (n32) at (h3.south east) [xshift=-\nm,yshift=\nm] {$\LAct$};
+\node[nodeB] (n33) at (h3.north east) [xshift=-\nm,yshift=-\nm] {$\LComm$};
+\node[nodeB] (n34) at (h3.north west) [xshift=\nm,yshift=-\nm] {$\LComp$};
+\node[nodeB] (n35) at (h3) [] {$\LState$};
+% logical node 4
+\node[nodeC] (n41) at (h4.south west) [xshift=\nm,yshift=\nm] {$\LSens$};
+\node[nodeC] (n42) at (h4.south east) [xshift=-\nm,yshift=\nm] {$\LAct$};
+\node[nodeC] (n43) at (h4.north east) [xshift=-\nm,yshift=-\nm] {$\LComp$};
+\node[nodeC] (n44) at (h4.north west) [xshift=\nm,yshift=-\nm] {$\LComm$};
+\node[nodeC] (n45) at (h4) [] {$\LState$};
+
+% physical links
+\draw[plink] 
+	(h1) -- (h2)
+	(h2) -- (h3)
+	(h2) -- (h4)
+	(h3) -- (h4)
+;
+% logical links
+\draw[llink]
+    (n12) to [] (n21)
+    (n21) to [] (n44)
+    (n21) to [bend right=30] (n33)
+    (n33) to [] (n44);
+\end{tikzpicture}
+\subcaption{Peer-to-peer architecture: one-to-one mapping between logical and physical devices, with no offloading.\label{fig:pulv:p2p}}
+\end{minipage}
+%\hfill
+%\begin{minipage}{0.48\columnwidth}\centering
+%\begin{tikzpicture}[node distance=1.0cm and 0.5cm,
+%every node/.style={scale=\tpscale}]
+%
+%% physical nodes
+%\node[host,minimum width=1.5cm] (h1) [anchor=north,label=above:{}] {};
+%\node[host,minimum width=1.5cm] (h2) [right=of h1,label=above:{}] {};
+%\node[host,minimum width=1.5cm] (h3) [right=of h2, label=above:{}] {};
+%\node[host,minimum width=2.5cm] (he) [above right=0.6cm and -1cm of h1,label=above:{}] {};
+%\node[host,minimum width=3.5cm] (h5) [above=2.6cm of h2,label=right:{}] {};
+%
+%\node[node] (n11) at (h1.south west) [xshift=\nm,yshift=\nm] {$\LSens$};
+%\node[node] (n12) at (h1.south east) [xshift=-\nm,yshift=\nm] {$\LAct$};
+%\node[node] (n13) at (h1.north east) [xshift=-\nm,yshift=-\nm] {$\LState$};
+%\node[node] (n14) at (h1.north west) [xshift=\nm,yshift=-\nm] {$\LComp$};
+%\node[node] (n15) at (he) [xshift=-2*\nm] {$\LComm$};
+%% logical node 2
+%\node[nodeA] (n21) at (h2.south west) [xshift=\nm,yshift=\nm] {$\LSens$};
+%\node[nodeA] (n22) at (h2.south east) [xshift=-\nm,yshift=\nm] {$\LAct$};
+%\node[nodeA] (n23) at (h2.north east) [xshift=-\nm,yshift=-\nm] {$\LState$};
+%\node[nodeA] (n24) at (h2.north west) [xshift=\nm,yshift=-\nm] {$\LComp$};
+%\node[nodeA] (n25) at (he) [xshift=0] {$\LComm$};
+%% logical node 3
+%\node[nodeB] (n31) at (h3.south west) [xshift=\nm,yshift=\nm] {$\LSens$};
+%\node[nodeB] (n32) at (h3.south east) [xshift=-\nm,yshift=\nm] {$\LAct$};
+%\node[nodeB] (n33) at (h3.north east) [xshift=-\nm,yshift=-\nm] {$\LState$};
+%\node[nodeB] (n34) at (h3.north west) [xshift=\nm,yshift=-\nm] {$\LComp$};
+%\node[nodeB] (n35) at (h5) [xshift=2*\nm] {$\LComm$};
+%
+%% physical links
+%\draw[plink] 
+%	(h1) -- (he)
+%	(h2) -- (he)
+%	(h3) -- (h5)
+%	(he) -- (h5);
+%
+%\draw[llink]
+%    %(n15) to [bend left=20] (n35)
+%    (n25) to [bend left=30] (n35);
+%\end{tikzpicture}
+%\subcaption{Communication component offloaded.\label{fig:pulv:comm}}
+%\end{minipage}
+%\\[0.2cm]
+%\begin{minipage}{0.48\columnwidth}\centering
+%\begin{tikzpicture}[node distance=1.0cm and 0.5cm,
+%every node/.style={scale=\tpscale}]
+%
+%% physical nodes
+%\node[host,minimum width=1.5cm] (h1) [anchor=north,label=above:{}] {};
+%\node[host,minimum width=1.5cm] (h2) [right=of h1,label=above:{}] {};
+%\node[host,minimum width=1.5cm] (h3) [right=of h2, label=above:{}] {};
+%\node[host,minimum width=2.5cm] (he) [above right=0.6cm and -1cm of h1,label=above:{}] {};
+%\node[host,minimum width=3.5cm] (h5) [above=2.6cm of h2,label=right:{}] {};
+%
+%\node[node] (n11) at (h1.south west) [xshift=\nm,yshift=\nm] {$\LSens$};
+%\node[node] (n12) at (h1.south east) [xshift=-\nm,yshift=\nm] {$\LAct$};
+%\node[node] (n13) at (he.north) [xshift=-2*\nm,yshift=-\nm] {$\LComm$};
+%\node[node] (n14) at (h1.north) [yshift=-\nm] {$\LComp$};
+%\node[node] (n15) at (he.south) [xshift=-2*\nm,yshift=\nm] {$\LState$};
+%% logical node 2
+%\node[nodeA] (n21) at (h2.south west) [xshift=\nm,yshift=\nm] {$\LSens$};
+%\node[nodeA] (n22) at (h2.south east) [xshift=-\nm,yshift=\nm] {$\LAct$};
+%\node[nodeA] (n23) at (he.north) [xshift=0*\nm,yshift=-\nm] {$\LComm$};
+%\node[nodeA] (n24) at (h2.north) [yshift=-\nm] {$\LComp$};
+%\node[nodeA] (n25) at (he.south) [xshift=0,yshift=\nm] {$\LState$};
+%% logical node 3
+%\node[nodeB] (n31) at (h3.south west) [xshift=\nm,yshift=\nm] {$\LSens$};
+%\node[nodeB] (n32) at (h3.south east) [xshift=-\nm,yshift=\nm] {$\LAct$};
+%\node[nodeB] (n33) at (h5.south) [xshift=2*\nm,yshift=\nm] {$\LComm$};
+%\node[nodeB] (n34) at (h3.north) [yshift=-\nm] {$\LComp$};
+%\node[nodeB] (n35) at (h5.south) [xshift=4*\nm,yshift=\nm] {$\LState$};
+%
+%% physical links
+%\draw[plink] 
+%	(h1) -- (he)
+%	(h2) -- (he)
+%	(h3) -- (h5)
+%	(he) -- (h5);
+%
+%\draw[llink]
+%    %(n15) to [bend left=20] (n35)
+%    (n23) to [bend left=20] (n33);
+%\end{tikzpicture}
+%\subcaption{Communication and state components offloaded.\label{fig:pulv:state}}
+%\end{minipage}
+\hfill
+\begin{minipage}{0.48\columnwidth}
+\begin{tikzpicture}[node distance=1.0cm and 0.5cm,
+every node/.style={scale=\tpscale}]
+
+% physical nodes
+\node[hostThin,minimum width=1.5cm] (h1) [anchor=north,label=above:{}] {};
+\node[hostThin,minimum width=1.5cm] (h2) [right=of h1,label=above:{}] {};
+\node[hostThin,minimum width=1.5cm] (h3) [right=of h2, label=above:{}] {};
+\node[host,minimum width=2.5cm] (he) [above right=0.6cm and -1cm of h1,label=above:{}] {};
+\node[host,minimum width=3.5cm] (h5) [above=2.6cm of h2,label=right:{}] {};
+
+\node[node] (n11) at (h1.south west) [xshift=\nm,yshift=\nm] {$\LSens$};
+\node[node] (n12) at (h1.south east) [xshift=-\nm,yshift=\nm] {$\LAct$};
+\node[node] (n13) at (he.south east) [xshift=-4*\nm,yshift=\nm] {$\LComm$};
+\node[node] (n14) at (he.south west) [xshift=1*\nm,yshift=\nm] {$\LComp$};
+\node[node] (n15) at (he.north) [xshift=-2*\nm,yshift=-\nm] {$\LState$};
+% logical node 2
+\node[nodeA] (n21) at (h2.south west) [xshift=\nm,yshift=\nm] {$\LSens$};
+\node[nodeA] (n22) at (h2.south east) [xshift=-\nm,yshift=\nm] {$\LAct$};
+\node[nodeA] (n23) at (he.south east) [xshift=-2.5*\nm,yshift=\nm] {$\LComm$};
+\node[nodeA] (n24) at (he.south west) [xshift=2.5*\nm,yshift=\nm] {$\LComp$};
+\node[nodeA] (n25) at (he.north) [xshift=0,yshift=-\nm] {$\LState$};
+% logical node 3
+\node[nodeB] (n31) at (h3.south west) [xshift=\nm,yshift=\nm] {$\LSens$};
+\node[nodeB] (n32) at (h3.south east) [xshift=-\nm,yshift=\nm] {$\LAct$};
+\node[nodeB] (n33) at (h5.south east) [xshift=-5*\nm,yshift=\nm] {$\LComm$};
+\node[nodeB] (n34) at (h5.south east) [xshift=-3*\nm,yshift=\nm] {$\LComp$};
+\node[nodeB] (n35) at (h5.south east) [xshift=-1*\nm,yshift=\nm] {$\LState$};
+
+% physical links
+\draw[plink] 
+	(h1) -- (he)
+	(h2) -- (he)
+	(h3) -- (h5)
+	(he) -- (h5);
+
+\draw[llink]
+    (n23) to [bend left=25] (n33);
+\end{tikzpicture}
+\subcaption{IoT hosts can be thin, with some components offloaded at the edge and cloud.\label{fig:pulv:full}}
+\end{minipage}
+\end{minipage}
+
+\caption{Pulverisation model and examples of deployments (Figure adapted from~\cite{DBLP:journals/fi/CasadeiPPVW20}).
+Notation: solid-border boxes denote physical hosts (bold borders are for thick devices); solid lines denote connections between hosts; dashed-border boxes denote software components; %, with letters indicating their function (\textbf{B}ehaviour, \textbf{C}ommunication, \textbf{K}nowledge, \textbf{S}ensors, \textbf{A}ctuators);
+differently colours denote (software components of) different logical devices;
+red dashed lines denote connections between the software components, i.e., neighbouring relationships (not shown for co-located software components).
+}
+\label{fig:pulv}
+\end{figure}
 
@@
 
-The goal of pulverisation~\cite{FI2020-pulverization}
+\emph{Pulverisation}~\cite{FI2020-pulverization,IEEE-IoTJ-pulverization-simulation}
+ is an approach to distributed application partitioning
+ and deployment, exemplified in \Cref{fig:pulv} and described in the following.
+%
+Its goal 
  is to provide application designers a way to specify the functional semantics of their software in a deployment-independent way.

### 2023-05-17 paper-2023-acsos-pulverization.tex
Fix ref
Intro: Minor clarificaiton

@@
 %
-One limitation of the current pulverisation approach regards dynamicity:
+One limitation of previous work on pulverisation regards dynamicity:
 although the application can be deployed on arbitrary systems,
@@
 
-\caption{Pulverisation model and examples of deployments (Figure adapted from~\cite{DBLP:journals/fi/CasadeiPPVW20}).
+\caption{Pulverisation model and examples of deployments (Figure adapted from~\cite{FI2020-pulverization}).
 Notation: solid-border boxes denote physical hosts (bold borders are for thick devices); solid lines denote connections between hosts; dashed-border boxes denote software components; %, with letters indicating their function (\textbf{B}ehaviour, \textbf{C}ommunication, \textbf{K}nowledge, \textbf{S}ensors, \textbf{A}ctuators);

### 2023-05-17 paper-2023-acsos-pulverization.tex
Terminology -> device in pulv fig
Comment meta out in intro
Minor

@@
 
-\meta{
-\begin{itemize}
-    \item need for deployment-independent specs
-    \item need for runtime reconfiguration (achieve \ac{qos} in face of changing conditions --- green computing?)
-    \item declarativity over imperative approaches (note: can be compared with the trend in general softeng/build systems)
-\end{itemize}
-}
+%\meta{
+%\begin{itemize}
+%    \item need for deployment-independent specs
+%    \item need for runtime reconfiguration (achieve \ac{qos} in face of changing conditions --- green computing?)
+%    \item declarativity over imperative approaches (note: can be compared with the trend in general softeng/build systems)
+%\end{itemize}
+%}
 
@@
 %
-\Cref{sec:background} provides background on \meta{deployment and} pulverisation.
+\Cref{sec:background} provides background on deployment and pulverisation.
 %
@@
 \def\tpscale{0.7}
-\newcommand{\agent}{node}
+\newcommand{\agent}{device}
 \newcommand{\LSens}{\boldsymbol{\sigma}}

### 2023-05-17 paper-2023-acsos-pulverization.tex
Minors

@@
 Examples include 
-  large-scale sensor-actuator networks, 
+  \ac{iot} deployments, 
   swarms of robots,
@@
 or the application is developed using a framework supporting automatic decomposition
-(one notable example are aggregate computing frameworks~\cite{BealIEEEComputer2015}).
+(one notable example are \emph{aggregate computing} frameworks~\cite{BealIEEEComputer2015,JLAMP2019}).
 %

### 2023-05-17 paper-2023-acsos-pulverization.tex
COMMON-WEARS + PON ack
Drop meta, already satisfied
adjustments to III and II.C

@@
 \title{Declarative Deployment Reconfiguration of Pulverised Collective Adaptive Systems at Runtime
-\thanks{Identify applicable funding agency here. If none, delete this.}
+\thanks{
+    This work has been supported by the Italian PRIN Project COMMON-WEARS (2020HCWWLP) and the EU/MUR FSE REACT-EU PON R\&I 2014-2020.
+}
 }
@@
 
-\meta{maybe import the image from the pulverisation paper?}
-
 Although the idea behind pulverisation is simple,
@@
 (as the host language ecosystem and tooling can be reused directly)
-and a gentle learning curve
+and a gentle learning curve compared to stand-alone \acp{dsl}
 (as the syntax will be largely familiar to the host language users).
 %
-The internal DSL approach has been successfully applied in several domains,
+The killer feature of internal \acp{dsl}, compared to external ones,
+is the counterpass of the limited syntax imposed by the host:
+if some peculiar entity or process cannot be express by the \ac{dsl},
+the designer can fall back to the host language directly,
+with no need to resort to interoperability features,
+and with the possibility to exploit all the libraries designed for the host language.
+%
+The internal DSL approach has thus been successfully applied in several domains,
 spanning from
@@
 logic~\cite{2pkt-swx16} and aggregate programming~\cite{DBLP:journals/softx/CasadeiVAP22}.
+%
 
@@
 %
-The~\Cref{lst:runtime-configuration} provides an example of the usage of the domain-specific language where is defined the runtime configuration of the system.
-%
-The framework define the concept of \texttt{Host} that is the representation of an available physical host where the components can be deployed;
-for each host is possible to specify the capabilities that it implements.
-%
-In this way, the framework can check if a host can execute a component by verifying if the host implements the required capabilities.
-%
-The runtime is configured be specifying for each implementation of a \emph{component},
-its associated startup host enabling the definition of the initial deployment of the system.
-%
-The most relevant part of the domain-specific language, is the definition of the reconfiguration rules:
-the framework models the concept of \texttt{ReconfigurationEvent}
-that is the event that triggers the reconfiguration of the system.
-%
-The DSL allows the association of a \texttt{ReconfigurationEvent}
-to a new deployment of the system which is defined as the new association between component and host.
+It works with the concepts of \texttt{Host}
+(associated to \texttt{Capability}),
+\texttt{ReconfigurationEvent},
+and \texttt{ReconfigurationRule}.
+%
+Using \Cref{lst:runtime-configuration} as a reference example,
+we show how reconfiguration events can be captured at the type system level
+by expressing them as a predicate over the asynchronous flow of \texttt{events} the system's state.
+%
+Similarly, hosts can be configured by defining their name and capabilities.
+%
+Once the hosts composing the system have been configured,
+the entire infrastructure can be set up by defining,
+for each logical device,
+which host should host each pulverised component.
+%
+Additionally,
+reconfiguration rules can be defined to specify how the system should react to reconfiguration events.
+%
+In this context, the \ac{dsl} exposes a special syntax to express the migration of a pulverised component on a different host.
+%
+Crucially, being the DSL internal,
+specialised behaviour not supported by the \ac{dsl} can be defined in the host language directly:
+configuration blocks surrounded by curly brackets are,
+in fact,
+plain lambda expressions, supporting any kind of computation.
 
@@
 Also mention that is multiplatform
+
+\subsection{Limitations}
 
@@
 
+openness
+
 global policies

### 2023-05-17 paper-2023-acsos-pulverization.tex
Fix line number style
pass on III.C

@@
 \newcommand{\meta}[1]{{\color{blue}#1}}
+
+\lstset{xleftmargin=0.3cm,numbersep=5pt}
 
@@
 we show how reconfiguration events can be captured at the type system level
-by expressing them as a predicate over the asynchronous flow of \texttt{events} the system's state.
+by expressing them as a predicate over an asynchronous flow of \texttt{events} influencing the system's state.
 %
@@
 
-A \texttt{ReconfigurationEvent} is characterized by the \emph{events}
-that it produces and the \emph{condition} that must be satisfied to trigger the reconfiguration.
-%
-The \emph{events} are represented as an asyncronous stream of valued produced by the device and the
-\emph{condition} is a predicate that is evaluated on the stream of events.
-%
-At the first match of the condition, the reconfiguration is triggered and the system is reconfigured with the new deployment.
-%
-The~\Cref{lst:reconfiguration-events} shows an example of the definition of a reconfiguration event.
-
-\lstinputlisting[
-    float=ht,
-    language=Kotlin,
-    caption={Example of the definition of a reconfiguration event.},
-    label=lst:reconfiguration-events,
-]{listings/ReconfigurationEvents.kt}
+% A \texttt{ReconfigurationEvent} is characterized by the \emph{events}
+% that it produces and the \emph{condition} that must be satisfied to trigger the reconfiguration.
+% %
+% The \emph{events} are represented as an asyncronous stream of valued produced by the device and the
+% \emph{condition} is a predicate that is evaluated on the stream of events.
+% %
+% At the first match of the condition, the reconfiguration is triggered and the system is reconfigured with the new deployment.
+% %
+% The~\Cref{lst:reconfiguration-events} shows an example of the definition of a reconfiguration event.
+
+% \lstinputlisting[
+%     float=ht,
+%     language=Kotlin,
+%     caption={Example of the definition of a reconfiguration event.},
+%     label=lst:reconfiguration-events,
+% ]{listings/ReconfigurationEvents.kt}
 

### 2023-05-17 paper-2023-acsos-pulverization.tex
Add simulation charts]

@@
 \subsubsection{Results}
+
+\begin{figure}[ht]
+    \begin{minipage}{\columnwidth}
+        \includegraphics[width=\linewidth]{figures/cloud_cost-device_consumption-cloud_consumption-distance-device=100.0.pdf}
+    \end{minipage}
+    \caption{
+        Results of the simulation with 100 devices.
+        %
+        The \emph{green} bar represents the mean power consumed by cloud;
+        the \emph{blue} bar represents the mean power consumed by all the devices in the system;
+        the \emph{yellow} bar represents the mean traveled distance by the people
+        and the \emph{purple} bar represents the average cost per hour of the cloud.
+        %
+        The chart shows two ``static'' scenarios: \emph{device-only} (on the left) and \emph{cloud-only} (on the right).
+        %
+        All the intermediate scenarios are obtained by exploiting the reconfiguration capabilities of the pulverization framework,
+        where different battery threshold are used to trigger the reconfiguration.
+    }
+    \label{fig:result-100-devices}
+\end{figure}
+
+\begin{figure}[ht]
+    \begin{minipage}{\columnwidth}
+        \includegraphics[width=\linewidth]{figures/cloud_cost-device_consumption-cloud_consumption-distance-device=300.0.pdf}
+    \end{minipage}
+    \caption{
+        Results of the simulation with 300 devices.
+        %
+        The \emph{green} bar represents the mean power consumed by cloud;
+        the \emph{blue} bar represents the mean power consumed by all the devices in the system;
+        the \emph{yellow} bar represents the mean traveled distance by the people
+        and the \emph{purple} bar represents the average cost per hour of the cloud.
+        %
+        The chart shows two ``static'' scenarios: \emph{device-only} (on the left) and \emph{cloud-only} (on the right).
+        %
+        All the intermediate scenarios are obtained by exploiting the reconfiguration capabilities of the pulverization framework,
+        where different battery threshold are used to trigger the reconfiguration.
+    }
+    \label{fig:result-300-devices}
+\end{figure}
+
+\begin{figure}[ht]
+    \begin{minipage}{\columnwidth}
+        \includegraphics[width=\linewidth]{figures/cloud_cost-device_consumption-cloud_consumption-distance-device=1000.0.pdf}
+    \end{minipage}
+    \caption{
+        Results of the simulation with 1000 devices.
+        %
+        The \emph{green} bar represents the mean power consumed by cloud;
+        the \emph{blue} bar represents the mean power consumed by all the devices in the system;
+        the \emph{yellow} bar represents the mean traveled distance by the people
+        and the \emph{purple} bar represents the average cost per hour of the cloud.
+        %
+        The chart shows two ``static'' scenarios: \emph{device-only} (on the left) and \emph{cloud-only} (on the right).
+        %
+        All the intermediate scenarios are obtained by exploiting the reconfiguration capabilities of the pulverization framework,
+        where different battery threshold are used to trigger the reconfiguration.
+    }
+    \label{fig:result-1000-devices}
+\end{figure}
 

### 2023-05-17 paper-2023-acsos-pulverization.tex
write something on the language choice

@@
 
-Also mention that is multiplatform
+We implemented the proposed \ac{acronym} in Kotlin.
+%
+We wanted a language with a strongly typed syntax to help developers with better contextual assist,
+and we wanted the possibility to target multiple runtimes
+(Kotlin currently targets the JVM, JavaScript, and native code for many platforms\footnote{\url{https://archive.is/JSiTg}},
+including mobile and wearable devices)
+and be as interoperable as possible with other languages.
+%
+Scala could have been a good alternative,
+but we deemed the multiplatform support of Kotlin a better fit for our needs.
+%
+The framework has been open sourced and released with a permissive license\footnote{\url{https://github.com/nicolasfara/pulverization-framework}}.
 

### 2023-05-17 paper-2023-acsos-pulverization.tex
fix dangling sentences

@@
     across pulverised components (intra-device) and among logical devices (inter-device);
-    \item \textbf{portability}: hosts will have very diverse hardware specifications, operang systems, and software stacks,
-    \item \textbf{runtime}: runtime reconf how to handle the communication between pulverised components across different runtimes?
+    \item \textbf{portability}: hosts will have very diverse hardware specifications, operating systems, and software stacks;
+    \item \textbf{runtime}: the system should be able to reconfigure its deployment at runtime without disrupting the application logic; and
     \item \textbf{language}: for the idea to be exploitable by designers is quintessential that
@@
 
-\section{A \ac{dsl} for runtime reconfiguration of pulverised systems}\label{sec:contribution}
+\section{A DSL for reconfigurable pulverised systems}\label{sec:contribution}
 
@@
 
-\subsection{Implementation details}
-
-We implemented the proposed \ac{acronym} in Kotlin.
+\subsection{Implementation details and underlying runtime}
+
+We implemented the proposed \ac{dsl} in Kotlin.
 %

### 2023-05-17 paper-2023-acsos-pulverization.tex
write a section on limitations

@@
 
-\subsection{Limitations}
+The configuration produced by the \ac{dsl} needs a platform that instructs the system on the way actual communication needs to be performed.
+%
+In our current implementation,
+we provide out of the box a JVM platform based on the MQTT protocol
+and supporting both intra- and inter-device communication.
+%
+For the large scale experiment,
+we implemented a second platform,
+this time to comply with the requirements of the simulation engine we used.
+%
+The implementation required little effort,
+and it has been made available as part of the evaluation repositories.
+
+\subsection{Limitations of the current version of the DSL}
+
+In its current form, the proposed \ac{dsl} has two major limitations.
+%
+First and foremost, currently it does not consider openness:
+hosts get specified in the runtime and reconfiguration \ac{dsl} as a closed set,
+so the specification is not robust to the addition of new hosts or the loss of existing ones.
+%
+This is a major limitation that,
+although planned to be addressed shortly with minor additions to the \ac{dsl},
+must be highlighted in this manuscript.
+%
+The second limitation is related to the policies that can be expressed in the reconfiguration \ac{dsl}:
+currently, only policies that can be assessed at the local host level can be assessed.
+%
+A policy that requires to consider the state of multiple hosts would require to resort to the host language primitive mechanisms and make explicit communications outside the framework,
+thus breaking the abstraction.
+%
+This limitation can be addressed by extending the current reconfiguration
+block of the \ac{dsl} in such a way that queries can be performed on any reachable host of the system,
+thus making all communications
+(including reconfiguration-related ones)
+pass through the channels controlled by the framework.
 
@@
 
+more platforms
+
 global policies

### 2023-05-17 paper-2023-acsos-pulverization.tex
begin a pass on section 4A

@@
 In this section,
-we exercise the proposed framework with two case studies.
+we exercise the proposed framework through two case studies.
 %
@@
 
-In this section we present a simple scenario where the pulverization framework is used to implement a crowd alert system:
-we simulate a laboratory where a safety system warns people when they are too close to each other.
-%
-Each person is equipped with a wearable device that is able to detect the distance from other devices and send the distances data to the
-alert system.
-%
-The people in the laboratory must observe a safety distance, and the system warns them when the distance is not respected by showing a red light
-when they are too close to each other, and a green color when the distance is respected (see~\Cref{fig:crowd-alert}).
-%
-\begin{figure*}[ht]
+In this example, we exercise the framework on a real-world testbed
+located in a laboratory.
+%
+In particular,
+we want to avoid situations in which too many people are too close to each other in the same room for safety reasons.
+%
+To do so,
+we equip each person with a wearable device that is able to detect the distance from other simialr devices via Bluetooth.
+%
+For the sake of simplicity,
+in the experiment we relied on smartphones to emulate wearables,
+and we used \meta{
+    NICOLAS please write here (within meta, so that it remains blue)
+    what you used to measure the signal strenght and estimate the distance.
+    Even formulas, calibration procedures, and stuff
+}
+%
+In our case the warning was issued through a monitor,
+whose color was set to tones ranging from green to red depending on the average distance among people:
+the shorter the distance, the more red the tone.
+%
+\Cref{fig:crowd-alert} shows a representation of the scenario.
+%
+\begin{figure}
     \centering
-    \includegraphics[width=.8\textwidth]{figures/crow-laboratory-demo.drawio.pdf}
+    \includegraphics[width=\columnwidth]{figures/crow-laboratory-demo.drawio.pdf}
     \caption{
-        Representation of the scenario described in the~\Cref{sec:small-scale-crowd-alert}
-        where more the people are far from each other, more the color of the screen is green.
-        More the people are close to each other, more the color of the screen is red.
-        The blue dot represents the wearable device and the link between the devices represents the distance between them.
-        The rectangle represents the monitor in the laboratory which shows the congestion level of the laboratory (via the color described above).
+        Representation of the scenario described in \Cref{sec:small-scale-crowd-alert}.
+        graph dots represent wearables,
+        and arcs lenghts represent the distance between them.
+        %
+        The rectangle at the bottom represents the monitor.
+        %
+        The closer the wearables are, the higher is the alert level
+        (reddish tones).
+        %
+        \meta{NICOLAS riduci lo spazio vuoto fra i tre casi, così da guadagnare spazio. La legenda potrebbe essere messa sotto o sopra per guadagnarne di ulteriore}
     }
     \label{fig:crowd-alert}
-\end{figure*}
+\end{figure}
 %

### 2023-05-17 paper-2023-acsos-pulverization.tex
describe the device types in 4a, leave comments for @nicolasfara

@@
 \end{figure}
-%
-The system is composed of two kind of devices: \emph{wearable} and \emph{laboratory}.
-%
-The former represents the wearable devices that is worn by the people, and consists of the following pulverized components:
-$\beta$, $\chi$, and $\sigma$.
-%
-The \emph{behaviour} ($\beta$) component is responsible for converting the raw data from the sensor into the distance
-between the device and the others; the \emph{sensors} ($\sigma$) component is responsible for perceiving the other devices and
-the \emph{communication} ($\chi$) component is responsible for sending the data to the alert system.
-%
-The latter represents the alert system and consists of the following pulverized components:
-$\beta$, $\kappa$, $\chi$, and $\alpha$.
-%
-The \emph{behaviour} ($\beta$) component is responsible for evaluating the meand distance between the devices,
-the \emph{actuators} ($\alpha$) is responsible for triggering the alarm showing a color in the screen,
-and the \emph{communication} ($\chi$) component is responsible for receiving the data from the wearable devices.
+
+The system is composed of two kinds of device: \emph{wearable} and \emph{alert}.
+%
+Wearables consist of the following pulverised components:
+sensors ($\sigma^w$), perceiving other devices by their signal strenght;
+behaviour ($\beta^w$), converting raw data from the sensor into messages for the alert system; and
+communication ($\chi^w$) sending data to the alert system.
+%
+They do not need state nor actuators.
+%
+The alert consists of the following pulverized components:
+$\beta^a$, computing the mean distance among devices;
+$\kappa^a$ \meta{
+    NICOLAS what does this do? write within this meta block
+};
+$\chi^a$, responsible for receiving the data from the wearable devices;
+$\alpha^a$, enacting the alarm by showing a color on the screen.
+%
+It needs no sensors.
 

### 2023-05-17 paper-2023-acsos-pulverization.tex
@danysk on pulverisation, distinction between app-level and infrastructural physical devices is important imho ... since in our CAS apps we generally want to 'control' a collective of physical systems

@@
 In this process,
-a single logical component could be (and usually does) end up being executed on multiple hosts.
+a single logical device could (and usually does) end up being executed on multiple hosts.
+%
+Indeed, though logical devices are generally associated to ``application-level'' physical devices that needs to be controlled or monitored (e.g., a drone, a sensor, or a person---a logical ensemble of wearables),
+ their execution can also be supported by other physical devices (e.g., purely infrastructural ones).
 

### 2023-05-17 paper-2023-acsos-pulverization.tex
Improve image, add RSSI to meter conversion formula

@@
 \usepackage[inline]{enumitem}
+\usepackage{siunitx}
+\DeclareSIUnit{\decibelmeter}{dBm}
 
@@
 To do so,
-we equip each person with a wearable device that is able to detect the distance from other simialr devices via Bluetooth.
+we equip each person with a wearable device that is able to detect the distance from other similar devices via Bluetooth.
 %
@@
 in the experiment we relied on smartphones to emulate wearables,
-and we used \meta{
-    NICOLAS please write here (within meta, so that it remains blue)
-    what you used to measure the signal strenght and estimate the distance.
-    Even formulas, calibration procedures, and stuff
+and we used
+% \meta{
+%     NICOLAS please write here (within meta, so that it remains blue)
+%     what you used to measure the signal strenght and estimate the distance.
+%     Even formulas, calibration procedures, and stuff
+% }
+%
+\meta{
+    the RSSI signal of the bluetooth during the scanning of the environment
+    to estimate the distance between the devices.
+    %
+    In particular, we used the following formula to estimate the distance:
+    \begin{equation*}
+        d = 10^{\cfrac{(A - A')}{(10 \cdot n)}}
+    \end{equation*}
+    where $A$ is the RSSI value at 1 meter,
+    $A'$ is the RSSI value measured by the device
+    and $n$ is the path loss exponent.
+    %
+    Since the definition of $A'$ and $n$ is not trivial,
+    and depends on the environment in which the system is deployed,
+    we have calibrated the system in a test environment by measuring the RSSI signal
+    at 1 meter using two smartphones.
+    %
+    The following values emerged from the calibration phase:
+    $A=\qty{-60}{\decibelmeter}$ and $n = \qty{2.0}{\decibel}$.
 }

### 2023-05-17 paper-2023-acsos-pulverization.tex
Add clarification for @danysk

@@
 Wearables consist of the following pulverised components:
-sensors ($\sigma^w$), perceiving other devices by their signal strenght;
+sensors ($\sigma^w$), perceiving other devices by their signal strength;
 behaviour ($\beta^w$), converting raw data from the sensor into messages for the alert system; and
@@
 $\kappa^a$ \meta{
-    NICOLAS what does this do? write within this meta block
+    NICOLAS what does this do? write within this meta block.
+    %
+    Hold in memory the updated distances between the devices.
 };
@@
 %
-Given the infrastracture defined above, we simulate also a condition where the \emph{Server} is overloaded and, via pulverization,
-we want to lighteninig the load of the \emph{Server} by moving the $\beta_{werable}$ in the \emph{Smartphones}.
+Given the infrastructure defined above, we simulate also a condition where the \emph{Server} is overloaded and, via pulverization,
+we want to lightening the load of the \emph{Server} by moving the $\beta_{werable}$ in the \emph{Smartphones}.
 %
@@
 
-At the start of the system, each \emph{Werable} device, with the $\sigma$ component,
+At the start of the system, each \emph{Wearable} device, with the $\sigma$ component,
 start to perceive the other devices in the laboratory using the RSSI signal from the Bluetooth.
@@
 Once the distances are computed, the $\chi$ component sends the data to the
-\emph{Laboratory} containing the dinstances perceived by the device.
+\emph{Laboratory} containing the distances perceived by the device.
 %
@@
 %
-The $\kappa$ component persist in memory the updated devices grpah based on the distances retreived by the $\chi$ component.
+The $\kappa$ component persist in memory the updated devices graph based on the distances retrieved by the $\chi$ component.
 and the $\alpha$ component show in the monitor the color based on the mean distance computed by the $\beta$ component.
 
-As soon as the reconfiguration event is triggered, the $\beta$ component of the \emph{Werable} is moved from the \emph{Server} to the \emph{Smartphone}
-lighteninig the server and preserving from one side the functional semantic of the system and from the other side maintain the system operational.
+As soon as the reconfiguration event is triggered, the $\beta$ component of the \emph{Wearable} is moved from the \emph{Server} to the \emph{Smartphone}
+lightening the server and preserving from one side the functional semantic of the system and from the other side maintain the system operational.
 %
@@
 %
-Moreover, we show how the pulverization framework can be used adapt the system to changing requirement without loosing the funcional semantic of the system
+Moreover, we show how the pulverization framework can be used adapt the system to changing requirement without loosing the functional semantic of the system
 by describing the system in a declarative way.

### 2023-05-17 paper-2023-acsos-pulverization.tex
write a bit more of IV.A

@@
 
-The infrastructure defined for this scenario is described as follow:
-
+We use the following hosts in the scenario:
 \begin{itemize}
-    \item \emph{Smartphones} are the devices used by the people to detect the distance from the others;
-    \item \emph{Local PC} is the device used to show the alarm in the laboratory;
-    \item \emph{Server} is the device used run the crowd algorithm and the wearable logic. 
+    \item multiple \emph{Smartphones} used by the people to detect the distance from the others;
+    \item a \emph{Monitor} used to show the alarm in the laboratory;
+    \item a \emph{Server} used run the crowd algorithm and the wearable logic.
 \end{itemize}
 
-The initial deployment of the system starts with the $\beta_{werable}$ executed in the \emph{Server},
-the $\sigma_{wearable}$ and $\chi_{werable}$ executed in the \emph{Smartphones},
-and all the \emph{Laboratory} components are executed in the \emph{Local PC}.
-%
-Given the infrastructure defined above, we simulate also a condition where the \emph{Server} is overloaded and, via pulverization,
-we want to lightening the load of the \emph{Server} by moving the $\beta_{werable}$ in the \emph{Smartphones}.
+Initially, $\beta^{w}$ is allocated to the \emph{Server} while
+$\sigma^{w}$ and $\chi^{w}$ are executed on the \emph{Smartphones},
+and all the \emph{alert} components are executed in the \emph{Monitor}.
+%
+In the experiment,
+we define a rule by which,
+if the \emph{Server} is overloaded,
+$\beta^{w}$ is moved to the \emph{Smartphones}
+(with each smartphone running the $\beta^{w}$ corresponding to its $\sigma^{w}$).
+%
+During the experiment,
+we launch several heavy-duty processes on the \emph{Server},
+and we observe how the system reacts to the overload.
+
+Given the infrastracture defined above, we simulate also a condition where the \emph{Server} is overloaded and, via pulverization,
+we want to lighteninig the load of the \emph{Server} by moving the $\beta_{werable}$ in the \emph{Smartphones}.
 %

### 2023-05-18 paper-2023-acsos-pulverization.tex
further pass on the snippets
reduce image size

@@
 \usepackage{cleveref}
+\usepackage{color}
 
@@
     \centering
-    \includegraphics[width=\columnwidth]{figures/crow-laboratory-demo.drawio.pdf}
+    \includegraphics[width=.9\columnwidth]{figures/crow-laboratory-demo.drawio.pdf}
     \caption{
@@
         (reddish tones).
-        %
-        \meta{NICOLAS riduci lo spazio vuoto fra i tre casi, così da guadagnare spazio. La legenda potrebbe essere messa sotto o sopra per guadagnarne di ulteriore}
     }

### 2023-05-18 paper-2023-acsos-pulverization.tex
add calibration

@@
 \acrodef{qos}[QoS]{quality of service}
+\acrodef{rssi}[RSSI]{received signal strength indicator}
 
@@
 %
-\meta{
-    the RSSI signal of the bluetooth during the scanning of the environment
-    to estimate the distance between the devices.
-    %
-    In particular, we used the following formula to estimate the distance:
-    \begin{equation*}
-        d = 10^{\cfrac{(A - A')}{(10 \cdot n)}}
-    \end{equation*}
-    where $A$ is the RSSI value at 1 meter,
-    $A'$ is the RSSI value measured by the device
-    and $n$ is the path loss exponent.
-    %
-    Since the definition of $A'$ and $n$ is not trivial,
-    and depends on the environment in which the system is deployed,
-    we have calibrated the system in a test environment by measuring the RSSI signal
-    at 1 meter using two smartphones.
-    %
-    The following values emerged from the calibration phase:
-    $A=\qty{-60}{\decibelmeter}$ and $n = \qty{2.0}{\decibel}$.
-}
-%
-In our case the warning was issued through a monitor,
-whose color was set to tones ranging from green to red depending on the average distance among people:
-the shorter the distance, the more red the tone.
+%\meta{
+    the Bluetooth \ac{rssi} to estimate the distance $d$ between devices as:
+    $$d=10^{\frac{R_{ref} - R}{10 \cdot n}}$$
+    where $R_{ref}$ is the \ac{rssi} reference value at 1 meter
+%    (calibrated in a test environment by measuring it with two smartphones),
+    R is the currently measured \ac{rssi},
+    and $n$ is the path loss exponent,
+    a parameter indicating the rate at which the \ac{rssi} decreases with distance.
+    %
+    $R_{ref}$ and $n$ are environment-sensitive parameters that require calibration;
+    we set up a test with two smartphones and measured,
+    for our setup,
+    $R_{ref}\qty{-60}{\decibelmeter}$ and $n=\qty{2.0}{\decibel}$.
+%    Since the definition of $A'$ and $n$ is not trivial,
+%    and depends on the environment in which the system is deployed,
+%    we have calibrated the system in a test environment by measuring the RSSI signal
+%    at 1 meter using two smartphones.
+%    %
+%    The following values emerged from the calibration phase:
+%    $A=\qty{-60}{\decibelmeter}$ and $n = \qty{2.0}{\decibel}$.
+%}
+%
+We used a monitor to issue the warning
+by setting its colour from green to red depending on the average distance among people:
+the shorter the distance, the redder the tone.
 %

### 2023-05-18 paper-2023-acsos-pulverization.tex
explain the role of the state

@@
 $\beta^a$, computing the mean distance among devices;
-$\kappa^a$ \meta{
-    NICOLAS what does this do? write within this meta block.
-    %
-    Hold in memory the updated distances between the devices.
-};
+$\kappa^a$, storing the distances measured by the device;
+%\meta{
+%    NICOLAS what does this do? write within this meta block.
+%    %
+%    Hold in memory the updated distances between the devices.
+%};
 $\chi^a$, responsible for receiving the data from the wearable devices;

### 2023-05-18 paper-2023-acsos-pulverization.tex
pass on IVA complete

@@
 we provide out of the box a JVM platform based on the MQTT protocol
-and supporting both intra- and inter-device communication.
+supporting both intra- and inter-device communication.
+%
+The platform also feature an optimization for pulverized components of the same logical device ending up running on the same host:
+as it would be expected,
+they would communicate directly by sharing memory,
+without the need to go through (de)serialisation processes and networking.
+\meta{NICOLAS please check if the last sentence is true, if so remove this comment}
 %
@@
 $\beta^a$, computing the mean distance among devices;
-$\kappa^a$, storing the distances measured by the device;
+$\kappa^a$, storing the distances measured by the device for comparison with the recent past;
 %\meta{
@@
 \begin{itemize}
-    \item multiple \emph{Smartphones} used by the people to detect the distance from the others;
-    \item a \emph{Monitor} used to show the alarm in the laboratory;
-    \item a \emph{Server} used run the crowd algorithm and the wearable logic.
+    \item multiple \emph{Smartphones}, one per user, emulating wearables and used as distance sensors;
+    \item a \emph{Monitor} used to show the alarm;
+    \item a \emph{Server} used run crowd estimation and wearable logic.
 \end{itemize}
@@
 and we observe how the system reacts to the overload.
-
-Given the infrastracture defined above, we simulate also a condition where the \emph{Server} is overloaded and, via pulverization,
-we want to lighteninig the load of the \emph{Server} by moving the $\beta_{werable}$ in the \emph{Smartphones}.
-%
-Formally, we can write the reconfiguration event as follow: $\beta^{werable}_{server} \xrightarrow{\mathcal{E}} \beta^{werable}_{smartphone}$
-where $\mathcal{E}$ represents the reconfiguration event due to the high load in the \emph{Server}.
-
-At the start of the system, each \emph{Wearable} device, with the $\sigma$ component,
-start to perceive the other devices in the laboratory using the RSSI signal from the Bluetooth.
-%
-The $\beta$ component, initially executed in the \emph{Server}, starts to compute the distances from the other devices
-based on the received RSSI signal from the $\sigma$ component.
-%
-Once the distances are computed, the $\chi$ component sends the data to the
-\emph{Laboratory} containing the distances perceived by the device.
-%
-
-Similarity, the \emph{Laboratory} device has all the components executed in the \emph{Local PC},
-where the $\chi$ component receives the data from the \emph{Wearable},
-the $\beta$ component computes the mean distance between the devices.
-%
-The $\kappa$ component persist in memory the updated devices graph based on the distances retrieved by the $\chi$ component.
-and the $\alpha$ component show in the monitor the color based on the mean distance computed by the $\beta$ component.
-
-As soon as the reconfiguration event is triggered, the $\beta$ component of the \emph{Wearable} is moved from the \emph{Server} to the \emph{Smartphone}
-lightening the server and preserving from one side the functional semantic of the system and from the other side maintain the system operational.
-%
-The reconfiguration of the system in entirely managed by the framework, which handle all the machinery and communication to move the component from
+%
+We rely on the MQTT platform module for all the communication,
+meaning that both intra- and inter-device communication is handled by the same protocol and broker.
+
+As soon as the reconfiguration event is triggered,
+the $\beta$ component of the \emph{Wearable} is moved from the \emph{Server} to the \emph{Smartphone} $s$ hosting $\sigma^{w}_{s}$,
+relieving the server and preserving while retaining the system in nominal conditions.
+%
+The reconfiguration of the system in entirely managed by the framework,
+which handles all the machinery and communication to move the component from
 one host to another.
-%
-With this simple example we show how the pulverization framework can be used in practice to manage a system with different devices and an heterogeneous
-infrastructure.
-%
-Moreover, we show how the pulverization framework can be used adapt the system to changing requirement without loosing the functional semantic of the system
-by describing the system in a declarative way.
+
+This example has been used as a real-world testbed driving our experimental implementation.
+%
+For the sake of brevity,
+we do not report the entire configuration in this manuscript, but it is available in the companion artifact~\cite{TBD}
+\meta{
+    NICOLAS please replace TBD with the artifact citation from zenodo. Use doi2bib to get the bibtex entry.
+}
+for those willing to exercise the whole system.
+%
+To simplify testing we also provide,
+in the same artifact,
+a simulation of the system where every device is containerized.
 

### 2023-05-18 paper-2023-acsos-pulverization.tex
Add simulation screenshot
Replace MQTT with AMQP protocol

@@
 In our current implementation,
-we provide out of the box a JVM platform based on the MQTT protocol
+we provide out of the box a JVM platform based on the AMQP protocol leveraging RabbitMQ,
 supporting both intra- and inter-device communication.
@@
 without the need to go through (de)serialisation processes and networking.
-\meta{NICOLAS please check if the last sentence is true, if so remove this comment}
 %
@@
 For the sake of brevity,
-we do not report the entire configuration in this manuscript, but it is available in the companion artifact~\cite{TBD}
-\meta{
-    NICOLAS please replace TBD with the artifact citation from zenodo. Use doi2bib to get the bibtex entry.
-}
+we do not report the entire configuration in this manuscript, but it is available in the companion artifact~\cite{https://doi.org/10.5281/zenodo.7933160}
+% \meta{
+%     NICOLAS please replace TBD with the artifact citation from zenodo. Use doi2bib to get the bibtex entry.
+% }
 for those willing to exercise the whole system.
@@
     \label{fig:result-1000-devices}
+\end{figure}
+
+\begin{figure}[ht]
+    \begin{minipage}{\columnwidth}
+        \includegraphics[width=\linewidth]{figures/simulation-screenshot.png}
+    \end{minipage}
+    \caption{TODO}
+    \label{fig:simulation-screenshot}   
 \end{figure}

### 2023-05-18 paper-2023-acsos-pulverization.tex
Add cite to zenodo artifacts

@@
 %
-Both experiments have been released with a permissive open-source license and have been archived for future reference on Zenodo~\cite{TBD}.
+Both experiments have been released with a permissive open-source license and have been archived for future reference on
+Zenodo~\cite{https://doi.org/10.5281/zenodo.7933160, https://doi.org/10.5281/zenodo.7948801}.
 

### 2023-05-18 paper-2023-acsos-pulverization.tex
Add power consumption model description

@@
 \cite{PianiniJOS2013}
+
+\meta{
+The simulation of this large-scale scenario should take into account several factors to be more realistic as possible:
+the power consumption of the werable devices (smartphones), the power consumption of the cloud server,
+and the weight of the application in terms of power dissipated by its execution.
+%
+The parameter described above represent the fundamental element of the power consumption model used in the simulation.
+%
+To be more adeherent as possible to a real consumption model of a device or cloud server,
+several research were conducted to figure out real value that can be used as parameter in the simulation.
+%
+Nevertheless,
+for our bes of knowledge,
+there are no precise and established studies regarding these specific parameters.
+%
+For this reason,
+we tried to identify likelihood parameters that could describe a plausible scenario in simulation.
+%
+In the following, we describe the parameters used in the simulation and the motivation behind their choice.
+%
+One popular metrics which relate the power consumption with the instruction is the \emph{Energy Per Instruction}
+(EPI)~\cite{Grochowski2006EnergyPI} measured in $J / instruction$,
+which associate the energy related to the execution of a CPU instruction.
+%
+From the metric introduced above, we should estimate the $EPI_{cloud}$ and $EPI_{device}$.
+%
+To do so, we rely our estimation based on the result obtained in~\cite{DBLP:conf/islped/ShaoB13},
+were the EPI of a server CPU is calculated.
+%
+From the value taken from~\cite{DBLP:conf/islped/ShaoB13} we defined a discrete variability range for
+$EPI_{cloud}$ equals to $300$, $30$, and $3$.
+%
+To estimate the $EPI_{device}$,
+an intermediate step could be determine the ratio between $EPI_{cloud}$ and $EPI_{device}$,
+which we define as $EPI_{ratio}$.
+%
+To do so we could rely on other characterizing elements:
+the \emph{TDP} of each CPU and the \emph{benchmark score} of each CPU,
+and for each one,
+define the ratio of this two metrics: $TDP_{ratio}$ and $Benachmark_{ratio}$.
+%
+In this way we are identifying the power ration and performance ratio between the CPU of the device
+and the CPU of the cloud.
+%
+The $EPI_{ratio}$ now can be define as:
+$$EPI_{ratio} = \frac{TDP_{ratio}}{Benchmark_{ratio}}$$
+%
+Supposing to have smartphones with the \emph{Snapdragon 888} (5W TDP) as a CPU and a cloud server
+with an \emph{Intel Xeon W-3375} (220W TDP), the respective benchmark results are $59,069$ and $9,362$.
+%
+Consequently, the $EPI_{ratio}$ is near to $7$.
+%
+To mitigate the uncertainty of this values we ranged the $EPI_{ratio}$ from 3 to 30.
+%
+The simulation consist of several scenarios in which we simulate different application load,
+and consequently different discharge time of the devices.
+%
+To do so, we estimate the number of instruction based on
+the $EPI_{device}$ and the required discharge time to simulate.
+%
+Given the total instruction of our application,
+we repartition that value for each component executed and the relative communication cost;
+in this way,
+we can estimate the consumption of each component either executed in cloud or in device.
+%
+The power consumption is trivially determined by:
+$$P = \frac{instructions * EPI}{time}$$
+where \emph{time} is the time interval between two measure,
+that in the simulation corresponds to the delta time between two simulation steps.
+}
 

### 2023-05-19 paper-2023-acsos-pulverization.tex
introduce the large scale case

@@
 \subsection{Large scale: urban event}
+
+To show the potential benefit of automatic reconfiguration in \acp{cas},
+we consider a large-scale scenario where a crowd of people is attending an event in a city
+to participate in a collective activity whose computational weight is relevant, for instance
+a tournament of a geo-located game similar to Ingress\footnote{\url{https://www.ingress.com/}}
+or Pokémon Go\footnote{\url{https://pokemongolive.com/}}).
+%
+We suppose the application to be pulverised,
+and its behaviour to be able to run either on the smartphones or in cloud.
+%
+We assume the event to start with users located in random positions in the city,
+and to have a recharging station available at that place.
+%
+The game requires to phisically move in the city,
+reach a \ac{poi} and perform some operations once there.
 

### 2023-05-19 paper-2023-acsos-pulverization.tex
add missing acronym
add cloud cost description

@@
 \acrodef{rssi}[RSSI]{received signal strength indicator}
+\acrodef{poi}[PoI]{point of interest}
 
@@
 Supposing to have smartphones with the \emph{Snapdragon 888} (5W TDP) as a CPU and a cloud server
-with an \emph{Intel Xeon W-3375} (220W TDP), the respective benchmark results are $59,069$ and $9,362$.
+with an \emph{Intel Xeon W-3375} (220W TDP), the respective benchmark results are $21,031$ and $9,362$.
 %
@@
 that in the simulation corresponds to the delta time between two simulation steps.
+%
+For what concern the cloud cost, the simulation is based on the \emph{AWS}
+pricing model\footnote{\url{https://aws.amazon.com/ec2/pricing/on-demand/}}
+using the \emph{m5.16xlarge} instance with a cost of \emph{\$3.584} per hour.
+%
+The number of instance required for running the system is calculated by
+dividing the total power consumption of the cloud by the $TDP_{cloud}$.
+%
+The total cost of the cloud is calculated by multiplying the number of instance by the cost per hour.
 }

### 2023-05-19 paper-2023-acsos-pulverization.tex
Fix some typos

@@
 Supposing to have smartphones with the \emph{Snapdragon 888} (5W TDP) as a CPU and a cloud server
-with an \emph{Intel Xeon W-3375} (220W TDP), the respective benchmark results are $21,031$ and $9,362$.
-%
-Consequently, the $EPI_{ratio}$ is near to $7$.
+with an \emph{Intel Xeon Platinum P-8124} (220W TDP), the respective benchmark results are $22,674$ and $9,362$.
+%
+Consequently, the $EPI_{ratio}$ is near to $18$.
 %

### 2023-05-19 paper-2023-acsos-pulverization.tex
Fix sentence

@@
 pricing model\footnote{\url{https://aws.amazon.com/ec2/pricing/on-demand/}}
-using the \emph{m5.16xlarge} instance with a cost of \emph{\$3.584} per hour.
+using the \emph{m5.16xlarge} instance (which use the CPU server described above)
+with a cost of \emph{\$3.584} per hour.
 %

### 2023-05-19 paper-2023-acsos-pulverization.tex
Background sec on middleware and DSLs: retitling + intro @danysk -- but I do not understand well what follows

@@
 \begin{itemize}
-    \item \textbf{communication}: spliting logical devices into small deployment units implies communication among them,
+    \item \textbf{communication}: splitting logical devices into small deployment units implies communication among them,
     thus requiring to consider networking at two levels:
@@
     \item \textbf{runtime}: the system should be able to reconfigure its deployment at runtime without disrupting the application logic; and
-    \item \textbf{language}: for the idea to be exploitable by designers is quintessential that
+    \item \textbf{language}: for the idea to be exploitable by designers it is important that
     the pulverised configuration
     and the mapping of pulverised components to the underlying infrastructure
-    can be expressed nimbly and, possibly, declaratively.
+    can be expressed expressively and, possibly, declaratively.
 \end{itemize}
 
-\subsection{Domain-specific languages}
+\subsection{Middlewares and Languages for Deployed Systems Specification and Execution}
+
+As mentioned in \Cref{sec:background:dep},
+ the deployment and re-configuration of a software system
+ may be supported -- following the well-known layered architectural style -- by a \emph{middleware} or \emph{runtime}~\cite{DBLP:journals/cacm/GazisK22}
+ encapsulating, among others, deployment and reconfiguration services.
 

### 2023-05-19 paper-2023-acsos-pulverization.tex
Need concept of 'deployment unit'

@@
 %
-A \emph{site} is a set of computers (\emph{hosts}) that may host a \emph{software system}, i.e., a coherent collection of artifacts.
+A \emph{site} is a set of computers (\emph{hosts}) that may host a \emph{software system}, i.e., a product described by a coherent collection of artefacts and consisting of a set of \emph{deployment units} that can be independently operated.
 %

### 2023-05-20 paper-2023-acsos-pulverization.tex
Pass on bib
Towards clarification of II.C on middlewares/runtimes and dsls -- what do you think @danysk

@@
 
+\acrodef{api}[API]{application-program interface}
 \acrodef{dsl}[DSL]{domain-specific language}
@@
 By an operational point of view, 
- deployment may be supported by a \emph{middleware}~\cite{DBLP:journals/cacm/GazisK22}.
+ deployment may be supported by a so-called \emph{runtime} or \emph{middleware}~\cite{DBLP:journals/cacm/GazisK22}.
 %
@@
  the deployment and re-configuration of a software system
- may be supported -- following the well-known layered architectural style -- by a \emph{middleware} or \emph{runtime}~\cite{DBLP:journals/cacm/GazisK22}
+ may be supported -- following the well-known layered architectural style -- by a \emph{runtime} or \emph{middleware}~\cite{DBLP:journals/cacm/GazisK22}
  encapsulating, among others, deployment and reconfiguration services.
-
-In principle, there are several possibilities to implement a pulverisation configurator and runtime:
-as a library
-configured by a program written in a general-purpose language,
-as a framework/runtime configured by means of configuration files,
-or as a \ac{dsl} configured with its own custom syntax.
+%
+The runtime could be reified into a deployment unit \emph{per se},
+ or can be implemented as part of the deployment units of the software system to be deployed---e.g., taking the form of an \emph{\ac{api}} or a \emph{framework}.
+%
+In order to take decisions about deployment and reconfiguration, 
+ such a runtime has to be \emph{configured}
+ for the software application at hand,
+ i.e., it has to be given a \emph{deployment plan} capturing deployment mappings, constraints, and policies.
+%
+
+In principle, there are several possibilities to implement a deployment plan:
+%
+(i) declaratively, through a set of \emph{configuration files};
+(ii) through a \emph{\ac{api}/library}, implemented in a general-purpose language;
+or 
+(iii) through a \ac{dsl}~\cite{dsl-book-voelter}. % configured with its own custom syntax.
 %
 These approaches have different trade-offs.
+%
+A framework/runtime configured via declarative files enforces declarativity at the configuration level,
+but its flexibility is limited,
+as options that have not been accounted for at design time can hardly be injected through configuration files.
 %
@@
 %
-A framework/runtime configured via declarative files enforces declarativity at the configuration level,
-but its flexibility is limited,
-as options that have not been accounted for at design time can hardly be injected through configuration files.
-%
-Finally, a \ac{dsl} can be designed to be as expressive as needed,
-but it has high maintenance cost (as the language maintenance costs stack upon those of the library/API) and, potentially,
+Finally, a \ac{dsl}~\cite{dsl-book-voelter} can be designed to be as expressive as needed,
+but it has high maintenance cost (as the language maintenance costs stack upon those of the library/\ac{api}) and, potentially,
 a steep learning curve due to the need of learning a new (custom) language.
@@
 the construction of \emph{internal} \acp{dsl},
-hybrids between libraries and stand-alone (external) \acp{dsl}.
+hybrids between libraries and stand-alone (external) \acp{dsl}~\cite{dsl-book-voelter}.
 %
@@
 
-\bibliographystyle{IEEEtran}
+\bibliographystyle{IEEEtranDOI}
 \bibliography{bibliography}

### 2023-05-20 paper-2023-acsos-pulverization.tex
Splitting into paragraph + minor title change --- @danysk Imho this subsec has an ISSUE: too much talking about DSLs in general and not specififcally to the matter at hand
Sec III - some terminology change - pls check @danysk @nicolasfara

@@
 
-\subsection{Middlewares and Languages for Deployed Systems Specification and Execution}
+\subsection{Middlewares and (Domain-Specific) Languages for Deployed Systems Specification and Execution}
 
@@
 These approaches have different trade-offs.
-%
+
 A framework/runtime configured via declarative files enforces declarativity at the configuration level,
@@
 a steep learning curve due to the need of learning a new (custom) language.
-%
+
 Recent evolution in programming languages,
@@
 
-\section{A DSL for reconfigurable pulverised systems}\label{sec:contribution}
+\section{A Runtime and \acsp{dsl} for Reconfigurable Pulverised Systems}\label{sec:contribution}
 
 From the previous section,
-we can identify two main phases in the pulverisation process:
+we can identify two main pieces of information for the definition of a deployment plan for a pulverised system:
 \begin{enumerate*}[label=\it{(\roman*)}]
-  \item the definition of the available devices and their capabilities (\emph{infrastructure}); and
-  \item the allocation of the pulverised components to them (\emph{runtime}).
+  \item the definition of the \emph{target} deployment domain (\emph{infrastructure}), i.e., the specification of the \emph{requirements} posed on hosts in terms of required (or preferred) capabilities; and
+  \item the definition of the \emph{deployment plan}, i.e., the allocation of the pulverised components to the infrastructural elements, together with constraints and reconfiguration policies.
 \end{enumerate*}
 %
-Coherently withush the single responsibility principle,
-we decided to capture these phases into two distinct \acp{dsl}.
-
-\subsection{Infrastructure and requirements}
+Coherently with the \emph{single responsibility principle},
+we capture these two descriptive needs into two distinct \acp{dsl}.
+
+\subsection{Infrastructure and Requirements}
 
@@
 
-\subsection{Runtime and reconfiguration}
+\subsection{Deployment and Reconfiguration}
 

### 2023-05-20 paper-2023-acsos-pulverization.tex
Figure of the approach - @danysk @nicolasfara pls check

@@
 we capture these two descriptive needs into two distinct \acp{dsl}.
+%
+An overall view of the approach in terms of the provided tools and the main modelling concepts is provided in \Cref{fig:approach}.
+
+\begin{figure}
+\includegraphics[width=\columnwidth]{figures/pulverization-framework-overview.drawio.pdf}
+\caption{The approach provides two \acp{dsl} and a runtime for specifying and executing deployment plans.
+Each deployment unit will be delivered to the target site and properly configured to interact with other deployment units, as required by the application at hand (see  \Cref{sec:evaluation} for examples).}
+\label{fig:approach}
+\end{figure}
 

### 2023-05-20 paper-2023-acsos-pulverization.tex
Update approach name

@@
 
-\newcommand{\ourframework}{\textsc{pulse-refrakt}}
+\acrodef{ourframework}[\textsc{PulvReAKt}]{Pulverisation with Reconfigurable Adaption in Kotlin}
+\newcommand{\ourframework}{\ac{ourframework}}
 
@@
 \emph{runtime configuration rules}, allowing pulverised components to move at runtime across different hosts;
-\textbf{second}, we provide a practical implementation in Kotlin called \ourframework{} which,
+\textbf{second}, we provide a practical implementation in Kotlin called \emph{\ourframework{}} which,
 to the best of our knowledge,

### 2023-05-20 paper-2023-acsos-pulverization.tex
write section 4b except the results

@@
 \usepackage{color}
+\usepackage{nicefrac}
 
@@
 %
-The game requires to phisically move in the city,
+The game requires to physically move in the city,
 reach a \ac{poi} and perform some operations once there.
-
-\cite{PianiniJOS2013}
-
-\meta{
-The simulation of this large-scale scenario should take into account several factors to be more realistic as possible:
-the power consumption of the werable devices (smartphones), the power consumption of the cloud server,
-and the weight of the application in terms of power dissipated by its execution.
-%
-The parameter described above represent the fundamental element of the power consumption model used in the simulation.
-%
-To be more adeherent as possible to a real consumption model of a device or cloud server,
-several research were conducted to figure out real value that can be used as parameter in the simulation.
-%
-Nevertheless,
-for our bes of knowledge,
-there are no precise and established studies regarding these specific parameters.
-%
-For this reason,
-we tried to identify likelihood parameters that could describe a plausible scenario in simulation.
-%
-In the following, we describe the parameters used in the simulation and the motivation behind their choice.
-%
-One popular metrics which relate the power consumption with the instruction is the \emph{Energy Per Instruction}
-(EPI)~\cite{Grochowski2006EnergyPI} measured in $J / instruction$,
-which associate the energy related to the execution of a CPU instruction.
-%
-From the metric introduced above, we should estimate the $EPI_{cloud}$ and $EPI_{device}$.
-%
-To do so, we rely our estimation based on the result obtained in~\cite{DBLP:conf/islped/ShaoB13},
-were the EPI of a server CPU is calculated.
-%
-From the value taken from~\cite{DBLP:conf/islped/ShaoB13} we defined a discrete variability range for
-$EPI_{cloud}$ equals to $300$, $30$, and $3$.
-%
-To estimate the $EPI_{device}$,
-an intermediate step could be determine the ratio between $EPI_{cloud}$ and $EPI_{device}$,
-which we define as $EPI_{ratio}$.
-%
-To do so we could rely on other characterizing elements:
-the \emph{TDP} of each CPU and the \emph{benchmark score} of each CPU,
-and for each one,
-define the ratio of this two metrics: $TDP_{ratio}$ and $Benachmark_{ratio}$.
-%
-In this way we are identifying the power ration and performance ratio between the CPU of the device
-and the CPU of the cloud.
-%
-The $EPI_{ratio}$ now can be define as:
-$$EPI_{ratio} = \frac{TDP_{ratio}}{Benchmark_{ratio}}$$
-%
-Supposing to have smartphones with the \emph{Snapdragon 888} (5W TDP) as a CPU and a cloud server
-with an \emph{Intel Xeon Platinum P-8124} (220W TDP), the respective benchmark results are $22,674$ and $9,362$.
-%
-Consequently, the $EPI_{ratio}$ is near to $18$.
-%
-To mitigate the uncertainty of this values we ranged the $EPI_{ratio}$ from 3 to 30.
-%
-The simulation consist of several scenarios in which we simulate different application load,
-and consequently different discharge time of the devices.
-%
-To do so, we estimate the number of instruction based on
-the $EPI_{device}$ and the required discharge time to simulate.
-%
-Given the total instruction of our application,
-we repartition that value for each component executed and the relative communication cost;
-in this way,
-we can estimate the consumption of each component either executed in cloud or in device.
-%
-The power consumption is trivially determined by:
-$$P = \frac{instructions * EPI}{time}$$
-where \emph{time} is the time interval between two measure,
-that in the simulation corresponds to the delta time between two simulation steps.
-%
-For what concern the cloud cost, the simulation is based on the \emph{AWS}
-pricing model\footnote{\url{https://aws.amazon.com/ec2/pricing/on-demand/}}
-using the \emph{m5.16xlarge} instance (which use the CPU server described above)
-with a cost of \emph{\$3.584} per hour.
-%
-The number of instance required for running the system is calculated by
-dividing the total power consumption of the cloud by the $TDP_{cloud}$.
-%
-The total cost of the cloud is calculated by multiplying the number of instance by the cost per hour.
+%
+We assume users with low battery to turn off the application and move to their start station to recharge the device
+before rejoining the game.
+%
+The application is pulverised and its behaviour can execute either on the device or in the cloud.
+%
+We configure the framework to move the behaviour to the cloud when the battery level
+(measured as a percentage of the maximum charge)
+is below a threshold $\lambda$,
+and to move it back to the device when the battery is charged enough to cross a threshold $\upsilon$
+(stable configuration have $\upsilon{}>\lambda$,
+otherwise the system continuously moves the behaviour between cloud and devices).
+%
+We thus indicate any configuration with $\Updownarrow^{\upsilon}_{\lambda}$,
+with the exception of two cases:
+$\upsilon, \lambda > 100$, forcing the behaviour to be always in the cloud, that we indicate with $\Uparrow$;
+and $\upsilon, \lambda < 0$, forcing the behaviour to be always on the device, that we indicate with $\Downarrow$.
+%
+We are interested in the following metrics:
+%
+\begin{itemize}
+    \item $P_{devices}$ ($kW$), the average device power consumption, a proxy for the battery life;
+    \item $P_{cloud}$ ($kW$), the average power consumption of the cloud, if added to $P_{devices}$
+    provides the total power consumption
+    (hence, assuming a non-carbon-neutral energy mix, it is also a proxy for the carbon footprint);
+    \item Distance ($km$), the total distance walked by the participants, a proxy for the \ac{qos},
+    as users that need to stop and recharge stop walking;
+    \item $\$_{cloud}$ (\$), the price paid to keep the required cloud instances up and running.
+\end{itemize}
+
+\subsubsection{Energy model}
+
+\acrodef{epi}[EPI]{Energy Per Instruction}
+\acrodef{tdp}[TDP]{Thermal Design Power}
+
+We assume that the power consumption of the device is linearly dependent on the CPU usage,
+and we thus decided to estimate the \ac{epi}~\cite{Grochowski2006EnergyPI}
+of common smartphone ($EPI_{device}$) and server ($EPI_{cloud}$) CPUs.
+%
+To do so, we take the \ac{tdp} of a
+Qualcomm Snapdragon 888 ($5W$)
+and an
+Intel Xeon Platinum P-8124 ($220W$)
+and we divide it by the score the CPU obtained in the popular Passmark benchmark\footnote{\url{https://www.cpubenchmark.net/}}
+($9362$ and $22674$, respectively)
+obtaining an indication of the power per benchmark point
+($\nicefrac{5W}{9362pts}$ and \nicefrac{220W}{22674pts} respectively).
+%
+We then used these ratios to estimate the relative \ac{epi} of the two CPUs,
+obtaining a $EPI_{ratio}$ close to 1:18.
+%
+Thus,
+we take the \ac{epi} estimation for the server CPU from~\cite{DBLP:conf/islped/ShaoB13},
+(approximately 10J per instruction, considering 36 logical cores of our reference processor)
+and we assume the \ac{epi} of the smartphone CPU to be 18 times lower.
+%
+Once the \acp{epi} of the two platforms are known,
+we estimate the number of instructions of the pulverised components as follows:
+we assume the application to be a heavy-duty game,
+capable to drain the battery of a smartphone
+(approximately 4500mAh in capacity, and thus, at 3.3V, 14.85Wh, 53460J)
+in 6 hours of continuous usage,
+which divided by the \ac{epi} of the smartphone CPU gives us the number of instructions.
+%
+Additionally,
+we account for a variable quota of instructions that captures other activities of the devices
+(screen, operating system, other applications, etc.).
+
+\subsubsection{Cost model}
+We assume the cloud to be composed by a set of Amazon \emph{AWS} instances,
+specifically, \emph{m5.16xlarge} instances
+(using our reference server CPU)
+the \emph{AWS}
+with a hourly cost of \$3.584\footnote{\url{https://aws.amazon.com/ec2/pricing/on-demand/}}
+%
+We estimate the number of instances required to run the pulverised components by
+dividing the total power consumption of the cloud by the \ac{tdp} of the server,
+assuming the CPU to be the dominant energy consumption component,
+and assuming that each vCPU on AWS maps to a logical core of the underlying hardware.
+
+\subsubsection{Implementation}
+We implemented our simulations by interfacing \ourframework{}
+with the Alchemist Simulator~\cite{PianiniJOS2013}
+through a dedicated platfrom module.
+%
+We ran experiments varying the number of participants (hence, smartphones): 100, 300, and 1000.
+%
+For each experiment, we run a total of 10 repetitions with a different random seed;
+all results are averaged over them.
+%
+The experiment has been open sourced\footnote{\url{https://github.com/nicolasfara/ACSOS-2023-pulverization-evaluation}}
+and the code has been archived on Zenodo~\cite{https://doi.org/10.5281/zenodo.7948801}.
+%
+\begin{figure}
+    \includegraphics[width=\linewidth]{figures/simulation-screenshot.png}
+    \caption{
+        A snapshot of the large-scale simulation.
+        Smartphones are represented as coloured circles,
+        with the color indicating their battery level
+        (green when charged, red when low).
     }
-
-\subsubsection{Free variables}
-
-As follow are reported the free variables used in the evaluation:
-\begin{itemize}
-    \item let $n$ be the number of devices;
-    \item let $t_d$ be the discharge time of the battery of the devices;
-    \item let $B_l$ be the low battery threshold;
-    \item let $B_h$ be the high battery threshold;
-    \item let $EPI_{cloud}$ be the energy per instruction of the cloud;
-    \item let $EPI_{ratio}$ be the ratio of the consumption of the cloud and the device;
-\end{itemize}
-
-\subsubsection{Metrics}
-
-From the variables above are derived the following metrics:
-
-\begin{itemize}
-    \item let $\beta_{device}$ be the number of \emph{behaviour} executed in the device;
-    \item let $\beta_{cloud}$ be the number of \emph{behaviour} executed in the cloud;
-    \item let $C_{\%}$ be the average battery percentage of the devices;
-    \item let $P_{devices}$ be the consumed power of the devices;
-    \item let $P_{cloud}$ be the consumed power of the cloud;
-    \item let $P_{total}$ be the total consumed power of the system;
-    \item let $I_{cloud}$ be the number of server instances of the cloud;
-    \item let $\$_{cloud}$ be the cost of the cloud;
-\end{itemize}
+    \label{fig:simulation-screenshot}
+\end{figure}
+%
+\Cref{fig:simulation-screenshot} shows a snapshot of the simulation
 
@@
     \label{fig:result-1000-devices}
-\end{figure}
-
-\begin{figure}[ht]
-    \begin{minipage}{\columnwidth}
-        \includegraphics[width=\linewidth]{figures/simulation-screenshot.png}
-    \end{minipage}
-    \caption{TODO}
-    \label{fig:simulation-screenshot}   
 \end{figure}

### 2023-05-20 paper-2023-acsos-pulverization.tex
Infrastructure DSL example: improvements and refs to lines of code

@@
 %
-The DSL provides simple means to define capabilities as types in the host language,
-and then to associate them to pulverised components for each device type.
+The DSL provides simple means to define capabilities as types in the host language (Lines~\ref{capab:start}--\ref{capab:end}),
+and then to associate them to pulverised components for each device type (Lines~\ref{devcap:start}--\ref{devcap:end}).
 %
@@
 %
-All combinations of components and capabilities that are defined as supported become instead amenable to
-\emph{runtime reconfiguration}.
-%
-Notice, for instance, that \texttt{Behavior} in \texttt{iot-sensor} is defined as executable on any of host exposing
-\texttt{HighCPU} and/or \texttt{EmbeddedDevice} as capabilities:
-it implies that the component can be moved from at runtime across any host implementing any such capability.
-
-\lstinputlisting[
+All combinations of components and capabilities that are defined as supported become instead amenable as \emph{targets} of deployment or reconfiguration---forming what we may call the \emph{deployment range}.
+%\emph{runtime reconfiguration}.
+%
+In the example, e.g., component \texttt{Behavior} in \texttt{iot-sensor} is defined as executable on any of host exposing
+\texttt{HighCPU} and/or \texttt{EmbeddedDevice} as capabilities (\Cref{devcap:b}):
+it implies that the component can be dynamically moved to any host offering any such capability.
+
+%\lstinputlisting[
+%    float=ht,
+%    language=Kotlin,
+%    caption={
+%        Example infrastructure configuration,
+%        defining three capabilities and a system composed of two device types,
+%        each component of which is then bound to the set of capabilities it needs to execute.
+%    },
+%    label=lst:system-configuration,
+%]{listings/SystemConfig.kt}
+
+\begin{lstlisting}[
     float=ht,
@@
     label=lst:system-configuration,
-]{listings/SystemConfig.kt}
+    escapechar=\%,
+    numbers=left,
+]
+object HighCPU : Capability%\label{capab:start}%
+object LowLatencyComm : Capability
+object EmbeddedDevice : Capability%\label{capab:end}%
+
+pulverizedSystem {
+  device("control-center") {%\label{devcap:start}%
+    Behavior and State deployableOn HighCPU
+    Communication deployableOn LowLatencyComm
+    Sensors deployableOn EmbeddedDevice
+  }
+  device("iot-sensor") {
+    Behavior deployableOn setOf(HighCPU, EmbeddedDevice)%\label{devcap:b}%
+    Communication deployableOn LowLatencyComm
+    Sensors and Actuators deployableOn EmbeddedDevice
+  }%\label{devcap:end}%
+}
+\end{lstlisting}
 

### 2023-05-20 paper-2023-acsos-pulverization.tex
Improve code style

@@
 \usepackage{nicefrac}
+
+\lstdefinelanguage{pulv}[]{Kotlin}{
+%  basicstyle=\small\ttfamily\lst@ifdisplaystyle\footnotesize\fi,
+%	frame=single,
+%	escapechar=\%,
+%  keywordstyle=\color{blue},
+  keywordstyle=[3]\bfseries{},
+  keywords=[3]{pulverizedSystem,device,and,deployableOn},
+%  sensitive=true,
+%  morecomment=[l]{//},
+%  morecomment=[n]{/*}{*/},
+%  commentstyle=\color{green!40!black},
+%  morestring=[b]",
+%  morestring=[b]',
+%  morestring=[b]""",
+%  stringstyle=\ttfamily\color{orange},
+}
+\lstset{language=pulv,
+basewidth=0.555em % decreases spacing between chars
+}
 
@@
     float=ht,
-    language=Kotlin,
+    language=pulv,
     caption={
@@
     escapechar=\%,
-    numbers=left,
+    numbers=left
 ]
@@
 Using \Cref{lst:runtime-configuration} as a reference example,
-we show how reconfiguration events can be captured at the type system level
+we show how reconfiguration events can be captured at the type-system level
 by expressing them as a predicate over an asynchronous flow of \texttt{events} influencing the system's state.

### 2023-05-20 paper-2023-acsos-pulverization.tex
finish up section 4

@@
 \usepackage{nicefrac}
+\usepackage{wasysym}
 
@@
 \makeatletter
-\newcommand{\linebreakand}{%
+\newcommand{\linebreakand}{
   \end{@IEEEauthorhalign}
@@
 with the Alchemist Simulator~\cite{PianiniJOS2013}
-through a dedicated platfrom module.
-%
-We ran experiments varying the number of participants (hence, smartphones): 100, 300, and 1000.
+through a dedicated platform module.
+%
+We ran experiments varying the number of participants (hence, smartphones): 100, 300, and 1000
+and the \ourframework{} reconfiguration rules:
+$\Uparrow$ (behaviour always in the smartphones),
+$\Updownarrow^{60}_{20}$,
+$\Updownarrow^{80}_{20}$,
+$\Updownarrow^{60}_{40}$,
+$\Updownarrow^{80}_{40}$,
+$\Downarrow$ (behaviour always in the cloud).
+%
+All the hybrid configurations
+(all except $\Uparrow$ and $\Downarrow$)
+require runtime reconfiguration to be realised.
 %
@@
 \subsubsection{Results}
-
-\begin{figure}[ht]
-    \begin{minipage}{\columnwidth}
-        \includegraphics[width=\linewidth]{figures/cloud_cost-device_consumption-cloud_consumption-distance-device=100.0.pdf}
-    \end{minipage}
+%
+\begin{figure*}
+    \includegraphics[width=.32\linewidth]{figures/cloud_cost-device_consumption-cloud_consumption-distance-device=100.0.pdf}~
+    \includegraphics[width=.32\linewidth]{figures/cloud_cost-device_consumption-cloud_consumption-distance-device=300.0.pdf}~
+    \includegraphics[width=.32\linewidth]{figures/cloud_cost-device_consumption-cloud_consumption-distance-device=1000.0.pdf}
     \caption{
-        Results of the simulation with 100 devices.
-        %
-        The \emph{green} bar represents the mean power consumed by cloud;
-        the \emph{blue} bar represents the mean power consumed by all the devices in the system;
-        the \emph{yellow} bar represents the mean traveled distance by the people
-        and the \emph{purple} bar represents the average cost per hour of the cloud.
-        %
-        The chart shows two ``static'' scenarios: \emph{device-only} (on the left) and \emph{cloud-only} (on the right).
-        %
-        All the intermediate scenarios are obtained by exploiting the reconfiguration capabilities of the pulverization framework,
-        where different battery threshold are used to trigger the reconfiguration.
+        Simulation results for 100 (left), 300 (center), and 1000 (right) devices;
+        showing all four metrics:
+        $P_{cloud}$ ($kW$) in green,
+        $P_{devices}$ ($kW$) in blue,
+        Distance in yellow, $\$_{cloud}$ in dark purple.
+        %
+        $P_{cloud}$ and $P_{devices}$ are stacked to also show the overall power consumption.
+        %
+        Different reconfiguration rules are indicated on the horizontal axis.
+        %
+        Runtime reconfiguration of the deployment opens to more opportunities to find the trade-off of cost,
+        energy use, and \ac{qos} that best fits the current context.
     }
-    \label{fig:result-100-devices}
-\end{figure}
-
-\begin{figure}[ht]
-    \begin{minipage}{\columnwidth}
-        \includegraphics[width=\linewidth]{figures/cloud_cost-device_consumption-cloud_consumption-distance-device=300.0.pdf}
-    \end{minipage}
-    \caption{
-        Results of the simulation with 300 devices.
-        %
-        The \emph{green} bar represents the mean power consumed by cloud;
-        the \emph{blue} bar represents the mean power consumed by all the devices in the system;
-        the \emph{yellow} bar represents the mean traveled distance by the people
-        and the \emph{purple} bar represents the average cost per hour of the cloud.
-        %
-        The chart shows two ``static'' scenarios: \emph{device-only} (on the left) and \emph{cloud-only} (on the right).
-        %
-        All the intermediate scenarios are obtained by exploiting the reconfiguration capabilities of the pulverization framework,
-        where different battery threshold are used to trigger the reconfiguration.
-    }
-    \label{fig:result-300-devices}
-\end{figure}
-
-\begin{figure}[ht]
-    \begin{minipage}{\columnwidth}
-        \includegraphics[width=\linewidth]{figures/cloud_cost-device_consumption-cloud_consumption-distance-device=1000.0.pdf}
-    \end{minipage}
-    \caption{
-        Results of the simulation with 1000 devices.
-        %
-        The \emph{green} bar represents the mean power consumed by cloud;
-        the \emph{blue} bar represents the mean power consumed by all the devices in the system;
-        the \emph{yellow} bar represents the mean traveled distance by the people
-        and the \emph{purple} bar represents the average cost per hour of the cloud.
-        %
-        The chart shows two ``static'' scenarios: \emph{device-only} (on the left) and \emph{cloud-only} (on the right).
-        %
-        All the intermediate scenarios are obtained by exploiting the reconfiguration capabilities of the pulverization framework,
-        where different battery threshold are used to trigger the reconfiguration.
-    }
-    \label{fig:result-1000-devices}
-\end{figure}
+    \label{fig:results}
+\end{figure*}
+%
+Results are summarised in \Cref{fig:results}.
+%
+$\Downarrow$ has the worst \ac{qos}, but also the lowest cloud cost and overall energy consumption.
+%
+Opposedly, $\Uparrow$ improves the \ac{qos}, but costs and energy use grow significantly.
+%
+The ability to reconfigure the deployment at runtime
+by moving the behaviour from the cloud to the smartphones and vice versa
+depending on the battery level allows to find different trade-offs between these two extremes.
 

### 2023-05-20 paper-2023-acsos-pulverization.tex
write the conclusion

@@
 
-openness
-
-more platforms
-
-global policies
+In this work we presented a practical \ac{dsl} and framework for the
+runtime re-configuration of pulverised \ac{cps} applications.
+%
+The original idea of pulverisation was to define applications considering arbitrary networks of logical devices,
+decompose (pulverise) these logical devices into small deployment units,
+and then define their desired deployment at a later time.
+%
+Through the specialised \ac{dsl} introduced in this work
+we have shown that this idea can be realised in practice.
+%
+Moreover, we extended the original idea of pulverisation with the possibility to specify runtime re-configuration policies,
+which can be leveraged to achieve better trade-offs in terms of performance, cost, and resource usage.
+%
+We provide a real-world demonstrator for the technology,
+and we show via simulation that the approach can be beneficial.
+
+In the future,
+we plan to extend the \ac{dsl} to support openness and collective reconfiguration policies.
+%
+The former will allow devices to join and leave the system at runtime,
+while the latter will allow reconfigurations to be performed on the basis of a collective state of the system.
+%
+Concurrently,
+we intend to add support for additional network protocols to be able to support a larger number of practical scenarios.
 

### 2023-05-20 paper-2023-acsos-pulverization.tex
tackle @metaphori's meta

@@
 is then evaluated in a small-scale real-world demo and in a city-scale simulated scenario,
- \meta{showing that ???}.
+demonstrating the feasibility of the approach
+and its potential benefits in achieving better trade-offs between performance and resource utilisation.
 \end{abstract}

### 2023-05-20 paper-2023-acsos-pulverization.tex
add missing bold keywords

@@
   keywordstyle=[3]\bfseries{},
-  keywords=[3]{pulverizedSystem,device,and,deployableOn},
+  keywords=[3]{pulverizedSystem,device,and,deployableOn,startsOn,reconfigures,pulverizationRuntime,onDevice,reconfigurationRules},
 %  sensitive=true,

### 2023-05-20 paper-2023-acsos-pulverization.tex
highlight runtime dsk keywords

--- 2023-05-20 paper-2023-acsos-pulverization.tex
@@
   keywordstyle=[3]\bfseries{},
-  keywords=[3]{pulverizedSystem,device,and,deployableOn},
+  keywords=[3]{pulverizedSystem,device,and,deployableOn,pulverizationRuntime,startsOn,onDevice,reconfigurationRules,reconfigures,movesTo},
+%  emph={DeviceBehaviour,DeviceCommunication,DeviceSensors,DeviceActuators,HighLoad,LowBattery},
 %  sensitive=true,
@@
 
-\lstinputlisting[
+\begin{lstlisting}[
     float=ht,
-    language=Kotlin,
+    language=pulv,
     caption={
@@
     label=lst:runtime-configuration,
-]{listings/RuntimeConfig.kt}
+    numbers=left,
+]
+// Reconfiguration events
+expect fun cpuLoad(): Double
+object HighLoad : ReconfigurationEvent<Double>() {
+  override val predicate = { it > 0.90 }
+  override val events: Flow<Double> = cpuLoad()
+}
+object LowBattery : ReconfigurationEvent<Double>() {
+  override val predicate = { it < 0.20 }
+  override val events: Flow<Double> = cpuLoad()
+}
+// Available hosts
+object Smartphone : Host {
+  override val hostname = "smartphone"
+  override val capabilities = setOf(EmbeddedDevice)
+}
+object Server : Host {
+  override val hostname = "amazon-aws"
+  override val capabilities=setOf(HighCPU,LowLatencyComm)
+}
+// Runtime initial setup and runtime reconfiguration rules
+val infrastructure = setOf(Smartphone, Server)
+pulverizationRuntime(conf, "iot-sensor", infrastructure) {
+  DeviceBehaviour() startsOn Server
+  DeviceCommunication() startsOn Server
+  DeviceSensors() startsOn Smartphone
+  DeviceActuators() startsOn Smartphone
+  reconfigurationRules {
+    onDevice {
+      HighLoad reconfigures {Behaviour movesTo Smartphone}
+      LowBattery reconfigures {Behaviour movesTo Server}
+    }
+  }
+}
+\end{lstlisting}
 

### 2023-05-20 paper-2023-acsos-pulverization.tex
remove repetition

@@
     and the mapping of pulverised components to the underlying infrastructure
-    can be expressed expressively and, possibly, declaratively.
+    can be expressed easily and, possibly, declaratively.
 \end{itemize}
@@
     float=ht,
+    numbers=left,
     language=pulv,

### 2023-05-20 paper-2023-acsos-pulverization.tex
fix broken acronym

@@
 %
-The killer feature of internal \acp{dsl}, compared to external ones,
-is the counterpass of the limited syntax imposed by the host:
-if some peculiar entity or process cannot be express by the \ac{dsl},
-the designer can fall back to the host language directly,
-with no need to resort to interoperability features,
-and with the possibility to exploit all the libraries designed for the host language.
-%
-The internal DSL approach has thus been successfully applied in several domains,
+Compared to external ones
+internal \acp{dsl} allow the designer to fall back to the host language directly
+to implement peculiar entities or processes that cannot be express by the \ac{dsl}.
+%
+The approach has thus been successfully applied in several domains,
 spanning from
@@
 In this work we presented a practical \ac{dsl} and framework for the
-runtime re-configuration of pulverised \ac{cps} applications.
+runtime re-configuration of pulverised \ac{cas} applications.
 %

### 2023-05-20 paper-2023-acsos-pulverization.tex
move limitations in the fw and moderate the tone, @metaphori check

@@
 and it has been made available as part of the evaluation repositories.
-
-\subsection{Limitations of the current version of the DSL}
-
-In its current form, the proposed \ac{dsl} has two major limitations.
-%
-First and foremost, currently it does not consider openness:
-hosts get specified in the runtime and reconfiguration \ac{dsl} as a closed set,
-so the specification is not robust to the addition of new hosts or the loss of existing ones.
-%
-This is a major limitation that,
-although planned to be addressed shortly with minor additions to the \ac{dsl},
-must be highlighted in this manuscript.
-%
-The second limitation is related to the policies that can be expressed in the reconfiguration \ac{dsl}:
-currently, only policies that can be assessed at the local host level can be assessed.
-%
-A policy that requires to consider the state of multiple hosts would require to resort to the host language primitive mechanisms and make explicit communications outside the framework,
-thus breaking the abstraction.
-%
-This limitation can be addressed by extending the current reconfiguration
-block of the \ac{dsl} in such a way that queries can be performed on any reachable host of the system,
-thus making all communications
-(including reconfiguration-related ones)
-pass through the channels controlled by the framework.
 
@@
 
-In the future,
-we plan to extend the \ac{dsl} to support openness and collective reconfiguration policies.
-%
-The former will allow devices to join and leave the system at runtime,
-while the latter will allow reconfigurations to be performed on the basis of a collective state of the system.
-%
-Concurrently,
+In its current form, the proposed \ac{dsl} has two limitations that will be addressed in the future.
+%
+First and foremost, it currently does not consider openness,
+as hosts get specified in the runtime and reconfiguration \acp{dsl} as a closed set.
+%
+The second limitation is related to the policies that can be expressed in the reconfiguration \ac{dsl}:
+currently, only policies that can be assessed at the local host level can be assessed.
+%
+A policy that requires to consider the state of multiple hosts would require to resort to the host language primitive mechanisms and make explicit communications outside the framework,
+thus breaking the abstraction.
+%
+This limitation can be addressed by extending the current reconfiguration
+block of the \ac{dsl} in such a way that queries can be performed on any reachable host of the system,
+thus making all communications
+(including reconfiguration-related ones)
+pass through the channels controlled by the framework.
+%
+Concurrently with these extensions,
 we intend to add support for additional network protocols to be able to support a larger number of practical scenarios.

### 2023-05-20 paper-2023-acsos-pulverization.tex
save some space, fix some typos

@@
 We assume that the power consumption of the device is linearly dependent on the CPU usage,
-and we thus decided to estimate the \ac{epi}~\cite{Grochowski2006EnergyPI}
+and we thus decided to estimate the \ac{epi}~\cite{DBLP:conf/islped/ShaoB13}
 of common smartphone ($EPI_{device}$) and server ($EPI_{cloud}$) CPUs.
@@
 %
-
 Specifically,
- this work builds on the \emph{pulverisation} component model~\cite{FI2020-pulverization,IEEE-IoTJ-pulverization-simulation},
- whereby a large-scale cyber-physical system
- is partitioned into a graph of devices
- where each device is split into five deployable components:
- (i) sensors interface, 
- (ii) actuators interface, 
- (iii) behaviour, 
- (iv) state, and 
- (v) communication component.
+ this work builds on the \emph{pulverisation} component model~\cite{FI2020-pulverization,IEEE-IoTJ-pulverization-simulation}.
+% whereby a large-scale cyber-physical system
+% is partitioned into a graph of devices
+% where each device is split into five deployable components:
+% (i) sensors interface,
+% (ii) actuators interface,
+% (iii) behaviour,
+% (iv) state, and
+% (v) communication component.
 %
@@
  and, in particular, to \emph{multi-tier programming} paradigm~\cite{DBLP:journals/csur/WeisenburgerWS20},
- where the deployment units for different tiers (e.g., client and server tiers; or view, business logic, and data tiers) are obrained by compilation or interpretation of a single codebase.
+ where the deployment units for different tiers
+(e.g., client and server tiers; or view, business logic, and data tiers)
+are obtained by compilation or interpretation of a single codebase.
 
@@
 %
-Given a deployable software system developed using the techniques of the previous subsection,
+Given a deployable software system
+developed using the techniques of the previous subsection,
  what is needed now is a language to describe a deployment domain 
@@
 Languages have been proposed 
- targetting specific kinds of deployment domains
- such as the smart grid (cf.
-the deployment specification language \emph{dspec} in the \emph{Resilient Information Architecture Platform for Smart Grids [RIAPS]} platform~\cite{DBLP:conf/coins/GhoshTKKL22})
- or the cloud (cf. the \emph{Cloud Application Modelling and Execution Language [CAMEL]} multi-\ac{dsl}~\cite{DBLP:journals/jcloudc/AchilleosKRKDOS19}).
+ targeting specific kinds of deployment domains
+ such as the smart grid (cf. \emph{dspec} in the \emph{RIAPS} platform~\cite{DBLP:conf/coins/GhoshTKKL22})
+ or the cloud (cf. the \emph{CAMEL} multi-\ac{dsl}~\cite{DBLP:journals/jcloudc/AchilleosKRKDOS19}).
 
 In this work,
- we are especially interested deployments over large-scale cyber-physical systems and the edge-fog-cloud continuum.
-%
-The \emph{MultiScale Autonomic Deployment Language (MuScADeL)}~\cite{DBLP:conf/compsac/BoujbelRLTAL14/muscadel}
+ we are especially interested in deployments over large-scale cyber-physical systems and the edge-cloud continuum.
+%
+\emph{MuScADeL}~\cite{DBLP:conf/compsac/BoujbelRLTAL14/muscadel}
  is a \ac{dsl} that allows to express deployment properties of 
- componentised applications on multi-scale deployment domains, i.e., large and heterogeneous infrastructures considered under several viewpoints (device, geography, network, administrative domains, etc.).
+applications on multi-scale deployment domains,
+i.e., large and heterogeneous infrastructures considered under several viewpoints
+(device, geography, network, administrative domains, etc.).
 %
@@
   a model-based deployment approach is proposed
-  targetting so-called \emph{fleets}, i.e.,
-  distributed and heterogeneous devices at the edge chaaracterised by different cyber-physical contexts (cf. resources, connectivity, etc.).
-%
-It is based on the \emph{Generation and Deployment
-of Smart IoT Systems (GeneSIS)} modelling language, where a \texttt{DeploymentModel} consists of sets of \texttt{Resources} (e.g., \texttt{Component}s, specialised by \texttt{InfrastructureComponent}s and \texttt{SoftwareComponent}s) with associated \texttt{Property}s and \texttt{Link}s.
+  targeting so-called \emph{fleets}, i.e.,
+  distributed and heterogeneous devices at the edge characterised by different cyber-physical contexts
+(cf. resources, connectivity, etc.).
+%
+%It is based on the \emph{GeneSIS} modelling language,
+%where a \texttt{DeploymentModel} consists of sets of \texttt{Resources} (e.g., \texttt{Component}s, specialised by \texttt{InfrastructureComponent}s and \texttt{SoftwareComponent}s) with associated \texttt{Property}s and \texttt{Link}s.
 
@@
  on automatic deployment of distributed systems
- is provided in a 2015 survey by Arcangeli et al. \cite{DBLP:journals/jss/ArcangeliBL15}.
-%
-A more recent survey by Coullon et al. focusses
+ is provided in~\cite{DBLP:journals/jss/ArcangeliBL15}.
+%
+A more recent survey~\cite{coullon2023swreconfig} focuses
  on formal techniques for verifying the correctness of reconfigurations
- in component-based distributed software systems~\cite{coullon2023swreconfig}.
-%
-Two formal models for specifying reconfigurable architectures are \emph{DR-BIP (Dynamically Reconfigurable - Behaviour Interaction Priority model)}~\cite{Ballouli18dr-bip}
-and \emph{DReAM (Dynamically Reconfigurable Architectural Modelling)}~\cite{denicola2020dream-dynamic-reconfig-arch-modelling}.
+ in component-based distributed software systems.
+%
+Two formal models for specifying reconfigurable architectures are
+\emph{DR-BIP}~\cite{Ballouli18dr-bip}
+and \emph{DReAM}~\cite{denicola2020dream-dynamic-reconfig-arch-modelling}.
 %
 Both are based on the same conceptual model: they 
- feature \emph{components} (capturing behaviour), \emph{connectors} (capturing interaction between components' \emph{ports}), 
+ feature \emph{components} (capturing behaviour),
+\emph{connectors} (capturing interaction between components' \emph{ports}),
 \emph{maps} (logical topologies),
@@
 %
-\meta{TODO: say something about our positioning wrt verification of reconfigurations}
+%\meta{TODO: say something about our positioning wrt verification of reconfigurations}
 
@@
 %
-The \emph{ASSL (Autonomic System Specification Language)}~\cite{DBLP:conf/birthday/VassevH97}
+\emph{ASSL}~\cite{DBLP:conf/birthday/VassevH97}
  supports the specification
@@
 %
-The \emph{AWaRE DSL (ADSL)}~\cite{DBLP:conf/aswec/ChhetriLUVKNR18/adsl} supports constraint-based self-management, leveraging managing agents.
-The language enables the specification of the domain model (in terms of components), the problem structure model (in terms of constraints), the agent architecture model (in terms of agents, roles, and their coordination), and the assignment model (in terms of management problem assignment strategies).
+The \emph{AWaRE DSL}~\cite{DBLP:conf/aswec/ChhetriLUVKNR18/adsl} supports constraint-based self-management,
+leveraging managing agents.
+%The language enables the specification of the domain model (in terms of components),
+%the problem structure model (in terms of constraints),
+%the agent architecture model (in terms of agents, roles, and their coordination), and the assignment model (in terms of management problem assignment strategies).
 %

### 2023-05-20 paper-2023-acsos-pulverization.tex
we actually do have space

@@
 
-A framework/runtime configured via declarative files enforces declarativity at the configuration level,
+A framework/runtime with configuration files enforces declarativity at the configuration level,
 but its flexibility is limited,
@@
 In the former,
-we show a small-scale demonstrator of a safety system warning people when a room capacity is exceeded
+we show a small-scale demonstrator of a safety system warning people when a room capacity is exceeded,
 to showcase the integration of the framework with established technologies\footnote{
@@
 %
-The experiment has been open sourced\footnote{\url{https://github.com/nicolasfara/ACSOS-2023-pulverization-evaluation}}
+The experiment has been open sourced
 and the code has been archived on Zenodo~\cite{https://doi.org/10.5281/zenodo.7948801}.
@@
 %
-\begin{figure*}
-    \includegraphics[width=.32\linewidth]{figures/cloud_cost-device_consumption-cloud_consumption-distance-device=100.0.pdf}~
-    \includegraphics[width=.32\linewidth]{figures/cloud_cost-device_consumption-cloud_consumption-distance-device=300.0.pdf}~
-    \includegraphics[width=.32\linewidth]{figures/cloud_cost-device_consumption-cloud_consumption-distance-device=1000.0.pdf}
+\begin{figure}
+    \includegraphics[width=\columnwidth]{figures/cloud_cost-device_consumption-cloud_consumption-distance-device=100.0.pdf}\\
+    \includegraphics[width=\columnwidth]{figures/cloud_cost-device_consumption-cloud_consumption-distance-device=300.0.pdf}\\
+    \includegraphics[width=\columnwidth]{figures/cloud_cost-device_consumption-cloud_consumption-distance-device=1000.0.pdf}
     \caption{
-        Simulation results for 100 (left), 300 (center), and 1000 (right) devices;
+        Simulation results for 100 (top), 300 (center), and 1000 (bottom) devices;
         showing all four metrics:
@@
     \label{fig:results}
-\end{figure*}
+\end{figure}
 %
@@
  this work builds on the \emph{pulverisation} component model~\cite{FI2020-pulverization,IEEE-IoTJ-pulverization-simulation}.
-% whereby a large-scale cyber-physical system
-% is partitioned into a graph of devices
-% where each device is split into five deployable components:
-% (i) sensors interface,
-% (ii) actuators interface,
-% (iii) behaviour,
-% (iv) state, and
-% (v) communication component.
+ whereby a large-scale cyber-physical system
+ is partitioned into a graph of devices
+ where each device is split into five deployable components:
+ (i) sensors interface,
+ (ii) actuators interface,
+ (iii) behaviour,
+ (iv) state, and
+ (v) communication component.
 %
@@
 %
-%It is based on the \emph{GeneSIS} modelling language,
-%where a \texttt{DeploymentModel} consists of sets of \texttt{Resources} (e.g., \texttt{Component}s, specialised by \texttt{InfrastructureComponent}s and \texttt{SoftwareComponent}s) with associated \texttt{Property}s and \texttt{Link}s.
+It is based on the \emph{GeneSIS} modelling language,
+where a \texttt{DeploymentModel} consists of sets of \texttt{Resources} (e.g., \texttt{Component}s, specialised by \texttt{InfrastructureComponent}s and \texttt{SoftwareComponent}s) with associated \texttt{Property}s and \texttt{Link}s.
 
@@
 leveraging managing agents.
-%The language enables the specification of the domain model (in terms of components),
-%the problem structure model (in terms of constraints),
-%the agent architecture model (in terms of agents, roles, and their coordination), and the assignment model (in terms of management problem assignment strategies).
+The language enables the specification of the domain model (in terms of components),
+the problem structure model (in terms of constraints),
+the agent architecture model (in terms of agents, roles, and their coordination), and the assignment model (in terms of management problem assignment strategies).
 %

### 2023-05-20 paper-2023-acsos-pulverization.tex
fix typos in abstract

@@
 the infrastructure supporting the execution of situated distributed computations
-evolved with a fast pace.
+evolved at a fast pace.
 %
@@
 %
-The availability of such an infrastructure opens to better resource utilisation and performance,
+The availability of such an infrastructure opens to better resource utilisation and performance
 but, contextually, introduces new challenges to software designers,
@@
 meant to neatly separate business logic and deployment concerns,
-allowing applications to be defined independent of the infrastructure they will execute upon.
+allowing applications to be defined independently of the infrastructure they will execute upon.
 %

### 2023-05-21 paper-2023-acsos-pulverization.tex
Put refs to lines of code in prose for second DSL @danysk @nicolasfara

@@
 Using \Cref{lst:runtime-configuration} as a reference example,
-we show how reconfiguration events can be captured at the type-system level
+we show how reconfiguration events (Lines~\ref{code:reconfigev:start}--\ref{code:reconfigev:end}) can be captured at the type-system level
 by expressing them as a predicate over an asynchronous flow of \texttt{events} influencing the system's state.
 %
-Similarly, hosts can be configured by defining their name and capabilities.
+Similarly, hosts can be configured by defining their name and capabilities (Lines~\ref{code:hosts:start}--\ref{code:hosts:end}).
 %
 Once the hosts composing the system have been configured,
-the entire infrastructure can be set up by defining,
+the entire infrastructure can be set up (Lines~\ref{code:infra:start}--\ref{code:infra:end}) by defining,
 for each logical device,
-which host should host each pulverised component.
+which host should host each pulverised component (Lines~\ref{code:mapping:start}--\ref{code:mapping:end}).
 %
 Additionally,
-reconfiguration rules can be defined to specify how the system should react to reconfiguration events.
+reconfiguration rules (Lines~\ref{code:reconfig:start}--\ref{code:reconfig:end}) can be defined to specify how the system should react to reconfiguration events.
 %
@@
     numbers=left,
+    escapechar=\%
 ]
-// Reconfiguration events
+// Reconfiguration events %\label{code:reconfigev:start}%
 expect fun cpuLoad(): Double
@@
   override val events: Flow<Double> = cpuLoad()
-}
-// Available hosts
+}%\label{code:reconfigev:end}%
+// Available hosts %\label{code:hosts:start}%
 object Smartphone : Host {
@@
   override val capabilities=setOf(HighCPU,LowLatencyComm)
-}
-// Runtime initial setup and runtime reconfiguration rules
+}%\label{code:hosts:end}%
+// Runtime initial setup and runtime reconfiguration rules %\label{code:infra:start}%
 val infrastructure = setOf(Smartphone, Server)
 pulverizationRuntime(conf, "iot-sensor", infrastructure) {
-  DeviceBehaviour() startsOn Server
+  DeviceBehaviour() startsOn Server%\label{code:mapping:start}%
   DeviceCommunication() startsOn Server
   DeviceSensors() startsOn Smartphone
-  DeviceActuators() startsOn Smartphone
-  reconfigurationRules {
+  DeviceActuators() startsOn Smartphone%\label{code:mapping:end}%
+  reconfigurationRules {%\label{code:reconfig:start}%
     onDevice {
@@
     }
-  }
-}
+  }%\label{code:reconfig:end}%
+}%\label{code:infra:end}%
 \end{lstlisting}

### 2023-05-21 paper-2023-acsos-pulverization.tex
Renaming DSLs + some fixes @danysk @nicolasfara

@@
 
+\newcommand{\dslSys}[0]{\emph{System \ac{dsl}}}
+\newcommand{\dslDep}[0]{\emph{Deployment \ac{dsl}}}
+
 \section{A Runtime and \acsp{dsl} for Reconfigurable Pulverised Systems}\label{sec:contribution}
@@
 Coherently with the \emph{single responsibility principle},
-we capture these two descriptive needs into two distinct \acp{dsl}.
+we capture these two descriptive needs into two distinct \acp{dsl}: the \dslSys{} and the \dslDep{}.
 %
@@
 
-\subsection{Infrastructure and Requirements}
-
-The infrastructure \ac{dsl} captures the concepts of
+\subsection{The \dslSys{}: Components and Required Capabilities}
+
+The \dslSys{} captures the concepts of
 \emph{device type}, \emph{component}, \emph{capability}, and \emph{requirement}.
@@
     caption={
-        Example infrastructure configuration,
-        defining three capabilities and a system composed of two device types,
-        each component of which is then bound to the set of capabilities it needs to execute.
+        Example of \dslSys{} usage.
+        This code snippet defines three capabilities and a logical system composed of two device types,
+        whose components are bound to the set of capabilities they needs to execute.
     },
@@
 
-\subsection{Deployment and Reconfiguration}
-
-The second \ac{dsl} supports the definition of the mapping between pulverised components and specific hosts,
+\subsection{The \dslDep{}: Deployment Domain, Mapping, and Reconfiguration}
+
+The \dslDep{} supports the definition of the mapping between pulverised components and specific hosts,
 as well as the definition of reactive reconfiguration policies.

### 2023-05-21 paper-2023-acsos-pulverization.tex
fix footnote link
Bring II.C into III.A, + refactor presentation of the DSLs + provide a sentence clarifying the section structure -- @danysk @nicolasfara pls check

@@
 \DeclareSIUnit{\decibelmeter}{dBm}
+\providecommand{\decibel}{dB}
+\providecommand{\decibelmeter}{dBm}
+\providecommand{\qty}[2]{#1#2}
 
@@
 
-\subsection{Middlewares and (Domain-Specific) Languages for Deployed Systems Specification and Execution}
+\newcommand{\dslSys}[0]{\emph{System \ac{dsl}}}
+\newcommand{\dslDep}[0]{\emph{Deployment \ac{dsl}}}
+
+\section{A Runtime and \acsp{dsl} for Reconfigurable Pulverised Systems}\label{sec:contribution}
+
+From the previous section,
+we can identify three main pieces of information for the definition of a pulverised system and its deployment:
+\begin{enumerate}[label=\it{(\roman*)}]
+  \item the description of the logical system and its partitioning into components and deployment units;
+  \item the definition of the \emph{target} deployment domain (\emph{infrastructure}), i.e., the specification of the \emph{requirements} posed on hosts in terms of required (or preferred) capabilities; and
+  \item the definition of the \emph{deployment plan}, i.e., the allocation of the pulverised components to the infrastructural elements, together with constraints and reconfiguration policies.
+\end{enumerate}
+%
+%Coherently with the \emph{single responsibility principle},
+We capture these %two 
+ descriptive needs into two distinct \acp{dsl}: the \dslSys{} and the \dslDep{}.
+%
+An overall view of the approach in terms of the provided tools and the main modelling concepts is provided in \Cref{fig:approach}.
+
+In the following,
+ we first clarify the approach (\Cref{sec:contrib:general}),
+ then cover the two \acp{dsl}
+ (\Cref{sec:contrib:dslsys,sec:contrib:dsldep}),
+ and finally provide details on the runtime system (\Cref{sec:contrib:runtime}).
+
+\begin{figure}
+\includegraphics[width=\columnwidth]{figures/pulverization-framework-overview.drawio.pdf}
+\caption{The approach provides two \acp{dsl} and a runtime for specifying and executing deployment plans.
+Each deployment unit will be delivered to the target site and properly configured to interact with other deployment units, as required by the application at hand (see  \Cref{sec:evaluation} for examples).}
+\label{fig:approach}
+\end{figure}
+
+\subsection{On Middlewares and (Domain-Specific) Languages for Deployed Systems Specification and Execution}
+\label{sec:contrib:general}
 
@@
 
-\newcommand{\dslSys}[0]{\emph{System \ac{dsl}}}
-\newcommand{\dslDep}[0]{\emph{Deployment \ac{dsl}}}
-
-\section{A Runtime and \acsp{dsl} for Reconfigurable Pulverised Systems}\label{sec:contribution}
-
-From the previous section,
-we can identify two main pieces of information for the definition of a deployment plan for a pulverised system:
-\begin{enumerate*}[label=\it{(\roman*)}]
-  \item the definition of the \emph{target} deployment domain (\emph{infrastructure}), i.e., the specification of the \emph{requirements} posed on hosts in terms of required (or preferred) capabilities; and
-  \item the definition of the \emph{deployment plan}, i.e., the allocation of the pulverised components to the infrastructural elements, together with constraints and reconfiguration policies.
-\end{enumerate*}
-%
-Coherently with the \emph{single responsibility principle},
-we capture these two descriptive needs into two distinct \acp{dsl}: the \dslSys{} and the \dslDep{}.
-%
-An overall view of the approach in terms of the provided tools and the main modelling concepts is provided in \Cref{fig:approach}.
-
-\begin{figure}
-\includegraphics[width=\columnwidth]{figures/pulverization-framework-overview.drawio.pdf}
-\caption{The approach provides two \acp{dsl} and a runtime for specifying and executing deployment plans.
-Each deployment unit will be delivered to the target site and properly configured to interact with other deployment units, as required by the application at hand (see  \Cref{sec:evaluation} for examples).}
-\label{fig:approach}
-\end{figure}
-
 \subsection{The \dslSys{}: Components and Required Capabilities}
+\label{sec:contrib:dslsys}
 
@@
 \subsection{The \dslDep{}: Deployment Domain, Mapping, and Reconfiguration}
+\label{sec:contrib:dsldep}
 
@@
 \subsection{Implementation details and underlying runtime}
+\label{sec:contrib:runtime}
 
@@
 %
-The framework has been open sourced and released with a permissive license\footnote{\url{https://github.com/nicolasfara/pulverization-framework}}.
+The framework has been open sourced and released with a permissive license\footnote{\url{https://github.com/pulvreakt/pulvreakt}}.
 

### 2023-05-21 paper-2023-acsos-pulverization.tex
add a more detailed description of the implementation details of the framework. Please check @metaphori @danysk

@@
 The framework has been open sourced and released with a permissive license\footnote{\url{https://github.com/pulvreakt/pulvreakt}}.
+
+\meta{
+The runtime bases its operation on three main components:
+the \texttt{Spawner}, which is responsible for executing the components;
+the \texttt{UnitReconfigurator}, which is responsible for reconfigure the \emph{deployment unit}
+according to the \texttt{ReconfigurationEvent}
+and the \texttt{Communicator}s which are used to establish the components' intra-communication.
+%
+To start, the \texttt{PulverisationRuntime} needs to know the name of the \emph{logical device},
+the \emph{host} on which is running, and the configurations produced by the two \ac{dsl}s.
+%
+At the start of the system the runtime,
+based on the provided configuration,
+spawn all the components that must start on the \emph{deployment unit}.
+%
+Similarly, the \texttt{UnitReconfigurator} start observing all the configured \texttt{ReconfigurationEvent}s,
+and register the listeners for incoming reconfigurations.
+%
+Each spawned component start to execute its logic and share data with other component via the \texttt{Communicator},
+which is responsible for the components' intra-communication by leveraging the communication protocol.
+%
+When a reconfiguration occurs, the \texttt{UnitReconfigurator} takes into account the new deployment configuration
+by interacting with the \texttt{Spawner} to start or stop the required components.
+}
 

### 2023-05-21 paper-2023-acsos-pulverization.tex
@nicolasfara ok

--- 2023-05-21 paper-2023-acsos-pulverization.tex
@@
 
+The runtime bases its operation on three main components:
+the \texttt{Spawner}, which is responsible for executing the components;
+the \texttt{UnitReconfigurator}, which is responsible for reconfigure the \emph{deployment unit}
+according to the \texttt{ReconfigurationEvent}
+and the \texttt{Communicator}s which are used to establish the components' intra-communication.
+%
+To start, the \texttt{PulverisationRuntime} needs to know the name of the \emph{logical device},
+the \emph{host} on which is running, and the configurations produced by the two \ac{dsl}s.
+%
+At the start of the system the runtime,
+based on the provided configuration,
+spawns all the components that must start on the \emph{deployment unit}.
+%
+Similarly, the \texttt{UnitReconfigurator} starts observing all the configured \texttt{ReconfigurationEvent}s,
+and registers the listeners for incoming reconfigurations.
+%
+Each spawned component starts to execute its logic and shares data with other component via the \texttt{Communicator},
+which is responsible for the components' intra-communication by leveraging the communication protocol.
+%
+When a reconfiguration occurs, the \texttt{UnitReconfigurator} takes into account the new deployment configuration,
+by interacting with the \texttt{Spawner}, to start or stop the required components.
+
 The configuration produced by the \ac{dsl} needs a platform that instructs the system on the way actual communication needs to be performed.

### 2023-05-21 paper-2023-acsos-pulverization.tex
Minors

@@
 %
-The platform also feature an optimization for pulverized components of the same logical device ending up running on the same host:
+The platform also features an optimisation for pulverised components of the same logical device ending up running on the same host:
 as it would be expected,

### 2023-05-21 paper-2023-acsos-pulverization.tex
Fix listing 1 and 2 @metaphori @danysk

@@
 
-pulverizedSystem {
+val conf = pulverizedSystem {
   device("control-center") {%\label{devcap:start}%
@@
 }%\label{code:infra:end}%
+pulverizationRuntime(conf, "control-center", infrastructure) {
+  /* ... */
+}
 \end{lstlisting}

### 2023-05-21 paper-2023-acsos-pulverization.tex
fix listing 2

@@
 expect fun cpuLoad(): Double
+expect fun batteryLevel(): Double
 object HighLoad : ReconfigurationEvent<Double>() {
@@
   override val predicate = { it < 0.20 }
-  override val events: Flow<Double> = cpuLoad()
+  override val events: Flow<Double> = batteryLevel()
 }%\label{code:reconfigev:end}%

### 2023-05-21 paper-2023-acsos-pulverization.tex
fix type annotation in listing 2

--- 2023-05-21 paper-2023-acsos-pulverization.tex
@@
 // Reconfiguration events %\label{code:reconfigev:start}%
-expect fun cpuLoad(): Double
+expect fun cpuLoad(): Flow<Double>
+expect fun batteryLevel(): Flow<Double>
 object HighLoad : ReconfigurationEvent<Double>() {
   override val predicate = { it > 0.90 }
-  override val events: Flow<Double> = cpuLoad()
+  override val events = cpuLoad()
 }
@@
   override val predicate = { it < 0.20 }
-  override val events: Flow<Double> = cpuLoad()
+  override val events = batteryLevel()
 }%\label{code:reconfigev:end}%

### 2023-05-21 paper-2023-acsos-pulverization.tex
Magic @nicolasfara

--- 2023-05-21 paper-2023-acsos-pulverization.tex
@@
 
+\makeatletter
+\newcommand\currentStyle@lstparam{}
+\lst@AddToHook{Output}{\global\let\currentStyle@lstparam\lst@thestyle}
+\lst@AddToHook{OutputOther}{\global\let\currentStyle@lstparam\lst@thestyle}
+\makeatother
+
+\makeatletter
+\newcommand{\crefe}[1]{\currentStyle@lstparam\Cref{#1}} %<-- try to change that to `xs'
+\makeatother
+
 \begin{lstlisting}[
@@
 val infrastructure = setOf(Smartphone, Server)
+val conf = // see %\crefe{lst:system-configuration}%
 pulverizationRuntime(conf, "iot-sensor", infrastructure) {

### 2023-05-21 paper-2023-acsos-pulverization.tex
Minors on bib

@@
 The general solution is to devise 
- a reasonable partitioning of the software system
+ a reasonable \emph{partitioning}~\cite{DBLP:journals/jnca/LiuASGBQ15} of the software system
  and a corresponding (dynamic) deployment plan 

### 2023-05-21 paper-2023-acsos-pulverization.tex
Minor

@@
  the approach of specifying in a single codebase 
- parts of the structure, behaviour, and/or interaction 
- for a whole distributed system
+ parts of the structure, behaviour, interaction 
+ and/or other aspects of a whole distributed system
  is also related to \emph{macroprogramming}~\cite{Casadei2023macro}

### 2023-05-21 paper-2023-acsos-pulverization.tex
RW: needs to mention how our approach fits or differs @nicolasfara @danysk some bits - pls check

@@
  for generating and testing deployments through simulation.
+%
+There, file descriptors and scripts are used to generate deployment plans; in \ourframework{}, instead, we provide \acp{dsl} supporting a specification-oriented, declarative approach.
+%
 
@@
 are obtained by compilation or interpretation of a single codebase.
+%
+\ourframework{} adopts the same idea,
+ where a logically centralised specification
+ of the distributed architecture
+ enables runtime checks and deployment activities.
 

### 2023-05-21 paper-2023-acsos-pulverization.tex
Capitalisation of titles @danysk pls check that I also retitled subsections in the eval (btw I also notice that it is not clear what the second scenario is all about... it is quite a generic 'a large-scale scenario where a crowd of people is attending an event in a city

@@
 
-\subsection{Implementation details and underlying runtime}
+\subsection{Implementation Details and the Runtime System}
 \label{sec:contrib:runtime}
@@
 
-\subsection{Small scale: crowd alert}
+\subsection{Small-Scale Scenario: Crowd Alert in a Lab}
 \label{sec:small-scale-crowd-alert}
@@
 
-\subsection{Large scale: urban event}
+\subsection{Large-Scale Scenario: Collective Computation at a Urban Event}
 
@@
 
-\subsection{Application description languages: component-based software engineering and the pulverisation model}
+\subsection{Application Description Languages: Component-based Software Engineering and the Pulverisation Model}
 \label{sec:rw:whatsw}
@@
 
-\subsection{Infrastructure and deployment description languages}
+\subsection{Infrastructure and Deployment Description Languages}
 \label{sec:rw:depdesc}
@@
 
-\subsection{Automatic/autonomic deployment and reconfiguration}
+\subsection{Automatic/Autonomic Deployment and Reconfiguration}
 \label{sec:rw:autodep}

### 2023-05-21 paper-2023-acsos-pulverization.tex
'Alert' is not a device -> alarm

@@
 
-The system is composed of two kinds of device: \emph{wearable} and \emph{alert}.
+The system is composed of two kinds of device: \emph{wearable} and \emph{alarm}.
 %
@@
 sensors ($\sigma^w$), perceiving other devices by their signal strength;
-behaviour ($\beta^w$), converting raw data from the sensor into messages for the alert system; and
-communication ($\chi^w$) sending data to the alert system.
+behaviour ($\beta^w$), converting raw data from the sensor into messages for the alarm; and
+communication ($\chi^w$) sending data to the alarm.
 %
@@
 %
-The alert consists of the following pulverized components:
+The alarm consists of the following pulverized components:
 $\beta^a$, computing the mean distance among devices;
@@
 $\sigma^{w}$ and $\chi^{w}$ are executed on the \emph{Smartphones},
-and all the \emph{alert} components are executed in the \emph{Monitor}.
+and all the \emph{alarm} components are executed in the \emph{Monitor}.
 %

### 2023-05-21 paper-2023-acsos-pulverization.tex
Assure the reader we'll explain those terms

@@
 %
-An overall view of the approach in terms of the provided tools and the main modelling concepts is provided in \Cref{fig:approach}.
+An overall view of the approach in terms of the provided tools and the main modelling concepts is provided in \Cref{fig:approach}: such elements are explained in the following sections.
 

### 2023-05-21 paper-2023-acsos-pulverization.tex
add a little description of the deployment unit. Pleas check @metaphori and @danysk

@@
 and \texttt{ReconfigurationRule}.
+%
+\meta{
+The main goal of the \dslDep is to configure the \emph{deployment unit},
+which represents the smallest deployable unit of the pulverised system.
+%
+It is characterized by a set of \emph{pulverised components},
+the \emph{host} on which they are deployed,
+and the \emph{logical device} they represent. 
+}
 %

### 2023-05-21 paper-2023-acsos-pulverization.tex
fix typo

@@
 \meta{
-The main goal of the \dslDep is to configure the \emph{deployment unit},
+The main goal of the \dslDep{} is to configure the \emph{deployment unit},
 which represents the smallest deployable unit of the pulverised system.

### 2023-05-21 paper-2023-acsos-pulverization.tex
RW: minors

@@
  or frameworks.
-%
+
 Specifically,
@@
 %
-It is based on the \emph{GeneSIS} modelling language,
+The main goal is to support the distribution of \emph{multiple variants} of the same application based on the different contexts of the different target devices.
+%
+The approach based on the \emph{GeneSIS} modelling language,
 where a \texttt{DeploymentModel} consists of sets of \texttt{Resources} (e.g., \texttt{Component}s, specialised by \texttt{InfrastructureComponent}s and \texttt{SoftwareComponent}s) with associated \texttt{Property}s and \texttt{Link}s.

### 2023-05-21 paper-2023-acsos-pulverization.tex
abstract tweaks

--- 2023-05-21 paper-2023-acsos-pulverization.tex
@@
 The availability of such an infrastructure opens to better resource utilisation and performance
-but, contextually, introduces new challenges to software designers,
-as applications must be conceived to be able to adapt their deployment to changing deployment domains and conditions.
+but, at the same time, introduces new challenges to software designers,
+as applications must be conceived to be able to adapt to changing deployment domains and conditions.
 %
@@
 %
-The framework is based on a domain-specific language capturing,
-in a declarative fashion,
+The framework is based on a domain-specific language capturing, in a declarative fashion:
 pulverised application components, device capabilities, resource allocation, and (runtime re-) configuration policies.

### 2023-05-21 paper-2023-acsos-pulverization.tex
Restructure RW

@@
  languages for 
- specifying distributed systems (\Cref{sec:rw:whatsw}),
- languages for specifying infrastructures and deployments (\Cref{sec:rw:depdesc}),
+ specifying distributed systems (\Cref{sec:rw:whatsw})
  and
- techniques for automatic deployment and reconfiguration (\Cref{sec:rw:autodep}).
+ languages for specifying infrastructures, deployments, and reconfigurations (\Cref{sec:rw:depdesc}). 
+% and
+% techniques for automatic deployment and reconfiguration (\Cref{sec:rw:autodep}).
 
@@
  or the cloud (cf. the \emph{CAMEL} multi-\ac{dsl}~\cite{DBLP:journals/jcloudc/AchilleosKRKDOS19}).
-
+%
 In this work,
  we are especially interested in deployments over large-scale cyber-physical systems and the edge-cloud continuum.
-%
-\emph{MuScADeL}~\cite{DBLP:conf/compsac/BoujbelRLTAL14/muscadel}
- is a \ac{dsl} that allows to express deployment properties of 
+
+A related \ac{dsl} is 
+\emph{MuScADeL}~\cite{DBLP:conf/compsac/BoujbelRLTAL14/muscadel}: allows to express deployment properties of 
 applications on multi-scale deployment domains,
@@
 %
-In~\cite{DBLP:journals/sosym/SongDFSF22},
-  a model-based deployment approach is proposed
-  targeting so-called \emph{fleets}, i.e.,
-  distributed and heterogeneous devices at the edge characterised by different cyber-physical contexts
-(cf. resources, connectivity, etc.).
-%
-The main goal is to support the distribution of \emph{multiple variants} of the same application based on the different contexts of the different target devices.
-%
-The approach based on the \emph{GeneSIS} modelling language,
-where a \texttt{DeploymentModel} consists of sets of \texttt{Resources} (e.g., \texttt{Component}s, specialised by \texttt{InfrastructureComponent}s and \texttt{SoftwareComponent}s) with associated \texttt{Property}s and \texttt{Link}s.
-
-\subsection{Automatic/Autonomic Deployment and Reconfiguration}
-\label{sec:rw:autodep}
+Component descriptions can include \emph{constraints} (both basic software/hardware constraint and multi-scale criteria).
+%
+The specifications are used to generate \emph{probe} artifacts supporting data collection at the level of the deployment middleware.
+%
+However, the work in~\cite{DBLP:conf/compsac/BoujbelRLTAL14/muscadel} is preliminary and does not cover runtime adaptation.
+
+%\subsection{Automatic/Autonomic Deployment and Reconfiguration}
+%\label{sec:rw:autodep}
 %
@@
 %
-Two formal models for specifying reconfigurable architectures are
+Indeed, formal models can help in specifying reconfigurable architectures: examples include 
 \emph{DR-BIP}~\cite{Ballouli18dr-bip}
@@
 
-In the \emph{osmotic computing} approach~\cite{DBLP:journals/computer/VillariFDRJR19},
- microservices (the solvent)
- can migrate across 
- the edge-cloud infrastructure (the solution),
- passing through layer boundaries (the semi-permeable membranes),
- in order to keep a balance in the desired properties (the solute).
-%
-\emph{ASSL}~\cite{DBLP:conf/birthday/VassevH97}
- supports the specification
- of autonomic elements,
- interaction protocols,
- and \emph{self-management policies} based on fluents (states), events, and actions.
+A number of approaches use languages together with middlewares to specify and execute reconfiguration policies.
+%
+%\emph{ASSL}~\cite{DBLP:conf/birthday/VassevH97}
+% supports the specification
+% of autonomic elements,
+% interaction protocols,
+% and \emph{self-management policies} based on fluents (states), events, and actions.
 %
@@
 %
-These approaches, however, typically suffer from scalability issues.
+%
+In~\cite{DBLP:journals/sosym/SongDFSF22},
+  a model-based deployment approach is proposed
+  targeting so-called \emph{fleets}, i.e.,
+  distributed and heterogeneous devices at the edge characterised by different cyber-physical contexts
+(cf. resources, connectivity, etc.).
+%
+The main goal is to support the distribution of \emph{multiple variants} of the same application based on the different contexts of the different target devices.
+%
+The approach based on the \emph{GeneSIS} modelling language,
+where a \texttt{DeploymentModel} consists of sets of \texttt{Resources} (e.g., \texttt{Component}s, specialised by \texttt{InfrastructureComponent}s and \texttt{SoftwareComponent}s) with associated \texttt{Property}s and \texttt{Link}s.
+%
+These approaches, however, being based on constraint solving, typically suffer from scalability issues.
+
+Another approach is \emph{osmotic computing}~\cite{DBLP:journals/computer/VillariFDRJR19},
+ whereby microservices (the solvent)
+ can migrate across 
+ the edge-cloud infrastructure (the solution),
+ passing through layer boundaries (the semi-permeable membranes),
+ in order to keep a balance in the desired properties (the solute).
+%
 

### 2023-05-21 paper-2023-acsos-pulverization.tex
fix author affiliations

@@
 
-\author{\IEEEauthorblockN{1\textsuperscript{st} Danilo Pianini}
+\author{
+\IEEEauthorblockN{
+    Nicolas Farabegoli\IEEEauthorrefmark{1},
+    Danilo Pianini\IEEEauthorrefmark{1},
+    Roberto Casadei\IEEEauthorrefmark{2},
+    Mirko Viroli\IEEEauthorrefmark{1}
+}
 \IEEEauthorblockA{\textit{Department of Computer Science and Engineering} \\
@@
 Cesena, Italy \\
-danilo.pianini@unibo.it}
-\and
-\IEEEauthorblockN{2\textsuperscript{nd} Roberto Casadei}
-\IEEEauthorblockA{\textit{Department of Computer Science and Engineering} \\
-\textit{Alma Mater Studiorum---Università di Bologna}\\
-Cesena, Italy \\
-roby.casadei@unibo.it}
-\linebreakand
-\IEEEauthorblockN{3\textsuperscript{rd} Nicolas Farabegoli}
-\IEEEauthorblockA{\textit{Department of Computer Science and Engineering} \\
-\textit{Alma Mater Studiorum---Università di Bologna}\\
-Cesena, Italy \\
-nicolas.farabegoli@unibo.it}
-\and
-\IEEEauthorblockN{4\textsuperscript{th} Mirko Viroli}
-\IEEEauthorblockA{\textit{Department of Computer Science and Engineering} \\
-\textit{Alma Mater Studiorum---Università di Bologna}\\
-Cesena, Italy \\
-mirko.viroli@unibo.it}
+\IEEEauthorrefmark{1}\{nicolas.farabegoli,danilo.pianini,mirko.viroli\}@unibo.it}
+\IEEEauthorrefmark{2}roby.casadei@unibo.it
 }

### 2023-05-21 paper-2023-acsos-pulverization.tex
intro

--- 2023-05-21 paper-2023-acsos-pulverization.tex
@@
 %
-However, exploiting this infrastructure,
+However, exploiting this infrastructure
  poses new challenges to \ac{cas} application designers.
@@
  called \emph{pulverisation}~\cite{FI2020-pulverization},
- the fosters the decoupling of business logic and deployment concerns.
+ that fosters the decoupling of business logic and deployment concerns.
 %
@@
 In this work,
-we continue the research on pulverised systems by providing two key contributions:
-\textbf{first}, we extend the theoretical framework of pulverisation by adding support for
+we further develop the research on pulverised systems by providing two key contributions:
+%
+\begin{enumerate}
+ \item we extend the theoretical framework of pulverisation by adding support for
 \emph{runtime configuration rules}, allowing pulverised components to move at runtime across different hosts;
-\textbf{second}, we provide a practical implementation in Kotlin called \emph{\ourframework{}} which,
-to the best of our knowledge,
-is the first \emph{practical framework supporting pulverisation}. % for a general-purpose language.
+ \item we accordigly provide a practical implementation of pulverisation (with reconfiguration rules) called \emph{\ourframework{}}, developed in the Kotlin programming language.
+\end{enumerate}
 %

### 2023-05-21 paper-2023-acsos-pulverization.tex
poor @metaphori bring him back into the group

@@
 \IEEEauthorblockN{
-    Nicolas Farabegoli\IEEEauthorrefmark{1},
-    Danilo Pianini\IEEEauthorrefmark{1},
-    Roberto Casadei\IEEEauthorrefmark{2},
-    Mirko Viroli\IEEEauthorrefmark{1}
+    Nicolas Farabegoli,
+    Danilo Pianini,
+    Roberto Casadei,
+    Mirko Viroli
 }
@@
 Cesena, Italy \\
-\IEEEauthorrefmark{1}\{nicolas.farabegoli,danilo.pianini,mirko.viroli\}@unibo.it}
-\IEEEauthorrefmark{2}roby.casadei@unibo.it
+\{nicolas.farabegoli, danilo.pianini, roby.casadei, mirko.viroli\}@unibo.it}
 }

### 2023-05-21 paper-2023-acsos-pulverization.tex
✂ improve conciseness

@@
  are extending 
- both the \emph{kinds of applications and systems} that are being addressed  
- as well as their \emph{supporting infrastructure}~\cite{DBLP:journals/iot/GillXOPBSGSWASM22}.
+ both the \emph{kinds of applications and systems} being addressed and
+their \emph{supporting infrastructure}~\cite{DBLP:journals/iot/GillXOPBSGSWASM22}.
 %

### 2023-05-21 paper-2023-acsos-pulverization.tex
typos, punctuations, language, conciseness

@@
  are extending 
- both the \emph{kinds of applications and systems} being addressed and
+ the \emph{kinds of applications and systems} being addressed and
 their \emph{supporting infrastructure}~\cite{DBLP:journals/iot/GillXOPBSGSWASM22}.
@@
 a multi-layer heterogeneous network of devices (ranging from large and powerful cloud servers to small connected \emph{things})
-on which software can compute, store, and exchange data in a distributed fashion, while 
+on which software can compute, store, and exchange data in a distributed fashion
+while
 optimising for performance and resource utilisation.
@@
 %
-Commonly, in fact, applications are designed with a specific infrastructure in mind,
-whose assumptions unavoidably leak into the application logic~\cite{Spolsky2004},
+Commonly, applications are designed with a specific infrastructure in mind,
+whose assumptions unavoidably leak into the application logic~\cite{Spolsky2004}
 unless captured and encapsulated away from it:
@@
 %
-In short, the core idea is to consider devices as \emph{logical} entities,
+In short, the core idea is to consider devices as \emph{logical} entities
 whose software can be designed as (or broken down to, if already existing) an ensemble of \emph{five components
@@
 This way, the designer can focus on the business logic at hand,
-delaying an optimised deployment to later stages of the sofware design,
+delaying an optimised deployment to later stages of the software design,
 thus gaining resilience to changes in the infrastructure or in non-functional requirements.
@@
 %
-It is characterized by a set of \emph{pulverised components},
+It is characterised by a set of \emph{pulverised components},
 the \emph{host} on which they are deployed,
@@
 %
-The alarm consists of the following pulverized components:
+The alarm consists of the following pulverised components:
 $\beta^a$, computing the mean distance among devices;
@@
 in the same artifact,
-a simulation of the system where every device is containerized.
+a simulation of the system where every device is containerised.
 

### 2023-05-21 paper-2023-acsos-pulverization.tex
intro with 3 contributions

--- 2023-05-21 paper-2023-acsos-pulverization.tex
@@
 \IEEEauthorblockN{
-    Nicolas Farabegoli,
-    Danilo Pianini,
-    Roberto Casadei,
-    Mirko Viroli
+    Nicolas Farabegoli\IEEEauthorrefmark{1},
+    Danilo Pianini\IEEEauthorrefmark{1},
+    Roberto Casadei\IEEEauthorrefmark{2},
+    Mirko Viroli\IEEEauthorrefmark{1}
 }
@@
 Cesena, Italy \\
-\{nicolas.farabegoli, danilo.pianini, roby.casadei, mirko.viroli\}@unibo.it}
+\IEEEauthorrefmark{1}\{nicolas.farabegoli,danilo.pianini,mirko.viroli\}@unibo.it}
+\IEEEauthorrefmark{2}roby.casadei@unibo.it
 }
@@
  it is mounting the idea of the \emph{cloud-edge continuum}~\cite{DBLP:journals/iot/BittencourtISFM18}: 
-a multi-layer heterogeneous network of devices (ranging from large and powerful cloud servers to small connected \emph{things})
-on which software can compute, store, and exchange data in a distributed fashion, while 
+a multi-layer heterogeneous network of devices (ranging from large and powerful cloud servers to small connected \emph{things}).
+There, software can compute, store, and exchange data in a distributed fashion, while also 
 optimising for performance and resource utilisation.
@@
 %
-Commonly, in fact, applications are designed with a specific infrastructure in mind,
-whose assumptions unavoidably leak into the application logic~\cite{Spolsky2004},
+In fact, typically applications are designed with a specific infrastructure in mind,
+whose assumptions unavoidably leak into the application logic~\cite{Spolsky2004}
 unless captured and encapsulated away from it:
@@
 In the context of \acp{cas},
- previous work proposes an application partitioning schema,
+ previous works propose an application partitioning schema,
  called \emph{pulverisation}~\cite{FI2020-pulverization},
@@
 %
-In short, the core idea is to consider devices as \emph{logical} entities,
+The core idea is to consider devices as \emph{logical} entities,
 whose software can be designed as (or broken down to, if already existing) an ensemble of \emph{five components
 (behaviour, state, communication, sensors, and actuators)}
- that can be deployed with flexibility on available infrastructure without (ideally) affecting application functionality. %whose actual host can be decided at deployment time.
+ that can be deployed with flexibility on the available infrastructure without affecting application functionality. %whose actual host can be decided at deployment time.
 %
@@
 although the application can be deployed on arbitrary systems,
-there is no support for the reconfiguration of components at runtime, as a reaction to changing conditions or requirements.
+there is no support for the reconfiguration of components at runtime, as a reaction to changing requirements or workload conditions.
 
 In this work,
-we further develop the research on pulverised systems by providing two key contributions:
-%
-\begin{enumerate}
+we further develop the research on pulverised systems by providing three key contributions:
+%
+\begin{enumerate}[label=\it{(\roman*)}]
  \item we extend the theoretical framework of pulverisation by adding support for
 \emph{runtime configuration rules}, allowing pulverised components to move at runtime across different hosts;
- \item we accordigly provide a practical implementation of pulverisation (with reconfiguration rules) called \emph{\ourframework{}}, developed in the Kotlin programming language.
+ \item we provide a practical implementation of pulverisation (with reconfiguration rules) called \emph{\ourframework{}}, developed in the Kotlin programming language.
+ \item we provide an evaluatuation of \ourframework{}, exercising it through two case studies:
+a small-scale real-wold demonstration, and a city-scale simulated scenario where we seek for achieving balance between energy consumption and \ac{qos}.
 \end{enumerate}
-%
-To evaluate \ourframework{}, we exercise it through two case studies:
-a small-scale real-wold demonstration, and a city-scale simulated scenario.
 

### 2023-05-21 paper-2023-acsos-pulverization.tex
typos, punctuations, language, conciseness, clarity

--- 2023-05-21 paper-2023-acsos-pulverization.tex
@@
  it is mounting the idea of the \emph{cloud-edge continuum}~\cite{DBLP:journals/iot/BittencourtISFM18}: 
-a multi-layer heterogeneous network of devices (ranging from large and powerful cloud servers to small connected \emph{things})
-on which software can compute, store, and exchange data in a distributed fashion
+a multi-layer heterogeneous network of devices (ranging from large and powerful cloud servers to small connected \emph{things}).
+There, software can compute, store, and exchange data in a distributed fashion
 while

### 2023-05-21 paper-2023-acsos-pulverization.tex
typos, punctuations, language, conciseness, clarity

@@
 The \emph{deployment view} is a well-known architectural viewpoint for software systems,
-concerned with the mapping of \emph{software components} to \emph{physical machines}~\cite{DBLP:journals/software/Kruchten95}, 
+concerned with the mapping of \emph{software components} to \emph{physical machines}~\cite{DBLP:journals/software/Kruchten95}
  and supported by modelling notations like \emph{UML Deployment diagrams}.

### 2023-05-21 paper-2023-acsos-pulverization.tex
typos, punctuations, language, conciseness, clarity

@@
 %
-Here, we provide a brief introduction to deployment and reconfiguration, based on the conceptual characterisation of~\cite{DBLP:journals/jss/ArcangeliBL15,carzaniga1998characterization}.
+Here, we briefly introduce deployment and reconfiguration based on the conceptual characterisation of~\cite{DBLP:journals/jss/ArcangeliBL15,carzaniga1998characterization}.
 %

### 2023-05-21 paper-2023-acsos-pulverization.tex
typos, punctuations, language, conciseness, clarity

@@
 %
-Common deployment-related activities 
- include: (i) \emph{release/update} of the software system at the producer site;
- (ii) \emph{installation/deinstallation} at/from the consumer sites;
+Common deployment-related activities include:
+(i) \emph{release/update} of the software system at the producer site;
+(ii) \emph{installation/uninstallation} at/from the consumer sites;
 (iii) \emph{activation/deactivation}, for starting/stopping the components;

### 2023-05-21 paper-2023-acsos-pulverization.tex
typos, punctuations, language, conciseness, clarity

@@
 According to the analytical framework of~\cite{DBLP:journals/jss/ArcangeliBL15},
-  the problem of (automatic) deployment of distributed software system can be addressed by considering
+  the problem of (automatic) deployment of distributed software systems can be addressed by considering
  (i) the nature of the software to be deployed (e.g., how the software system is split into components);
@@
 %
-By an operational point of view, 
+From an operational point of view,
  deployment may be supported by a so-called \emph{runtime} or \emph{middleware}~\cite{DBLP:journals/cacm/GazisK22}.

### 2023-05-21 paper-2023-acsos-pulverization.tex
typos, punctuations, language, conciseness, clarity

@@
 Notation: solid-border boxes denote physical hosts (bold borders are for thick devices); solid lines denote connections between hosts; dashed-border boxes denote software components; %, with letters indicating their function (\textbf{B}ehaviour, \textbf{C}ommunication, \textbf{K}nowledge, \textbf{S}ensors, \textbf{A}ctuators);
-differently colours denote (software components of) different logical devices;
+different colours denote (software components of) different logical devices;
 red dashed lines denote connections between the software components, i.e., neighbouring relationships (not shown for co-located software components).

### 2023-05-21 paper-2023-acsos-pulverization.tex
typos, punctuations, language, conciseness, clarity

@@
 Its goal 
- is to provide application designers a way to specify the functional semantics of their software in a deployment-independent way.
+ is to provide application designers with a way to specify the functional semantics of their software in a deployment-independent way.
 %
 To do so,
-the application logic should be designed considering a \emph{logical system},
-namely, a network of \emph{logical} devices forming a network with \emph{arbitrary} topology.
+the application logic should be designed considering a \emph{logical system}:
+a network of \emph{logical} devices forming a network with \emph{arbitrary} topology.
 %

### 2023-05-21 paper-2023-acsos-pulverization.tex
typos, punctuations, language, conciseness, clarity

@@
 %
-Indeed, though logical devices are generally associated to ``application-level'' physical devices that needs to be controlled or monitored (e.g., a drone, a sensor, or a person---a logical ensemble of wearables),
- their execution can also be supported by other physical devices (e.g., purely infrastructural ones).
+Indeed, though logical devices are generally associated with ``application-level'' physical devices that need to be controlled or monitored
+(e.g., a drone, a sensor, or a person---a logical ensemble of wearables),
+their execution can also be supported by other physical devices
+(e.g., purely infrastructural ones).
 

### 2023-05-21 paper-2023-acsos-pulverization.tex
typos, punctuations, language, conciseness, clarity

@@
 %
-Once pulverised, the logical system would be split in ten deployable pulverised components
+Once pulverised, the logical system would be split into ten deployable pulverised components
 (that we indicate with their component symbol and an index, e.g., $\sigma_1$ is the logical sensor of the first device).
 %
-These pulverised components can be deployed in arbitrary hosts,
+These pulverised components can be deployed in arbitrary hosts
 as far as they have the \emph{capabilities} required to host them

### 2023-05-21 paper-2023-acsos-pulverization.tex
typos, punctuations, language, conciseness, clarity

@@
 (which are too weak and energy-critical to execute the logic),
-the valves are controlled by a dedicated device,
+a dedicated device controls the valves,
 the levels must be logged on a database hosted on a cloud server reachable via HTTPS,

### 2023-05-21 paper-2023-acsos-pulverization.tex
typos, punctuations, language, conciseness, clarity

@@
 let us assume, for instance,
-that our actual target infrastructure is not composed of two connected devices,
-but it is an existing \ac{iot} system
+that our actual target infrastructure is not composed of two connected devices
+but is an existing \ac{iot} system
 where the rain gauges are connected to two LoRaWAN motes

### 2023-05-21 paper-2023-acsos-pulverization.tex
typos, punctuations, language, conciseness, clarity

@@
     \item \textbf{runtime}: the system should be able to reconfigure its deployment at runtime without disrupting the application logic; and
-    \item \textbf{language}: for the idea to be exploitable by designers it is important that
+    \item \textbf{language}: for the idea to be exploitable by designers,
+    it is essential that
     the pulverised configuration

### 2023-05-21 paper-2023-acsos-pulverization.tex
typos, punctuations, language, conciseness, clarity

@@
   \item the definition of the \emph{target} deployment domain (\emph{infrastructure}), i.e., the specification of the \emph{requirements} posed on hosts in terms of required (or preferred) capabilities; and
-  \item the definition of the \emph{deployment plan}, i.e., the allocation of the pulverised components to the infrastructural elements, together with constraints and reconfiguration policies.
+  \item the definition of the \emph{deployment plan}, i.e., allocating the pulverised components to the infrastructural elements, together with constraints and reconfiguration policies.
 \end{enumerate}

### 2023-05-21 paper-2023-acsos-pulverization.tex
typos, punctuations, language, conciseness, clarity

@@
 %
-The runtime could be reified into a deployment unit \emph{per se},
+The runtime could be reified into a deployment unit \emph{per se}
  or can be implemented as part of the deployment units of the software system to be deployed---e.g., taking the form of an \emph{\ac{api}} or a \emph{framework}.
@@
 %
-(i) declaratively, through a set of \emph{configuration files};
-(ii) through a \emph{\ac{api}/library}, implemented in a general-purpose language;
+(i) declaratively through a set of \emph{configuration files};
+(ii) through an \emph{\ac{api}/library} implemented in a general-purpose language;
 or 
@@
 %
-These approaches have different trade-offs.
+Naturally, these approaches have different trade-offs.
 

### 2023-05-21 paper-2023-acsos-pulverization.tex
typos, punctuations, language, conciseness, clarity

@@
 %
-A (well designed) library can be flexible as the host language can be used to express peculiar configurations
+On the other hand,
+a (well designed) library can be flexible as the host language can be used to express peculiar configurations
 and easier to adopt to users acquainted with the host language;

### 2023-05-21 paper-2023-acsos-pulverization.tex
typos, punctuations, language, conciseness, clarity

--- 2023-05-21 paper-2023-acsos-pulverization.tex
@@
 %
-A (well designed) library can be flexible as the host language can be used to express peculiar configurations
-and easier to adopt to users acquainted with the host language;
-but its configuration may quickly become imperative and harder to maintain as complexity grows.
+On the other hand,
+a (well-designed) library can be flexible as the host language can be used to express peculiar configurations
+and is easier to adopt for users acquainted with the host language.
+%
+However,
+its configuration may quickly become imperative and harder to maintain as complexity grows.
 %

### 2023-05-21 paper-2023-acsos-pulverization.tex
typos, punctuations, language, conciseness, clarity

@@
 However,
-its configuration may quickly become imperative and harder to maintain as complexity grows.
+its configuration may quickly become imperative and more challenging to maintain as complexity grows.
 %
 Finally, a \ac{dsl}~\cite{dsl-book-voelter} can be designed to be as expressive as needed,
-but it has high maintenance cost (as the language maintenance costs stack upon those of the library/\ac{api}) and, potentially,
-a steep learning curve due to the need of learning a new (custom) language.
+but it has high maintenance cost
+(as the language maintenance stacks upon the library/\ac{api}) and, potentially,
+a steep learning curve due to the need to learn a new (custom) language.
 

### 2023-05-21 paper-2023-acsos-pulverization.tex
typos, punctuations, language, conciseness, clarity

@@
 
-Recent evolution in programming languages,
-however,
-open to an additional strategy:
+However, recent evolution in programming languages opened an additional strategy:
 the construction of \emph{internal} \acp{dsl},

### 2023-05-21 paper-2023-acsos-pulverization.tex
typos, punctuations, language, conciseness, clarity

@@
 However, recent evolution in programming languages opened an additional strategy:
-the construction of \emph{internal} \acp{dsl},
+\emph{internal} \acp{dsl},
 hybrids between libraries and stand-alone (external) \acp{dsl}~\cite{dsl-book-voelter}.

### 2023-05-21 paper-2023-acsos-pulverization.tex
typos, punctuations, language, conciseness, clarity

@@
 %
-In fact, modern languages such as Kotlin, Groovy, Ruby, and Scala~\cite{Riti2018}
+Modern languages such as Kotlin, Groovy, Ruby, and Scala~\cite{Riti2018}
 provide specific syntactic features to enable the construction of APIs whose ergonomy is akin to the one of a dedicated language,

### 2023-05-21 paper-2023-acsos-pulverization.tex
typos, punctuations, language, conciseness, clarity

@@
 %
-Compared to external ones
+Compared to external ones,
 internal \acp{dsl} allow the designer to fall back to the host language directly
-to implement peculiar entities or processes that cannot be express by the \ac{dsl}.
+to implement peculiar entities or processes that cannot be expressed by the \ac{dsl}.
 %

### 2023-05-21 paper-2023-acsos-pulverization.tex
typos, punctuations, language, conciseness, clarity

--- 2023-05-21 paper-2023-acsos-pulverization.tex
@@
 %
-Compared to external ones
+Compared to external ones,
 internal \acp{dsl} allow the designer to fall back to the host language directly
-to implement peculiar entities or processes that cannot be express by the \ac{dsl}.
+to implement peculiar entities or processes that the \ac{dsl} cannot express.
 %
 The approach has thus been successfully applied in several domains,
-spanning from
+from
 build systems\footnote{\url{https://archive.is/5xtaN}}

### 2023-05-21 paper-2023-acsos-pulverization.tex
typos, punctuations, language, conciseness, clarity

@@
 Compared to external ones,
-internal \acp{dsl} allow the designer to fall back to the host language directly
+internal \acp{dsl} allow the designer to directly fall back to the host language
 to implement peculiar entities or processes that the \ac{dsl} cannot express.

### 2023-05-21 paper-2023-acsos-pulverization.tex
typos, punctuations, language, conciseness, clarity

@@
 %
-The DSL provides simple means to define capabilities as types in the host language (Lines~\ref{capab:start}--\ref{capab:end}),
-and then to associate them to pulverised components for each device type (Lines~\ref{devcap:start}--\ref{devcap:end}).
+The DSL provides simple means to define capabilities as types in the host language (Lines~\ref{capab:start}--\ref{capab:end})
+and associate them with pulverised components for each device type (Lines~\ref{devcap:start}--\ref{devcap:end}).
 %

### 2023-05-21 paper-2023-acsos-pulverization.tex
typos, punctuations, language, conciseness, clarity

@@
 This \ac{dsl} is meant to define constraints that the application imposes on the infrastructure,
-ruling out configurations that are not capable to support the system
+ruling out configurations that cannot support the system
 (for instance, deploying the sensor component of a device in a host that exposes no sensor).

### 2023-05-21 paper-2023-acsos-pulverization.tex
typos, punctuations, language, conciseness, clarity

@@
 %
-All combinations of components and capabilities that are defined as supported become instead amenable as \emph{targets} of deployment or reconfiguration---forming what we may call the \emph{deployment range}.
+All combinations of components and capabilities defined as supported become amenable instead as \emph{targets} of deployment or reconfiguration---forming what we may call the \emph{deployment range}.
 %\emph{runtime reconfiguration}.

### 2023-05-21 paper-2023-acsos-pulverization.tex
typos, punctuations, language, conciseness, clarity

@@
 %
-In the example, e.g., component \texttt{Behavior} in \texttt{iot-sensor} is defined as executable on any of host exposing
+In the example, e.g., component \texttt{Behavior} in \texttt{iot-sensor} is defined as executable on any host exposing
 \texttt{HighCPU} and/or \texttt{EmbeddedDevice} as capabilities (\Cref{devcap:b}):
@@
 
-The \dslDep{} supports the definition of the mapping between pulverised components and specific hosts,
-as well as the definition of reactive reconfiguration policies.
+The \dslDep{} supports the definition of the mapping between pulverised components and specific hosts
+and the definition of reactive reconfiguration policies.
 %
 It works with the concepts of \texttt{Host}
-(associated to \texttt{Capability}),
+(associated with \texttt{Capability}),
 \texttt{ReconfigurationEvent},
@@
 %
-Similarly, hosts can be configured by defining their name and capabilities (Lines~\ref{code:hosts:start}--\ref{code:hosts:end}).
+Similarly, hosts can be configured by defining their names and capabilities (Lines~\ref{code:hosts:start}--\ref{code:hosts:end}).
 %

### 2023-05-21 paper-2023-acsos-pulverization.tex
typos, punctuations, language, conciseness, clarity

@@
 in fact,
-plain lambda expressions, supporting any kind of computation.
+plain lambda expressions supporting any kind of computation.
 

### 2023-05-21 paper-2023-acsos-pulverization.tex
typos, punctuations, language, conciseness, clarity

@@
 %
-We wanted a language with a strongly typed syntax to help developers with better contextual assist,
+We wanted a language with a strongly typed syntax to help developers with a better contextual assist,
 and we wanted the possibility to target multiple runtimes
@@
 %
-The framework has been open sourced and released with a permissive license\footnote{\url{https://github.com/pulvreakt/pulvreakt}}.
+The framework has been open-sourced and released with a permissive license\footnote{\url{https://github.com/pulvreakt/pulvreakt}}.
 

### 2023-05-21 paper-2023-acsos-pulverization.tex
typos, punctuations, language, conciseness, clarity

@@
 the \texttt{Spawner}, which is responsible for executing the components;
-the \texttt{UnitReconfigurator}, which is responsible for reconfigure the \emph{deployment unit}
+the \texttt{UnitReconfigurator}, which is responsible for reconfiguring the \emph{deployment unit}
 according to the \texttt{ReconfigurationEvent}

### 2023-05-21 paper-2023-acsos-pulverization.tex
typos, punctuations, language, conciseness, clarity

--- 2023-05-21 paper-2023-acsos-pulverization.tex
@@
 the \texttt{Spawner}, which is responsible for executing the components;
-the \texttt{UnitReconfigurator}, which is responsible for reconfigure the \emph{deployment unit}
+the \texttt{UnitReconfigurator}, which is responsible for reconfiguring the \emph{deployment unit}
 according to the \texttt{ReconfigurationEvent}
-and the \texttt{Communicator}s which are used to establish the components' intra-communication.
+and the \texttt{Communicator}s,
+which are used to establish the components' intra-communication.
 %
 To start, the \texttt{PulverisationRuntime} needs to know the name of the \emph{logical device},
-the \emph{host} on which is running, and the configurations produced by the two \ac{dsl}s.
-%
-At the start of the system the runtime,
+the \emph{host} on which it is running, and the configurations produced by the two \ac{dsl}s.
+%
+At the start of the system, the runtime,
 based on the provided configuration,
@@
 %
-Similarly, the \texttt{UnitReconfigurator} starts observing all the configured \texttt{ReconfigurationEvent}s,
+Similarly, the \texttt{UnitReconfigurator} starts observing all the configured \texttt{ReconfigurationEvent}s
 and registers the listeners for incoming reconfigurations.

### 2023-05-21 paper-2023-acsos-pulverization.tex
typos, punctuations, language, conciseness, clarity

@@
 %
-Each spawned component starts to execute its logic and shares data with other component via the \texttt{Communicator},
+Then,
+each spawned component starts
+executing its logic and sharing data with other components via the \texttt{Communicator},
 which is responsible for the components' intra-communication by leveraging the communication protocol.
 %
-When a reconfiguration occurs, the \texttt{UnitReconfigurator} takes into account the new deployment configuration,
-by interacting with the \texttt{Spawner}, to start or stop the required components.
+Finally,
+when a reconfiguration occurs, the \texttt{UnitReconfigurator} considers the new deployment configuration and,
+by interacting with the \texttt{Spawner},
+starts or stops components in order to achieve the new setup.
 

### 2023-05-21 paper-2023-acsos-pulverization.tex
typos, punctuations, language, conciseness, clarity

@@
 
-The configuration produced by the \ac{dsl} needs a platform that instructs the system on the way actual communication needs to be performed.
+The configuration produced by the \ac{dsl} needs a platform that instructs the system on
+how actual communication must be performed.
 %
 In our current implementation,
-we provide out of the box a JVM platform based on the AMQP protocol leveraging RabbitMQ,
+we provide out-of-the-box a JVM platform based on the AMQP protocol leveraging RabbitMQ,
 supporting both intra- and inter-device communication.

### 2023-05-21 paper-2023-acsos-pulverization.tex
typos, punctuations, language, conciseness, clarity

@@
 %
-The platform also features an optimisation for pulverised components of the same logical device ending up running on the same host:
-as it would be expected,
-they would communicate directly by sharing memory,
+The platform also optimises pulverised components of the same logical device ending up running on the same host:
+their communication happens directly by sharing memory
 without the need to go through (de)serialisation processes and networking.

### 2023-05-21 paper-2023-acsos-pulverization.tex
typos, punctuations, language, conciseness, clarity

@@
 %
-For the large scale experiment,
+For the large-scale experiment,
 we implemented a second platform,
@@
 %
-The implementation required little effort,
-and it has been made available as part of the evaluation repositories.
+The implementation required little effort
+and has been made available as part of the evaluation repositories.
 

### 2023-05-21 paper-2023-acsos-pulverization.tex
typos, punctuations, language, conciseness, clarity

@@
 In the former,
-we show a small-scale demonstrator of a safety system warning people when a room capacity is exceeded,
+we show a small-scale demonstrator of a safety system warning people when a room capacity is exceeded
 to showcase the integration of the framework with established technologies\footnote{
@@
 in the latter,
-we simulate a city-scale distributed application deployed for a urban event
-and show how the dynamic reconfiguration may help achieving balance between energy consumption and \ac{qos}\footnote{
+we simulate a city-scale distributed application deployed for an urban event
+and show how the dynamic reconfiguration may help to balance
+cost, energy consumption and \ac{qos}\footnote{
     \url{https://github.com/nicolasfara/ACSOS-2023-pulverization-evaluation}

### 2023-05-21 paper-2023-acsos-pulverization.tex
typos, punctuations, language, conciseness, clarity

@@
 %
-Both experiments have been released with a permissive open-source license and have been archived for future reference on
+Both experiments were released with a permissive open-source license and archived for future reference on
 Zenodo~\cite{https://doi.org/10.5281/zenodo.7933160, https://doi.org/10.5281/zenodo.7948801}.
@@
 In particular,
-we want to avoid situations in which too many people are too close to each other in the same room for safety reasons.
+we want to avoid situations where too many people are too close to each other in the same room for safety reasons.
 %
 To do so,
-we equip each person with a wearable device that is able to detect the distance from other similar devices via Bluetooth.
+we equip each person with a wearable device that can detect the distance from other similar devices via Bluetooth.
 %

### 2023-05-21 paper-2023-acsos-pulverization.tex
typos, punctuations, language, conciseness, clarity

@@
 For the sake of simplicity,
-in the experiment we relied on smartphones to emulate wearables,
+we relied on smartphones to emulate wearables,
 and we used
@@
     $$d=10^{\frac{R_{ref} - R}{10 \cdot n}}$$
-    where $R_{ref}$ is the \ac{rssi} reference value at 1 meter
+    where $R_{ref}$ is the \ac{rssi} reference value at 1 meter,
 %    (calibrated in a test environment by measuring it with two smartphones),

### 2023-05-21 paper-2023-acsos-pulverization.tex
typos, punctuations, language, conciseness, clarity

@@
         graph dots represent wearables,
-        and arcs lenghts represent the distance between them.
+        and arc lengths represent the distance between them.
         %
@@
         %
-        The closer the wearables are, the higher is the alert level
+        The closer the wearables are, the higher the alert level
         (reddish tones).

### 2023-05-21 paper-2023-acsos-pulverization.tex
typos, punctuations, language, conciseness, clarity

@@
 
-The system is composed of two kinds of device: \emph{wearable} and \emph{alarm}.
+The system is composed of two kinds of devices: \emph{wearable} and \emph{alarm}.
 %
@@
 %
-They do not need state nor actuators.
+They do not need a state or actuators.
 %
@@
 $\chi^a$, responsible for receiving the data from the wearable devices;
-$\alpha^a$, enacting the alarm by showing a color on the screen.
+$\alpha^a$, enacting the alarm by showing a colour on the screen.
 %
@@
     \item a \emph{Monitor} used to show the alarm;
-    \item a \emph{Server} used run crowd estimation and wearable logic.
+    \item a \emph{Server} used to run crowd estimation and wearable logic.
 \end{itemize}

### 2023-05-21 paper-2023-acsos-pulverization.tex
typos, punctuations, language, conciseness, clarity

@@
 During the experiment,
-we launch several heavy-duty processes on the \emph{Server},
-and we observe how the system reacts to the overload.
+we launch several heavy-duty processes on the \emph{Server}
+and observe how the system reacts to the overload.
 %
 We rely on the MQTT platform module for all the communication,
-meaning that both intra- and inter-device communication is handled by the same protocol and broker.
+meaning that the same protocol and broker handle both intra- and inter-device communication.
 
@@
 %
-The reconfiguration of the system in entirely managed by the framework,
+The reconfiguration of the system is entirely managed by the framework,
 which handles all the machinery and communication to move the component from

### 2023-05-21 paper-2023-acsos-pulverization.tex
typos, punctuations, language, conciseness, clarity

@@
 %
-For the sake of brevity,
+For brevity,
 we do not report the entire configuration in this manuscript, but it is available in the companion artifact~\cite{https://doi.org/10.5281/zenodo.7933160}

### 2023-05-21 paper-2023-acsos-pulverization.tex
typos, punctuations, language, conciseness, clarity

@@
 %
-For brevity,
-we do not report the entire configuration in this manuscript, but it is available in the companion artifact~\cite{https://doi.org/10.5281/zenodo.7933160}
+We do not report the entire configuration in this manuscript for brevity,
+but it is available in the companion artefact~\cite{https://doi.org/10.5281/zenodo.7933160}
 % \meta{

### 2023-05-21 paper-2023-acsos-pulverization.tex
typos, punctuations, language, conciseness, clarity

@@
 %
-To simplify testing we also provide,
-in the same artifact,
+To simplify testing,
+we also provide,
+in the same artefact,
 a simulation of the system where every device is containerised.
 
-\subsection{Large-Scale Scenario: Collective Computation at a Urban Event}
+\subsection{Large-Scale Scenario: Collective Computation at an Urban Event}
 

### 2023-05-21 paper-2023-acsos-pulverization.tex
typos, punctuations, language, conciseness, clarity

@@
 \includegraphics[width=\columnwidth]{figures/pulverization-framework-overview.drawio.pdf}
-\caption{The approach provides two \acp{dsl} and a runtime for specifying and executing deployment plans.
-Each deployment unit will be delivered to the target site and properly configured to interact with other deployment units, as required by the application at hand (see  \Cref{sec:evaluation} for examples).}
+\caption{
+    The approach provides two \acp{dsl}
+    and a runtime for specifying and executing deployment plans.
+    %
+    Each deployment unit will be delivered to the target site
+    and properly configured to interact with other deployment units
+    as the application at hand requires (see  \Cref{sec:evaluation} for examples).
+}
 \label{fig:approach}

### 2023-05-21 paper-2023-acsos-pulverization.tex
typos, punctuations, language, conciseness, clarity

@@
 but its flexibility is limited,
-as options that have not been accounted for at design time can hardly be injected through configuration files.
+as options not accounted for at design time can hardly be injected through configuration files.
 %
@@
 %
-Compared to external ones,
+Unlike external ones,
 internal \acp{dsl} allow the designer to directly fall back to the host language

### 2023-05-21 paper-2023-acsos-pulverization.tex
promote @nicolasfara's meta, tag @metaphori for reference and double check

@@
 %
-\meta{
+%\meta{
     The main goal of the \dslDep{} is to configure the \emph{deployment unit},
@@
     and the \emph{logical device} they represent.
-}
+%}
 %
@@
 we consider a large-scale scenario where a crowd of people is attending an event in a city
-to participate in a collective activity whose computational weight is relevant, for instance
+to participate in a collective activity whose computational weight is relevant,
+for instance,
 a tournament of a geo-located game similar to Ingress\footnote{\url{https://www.ingress.com/}}
@@
 %
-We suppose the application to be pulverised,
-and its behaviour to be able to run either on the smartphones or in cloud.
-%
-We assume the event to start with users located in random positions in the city,
-and to have a recharging station available at that place.
-%
-The game requires to physically move in the city,
-reach a \ac{poi} and perform some operations once there.
+We suppose the application to be pulverised
+and its behaviour to be able to run either on smartphones or in the cloud.
+%
+We assume the event starts with users in random positions in the city,
+where their respective recharging station is located.
+%
+The game requires physically moving into the city,
+reaching a \ac{poi} and performing some operations there.
 %
@@
 %
-The application is pulverised and its behaviour can execute either on the device or in the cloud.
+The application is pulverised,
+and its behaviour can execute either on the device or in the cloud.
 %

### 2023-05-21 paper-2023-acsos-pulverization.tex
promote @nicolasfara's meta, tag @metaphori for reference and double check

@@
 We thus indicate any configuration with $\Updownarrow^{\upsilon}_{\lambda}$,
-with the exception of two cases:
+except for two cases:
 $\upsilon, \lambda > 100$, forcing the behaviour to be always in the cloud, that we indicate with $\Uparrow$;

### 2023-05-21 paper-2023-acsos-pulverization.tex
promote @nicolasfara's meta, tag @metaphori for reference and double check

@@
     \item $P_{devices}$ ($kW$), the average device power consumption, a proxy for the battery life;
-    \item $P_{cloud}$ ($kW$), the average power consumption of the cloud, if added to $P_{devices}$
+    \item $P_{cloud}$ ($kW$), the average power consumption of the cloud, if added to $P_{devices}$,
     provides the total power consumption
@@
 and we thus decided to estimate the \ac{epi}~\cite{DBLP:conf/islped/ShaoB13}
-of common smartphone ($EPI_{device}$) and server ($EPI_{cloud}$) CPUs.
+of a typical smartphone ($EPI_{device}$) and server ($EPI_{cloud}$) CPUs.
 %
@@
 and an
-Intel Xeon Platinum P-8124 ($220W$)
+Intel Xeon Platinum P-8124 ($220W$),
 and we divide it by the score the CPU obtained in the popular Passmark benchmark\footnote{\url{https://www.cpubenchmark.net/}}
@@
 We then used these ratios to estimate the relative \ac{epi} of the two CPUs,
-obtaining a $EPI_{ratio}$ close to 1:18.
+obtaining an $EPI_{ratio}$ close to 1:18.
 %
@@
 we assume the application to be a heavy-duty game,
-capable to drain the battery of a smartphone
+capable of draining the battery of a smartphone
 (approximately 4500mAh in capacity, and thus, at 3.3V, 14.85Wh, 53460J)

### 2023-05-21 paper-2023-acsos-pulverization.tex
typos, punctuations, language, conciseness, clarity

@@
 \subsubsection{Cost model}
-We assume the cloud to be composed by a set of Amazon \emph{AWS} instances,
-specifically, \emph{m5.16xlarge} instances
+We assume the cloud to be composed of a set of Amazon \emph{AWS} instances,
+precisely, \emph{m5.16xlarge} instances
 (using our reference server CPU)
 the \emph{AWS}
-with a hourly cost of \$3.584\footnote{\url{https://aws.amazon.com/ec2/pricing/on-demand/}}
+with an hourly cost of \$3.584\footnote{\url{https://aws.amazon.com/ec2/pricing/on-demand/}}.
 %

### 2023-05-21 paper-2023-acsos-pulverization.tex
typos, punctuations, language, conciseness, clarity

@@
 All the hybrid configurations
-(all except $\Uparrow$ and $\Downarrow$)
-require runtime reconfiguration to be realised.
-%
-For each experiment, we run a total of 10 repetitions with a different random seed;
+(every configuration except $\Uparrow$ and $\Downarrow$)
+require runtime reconfiguration.
+%
+We run 10 repetitions or each experiment with a different random seed;
 all results are averaged over them.

### 2023-05-21 paper-2023-acsos-pulverization.tex
typos, punctuations, language, conciseness, clarity

@@
 %
-The experiment has been open sourced
+The experiment has been open-sourced,
 and the code has been archived on Zenodo~\cite{https://doi.org/10.5281/zenodo.7948801}.
@@
         Smartphones are represented as coloured circles,
-        with the color indicating their battery level
+        with the colour indicating their battery level
         (green when charged, red when low).

### 2023-05-21 paper-2023-acsos-pulverization.tex
typos, punctuations, language, conciseness, clarity

@@
     \caption{
-        Simulation results for 100 (top), 300 (center), and 1000 (bottom) devices;
+        Simulation results for 100 (top), 300 (centre), and 1000 (bottom) devices;
         showing all four metrics:
@@
         %
-        $P_{cloud}$ and $P_{devices}$ are stacked to also show the overall power consumption.
+        $P_{cloud}$ and $P_{devices}$ are stacked to show the overall power consumption.
         %
@@
         %
-        Runtime reconfiguration of the deployment opens to more opportunities to find the trade-off of cost,
-        energy use, and \ac{qos} that best fits the current context.
+        Runtime reconfiguration of the deployment opens more opportunities to find the
+        cost, energy use, and \ac{qos} trade-off that best fits the current context.
     }
@@
 %
-$\Downarrow$ has the worst \ac{qos}, but also the lowest cloud cost and overall energy consumption.
+$\Downarrow$ has the worst \ac{qos}
+but also the lowest cloud cost and overall energy consumption.
 %
@@
 by moving the behaviour from the cloud to the smartphones and vice versa
-depending on the battery level allows to find different trade-offs between these two extremes.
+depending on the battery level allows us to find different trade-offs between these two extremes.
 

### 2023-05-21 paper-2023-acsos-pulverization.tex
typos, punctuations, language, conciseness, clarity

@@
 Therefore, it can be framed in the context of \emph{component-based software engineering},
- with ~\cite{vale2016component-based-se} providing a comprehensive review.
+with~\cite{vale2016component-based-se} providing a comprehensive review.
 %
@@
 Specifically,
- this work builds on the \emph{pulverisation} component model~\cite{FI2020-pulverization,IEEE-IoTJ-pulverization-simulation}.
+ this work builds on the \emph{pulverisation} component model~\cite{FI2020-pulverization,IEEE-IoTJ-pulverization-simulation},
  whereby a large-scale cyber-physical system
@@
 Languages have been proposed
- targeting specific kinds of deployment domains
+ targeting specific deployment domains
  such as the smart grid (cf. \emph{dspec} in the \emph{RIAPS} platform~\cite{DBLP:conf/coins/GhoshTKKL22})
@@
 A related \ac{dsl} is 
-\emph{MuScADeL}~\cite{DBLP:conf/compsac/BoujbelRLTAL14/muscadel}: allows to express deployment properties of 
+\emph{MuScADeL}~\cite{DBLP:conf/compsac/BoujbelRLTAL14/muscadel}:
+it allows to express deployment properties of
 applications on multi-scale deployment domains,
@@
 %
-Component descriptions can include \emph{constraints} (both basic software/hardware constraint and multi-scale criteria).
-%
-The specifications are used to generate \emph{probe} artifacts supporting data collection at the level of the deployment middleware.
+Component descriptions can include \emph{constraints}
+(both basic software/hardware constraints and multi-scale criteria).
+%
+The specifications are used to generate \emph{probe} artefacts supporting data collection at the level of the deployment middleware.
 %
@@
 %
-Indeed, formal models can help in specifying reconfigurable architectures: examples include 
+Indeed, formal models can help specify reconfigurable architectures: examples include
 \emph{DR-BIP}~\cite{Ballouli18dr-bip}
@@
  feature \emph{components} (capturing behaviour),
-\emph{connectors} (capturing interaction between components' \emph{ports}),
+\emph{connectors} (capturing the interaction between components' \emph{ports}),
 \emph{maps} (logical topologies),
@@
 
-A number of approaches use languages together with middlewares to specify and execute reconfiguration policies.
+Several approaches use languages together with middlewares to specify and execute reconfiguration policies.
 %
@@
 %
-Similar, earlier approaches based on the specification of \emph{constraint-sets} include \emph{DELADAS}~\cite{Dearle2004deladas}
+Similarly,
+earlier approaches based on the specification of \emph{constraint-sets} include \emph{DELADAS}~\cite{Dearle2004deladas}
 and \emph{ConfSolve}~\cite{DBLP:conf/saso/HewsonAG13}.
@@
  with constructs for specifying adaptive behaviour 
- and policies (constraints) for the reconfiguration of system components.
+ and policies (constraints) for reconfiguring system components.
 %
@@
 %
-The approach based on the \emph{GeneSIS} modelling language,
+The approach is based on the \emph{GeneSIS} modelling language,
 where a \texttt{DeploymentModel} consists of sets of \texttt{Resources} (e.g., \texttt{Component}s, specialised by \texttt{InfrastructureComponent}s and \texttt{SoftwareComponent}s) with associated \texttt{Property}s and \texttt{Link}s.
 %
-These approaches, however, being based on constraint solving, typically suffer from scalability issues.
+However,
+these approaches, being based on constraint solving, typically suffer from scalability issues.
 
@@
  the edge-cloud infrastructure (the solution),
- passing through layer boundaries (the semi-permeable membranes),
+ passing through layer boundaries (the semi-permeable membranes)
  in order to keep a balance in the desired properties (the solute).
@@
 
-In this work we presented a practical \ac{dsl} and framework for the
+In this work,
+we presented a practical \ac{dsl} and framework for the
 runtime re-configuration of pulverised \ac{cas} applications.
@@
 %
-Through the specialised \ac{dsl} introduced in this work
-we have shown that this idea can be realised in practice.
-%
-Moreover, we extended the original idea of pulverisation with the possibility to specify runtime re-configuration policies,
-which can be leveraged to achieve better trade-offs in terms of performance, cost, and resource usage.
-%
-We provide a real-world demonstrator for the technology,
-and we show via simulation that the approach can be beneficial.
+The specialised \ac{dsl} introduced in this work
+shows that this idea can be realised in practice.
+%
+Moreover, we extended the original idea of pulverisation with the possibility of specifying runtime re-configuration policies,
+which can be leveraged to achieve better performance, cost, and resource usage trade-offs.
+%
+Finally, we provide a real-world demonstrator for the technology
+and show via simulation that the approach can be beneficial.
 
@@
 %
-A policy that requires to consider the state of multiple hosts would require to resort to the host language primitive mechanisms and make explicit communications outside the framework,
+A policy that requires to consider the state of multiple hosts would require to resort to the host language's primitive mechanisms and make explicit communications outside the framework,
 thus breaking the abstraction.
@@
 This limitation can be addressed by extending the current reconfiguration
-block of the \ac{dsl} in such a way that queries can be performed on any reachable host of the system,
+block of the \ac{dsl} so that queries can be performed on any reachable host of the system,
 thus making all communications
@@
 Concurrently with these extensions,
-we intend to add support for additional network protocols to be able to support a larger number of practical scenarios.
+we intend to add support for additional network protocols to support a larger number of practical scenarios.
 

### 2023-05-21 paper-2023-acsos-pulverization.tex
fix some typos

@@
 To evaluate \ourframework{}, we exercise it through two case studies:
-a small-scale real-wold demonstration, and a city-scale simulated scenario.
+a small-scale real-world demonstration, and a city-scale simulated scenario.
 
@@
 concerned with the mapping of \emph{software components} to \emph{physical machines}~\cite{DBLP:journals/software/Kruchten95}
- and supported by modelling notations like \emph{UML Deployment diagrams}.
+ and supported by modeling notations like \emph{UML Deployment diagrams}.
 %
@@
 Finally, a \ac{dsl}~\cite{dsl-book-voelter} can be designed to be as expressive as needed,
-but it has high maintenance cost
+but it has a high maintenance cost
 (as the language maintenance stacks upon the library/\ac{api}) and, potentially,
@@
         This code snippet defines three capabilities and a logical system composed of two device types,
-        whose components are bound to the set of capabilities they needs to execute.
+        whose components are bound to the set of capabilities they need to execute.
     },
@@
 Once the \acp{epi} of the two platforms are known,
-we estimate the number of instructions of the pulverised components as follows:
+we estimate the number of instructions for the pulverised components as follows:
 we assume the application to be a heavy-duty game,
@@
 %
-We run 10 repetitions or each experiment with a different random seed;
+We run 10 repetitions of each experiment with a different random seed;
 all results are averaged over them.
@@
     \caption{
-        Simulation results for 100 (top), 300 (centre), and 1000 (bottom) devices;
+        Simulation results for 100 (top), 300 (center), and 1000 (bottom) devices;
         showing all four metrics:

### 2023-05-21 paper-2023-acsos-pulverization.tex
Minors

@@
 
+%\usepackage[none]{hyphenat}
+
 \acrodef{api}[API]{application-program interface}
@@
 \emph{runtime configuration rules}, allowing pulverised components to move at runtime across different hosts;
- \item we accordigly provide a practical implementation of pulverisation (with reconfiguration rules) called \emph{\ourframework{}}, developed in the Kotlin programming language.
+ \item we accordingly provide a practical implementation of pulverisation (with reconfiguration rules) called \emph{\ourframework{}}, developed in the Kotlin programming language.
 \end{enumerate}
@@
 concerned with the mapping of \emph{software components} to \emph{physical machines}~\cite{DBLP:journals/software/Kruchten95}
- and supported by modeling notations like \emph{UML Deployment diagrams}.
+ and supported by modelling notations like \emph{UML Deployment diagrams}.
 %
@@
 %
-The framework has been open-sourced and released with a permissive license\footnote{\url{https://github.com/pulvreakt/pulvreakt}}.
+The framework has been open-sourced and released with a permissive licence\footnote{\url{https://github.com/pulvreakt/pulvreakt}}.
 
@@
 in the latter,
-we simulate a city-scale distributed application deployed for an urban event
+we simulate a city-scale distributed application deployed for a urban event
 and show how the dynamic reconfiguration may help to balance
@@
 %
-Both experiments were released with a permissive open-source license and archived for future reference on
+Both experiments were released with a permissive open-source licence and archived for future reference on
 Zenodo~\cite{https://doi.org/10.5281/zenodo.7933160, https://doi.org/10.5281/zenodo.7948801}.
@@
 
-\subsection{Large-Scale Scenario: Collective Computation at an Urban Event}
+\subsection{Large-Scale Scenario: Collective Computation at a Urban Event}
 
@@
 %
-Opposedly, $\Uparrow$ improves the \ac{qos}, but costs and energy use grow significantly.
+Conversely, $\Uparrow$ improves the \ac{qos}, but costs and energy use grow significantly.
 %
@@
 \emph{MuScADeL}~\cite{DBLP:conf/compsac/BoujbelRLTAL14/muscadel}:
-it allows to express deployment properties of
+it allows expressing deployment properties of
 applications on multi-scale deployment domains,
@@
 %
-A policy that requires to consider the state of multiple hosts would require to resort to the host language's primitive mechanisms and make explicit communications outside the framework,
+A policy that requires to consider the state of multiple hosts would require resorting to the host language's primitive mechanisms and make explicit communications outside the framework,
 thus breaking the abstraction.

### 2023-05-21 paper-2023-acsos-pulverization.tex
Minors

@@
 val infrastructure = setOf(Smartphone, Server)
-val conf = // see %\crefe{lst:system-configuration}%
+val conf = ... // see %\crefe{lst:system-configuration}%
 pulverizationRuntime(conf, "iot-sensor", infrastructure) {
@@
 the \texttt{UnitReconfigurator}, which is responsible for reconfiguring the \emph{deployment unit}
-according to the \texttt{ReconfigurationEvent}
+according to the \texttt{ReconfigurationEvent}; 
 and the \texttt{Communicator}s,
@@
 %
-At the start of the system, the runtime,
-based on the provided configuration,
-spawns all the components that must start on the \emph{deployment unit}.
+At the start of the system, using the provided configuration, the runtime
+spawns all the components that must execute on the \emph{deployment unit}.
 %

### 2023-05-21 paper-2023-acsos-pulverization.tex
Minors

@@
 In the former,
-we show a small-scale demonstrator of a safety system warning people when a room capacity is exceeded
+we show a small-scale demonstrator of a safety system warning people when a room capacity is exceeded, 
 to showcase the integration of the framework with established technologies\footnote{
@@
 and show how the dynamic reconfiguration may help to balance
-cost, energy consumption and \ac{qos}\footnote{
+cost, energy consumption, and \ac{qos}\footnote{
     \url{https://github.com/nicolasfara/ACSOS-2023-pulverization-evaluation}
@@
 %
+These tools can support \emph{reasoning} about reconfigurations and \emph{verification}.
+
 %\meta{TODO: say something about our positioning wrt verification of reconfigurations}

### 2023-05-22 paper-2023-acsos-pulverization.tex
A clear statement in RW about difference
larger code with little decrease in char spacing

@@
 \lstdefinelanguage{pulv}[]{Kotlin}{
-%  basicstyle=\small\ttfamily\lst@ifdisplaystyle\footnotesize\fi,
+ basicstyle=\small\ttfamily\lst@ifdisplaystyle\footnotesize\fi,
 %	frame=single,
@@
 \lstset{language=pulv,
-basewidth=0.555em % decreases spacing between chars
+basewidth=0.52em % decreases spacing between chars
 }
@@
  we are especially interested in deployments over large-scale cyber-physical systems and the edge-cloud continuum.
-
-A related \ac{dsl} is 
+%
+The main aspects distinguishing \ourframework{} from other approaches are that it targets \emph{pulverised systems} (and not general partitionings) and that it uses internal \acp{dsl} embedded in Kotlin.
+
+A related (external) \ac{dsl} is 
 \emph{MuScADeL}~\cite{DBLP:conf/compsac/BoujbelRLTAL14/muscadel}:

### 2023-05-22 paper-2023-acsos-pulverization.tex
Fig on approach: minor fixes + larger text

@@
     Each deployment unit will be delivered to the target site
-    and properly configured to interact with other deployment units
-    as the application at hand requires (see  \Cref{sec:evaluation} for examples).
+    and properly configured to interact with other deployment units according to the protocal 
+    that the application at hand requires (see  \Cref{sec:evaluation} for examples).
 }

### 2023-05-22 paper-2023-acsos-pulverization.tex
Increase char spacing in sources + readjust a bit

@@
 \lstset{language=pulv,
-basewidth=0.52em % decreases spacing between chars
+basewidth=0.55em % decreases spacing between chars
 }
@@
   override val hostname = "amazon-aws"
-  override val capabilities=setOf(HighCPU,LowLatencyComm)
+  override val capabilities = 
+    setOf(HighCPU,LowLatencyComm)
 }%\label{code:hosts:end}%
-// Runtime initial setup and runtime reconfiguration rules %\label{code:infra:start}%
+// Runtime initial setup and runtime reconfig. rules %\label{code:infra:start}%
 val infrastructure = setOf(Smartphone, Server)
 val conf = ... // see %\crefe{lst:system-configuration}%
-pulverizationRuntime(conf, "iot-sensor", infrastructure) {
+pulverizationRuntime(conf,"iot-sensor",infrastructure){
   DeviceBehaviour() startsOn Server%\label{code:mapping:start}%
@@
     onDevice {
-      HighLoad reconfigures {Behaviour movesTo Smartphone}
-      LowBattery reconfigures {Behaviour movesTo Server}
+      HighLoad reconfigures 
+        { Behaviour movesTo Smartphone }
+      LowBattery reconfigures 
+        { Behaviour movesTo Server }
     }

### 2023-05-22 paper-2023-acsos-pulverization.tex
RW: minors

@@
 %
+The idea of its \ac{dsl} to reconfiguration is to let designers specify behaviours that regulate the transitions between explicitly specified \emph{configurations}.
 %
@@
 
-Another approach is \emph{osmotic computing}~\cite{DBLP:journals/computer/VillariFDRJR19},
+Finally, there is a plethora of approaches for automatic deployment or offloading across the edge-cloud continuum. One example is \emph{osmotic computing}~\cite{DBLP:journals/computer/VillariFDRJR19},
  whereby microservices (the solvent)
